---
title: "ADNIMERGE2-Analysis-Data"
output: 
   rmarkdown::html_vignette:
     toc: true
     code_folding: hide
date: "Last Updated: `r format(Sys.Date(), '%B %d, %Y')`"
vignette: >
  %\VignetteIndexEntry{ADNIMERGE2-Analysis-Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  class.source = "fold-show"
)
options(rmarkdown.html_vignette.check_title = FALSE)
# Data print function
datatable <- function(data, paging = FALSE, searchable = TRUE, bInfo = FALSE, ...) {
  DT::datatable(
    data = data, ...,
    options = list(
      paging = paging, searchable = searchable, bInfo = bInfo,
      ...
    )
  )
}
```

# Introduction

This article describes creating derived analysis ready dataset using the [PHARMERVERSE](https://pharmaverse.org/) workflow. It uses some pre-generated standardized dataset and a metadata-specs as input. Please refer to the following articles to see how those derived dataset and metadata-specs were created, respectively. 

* [ADNIMERGE2-Derived-Data]()
* [ADNIMERGE2-Analysis-Meta-Specs]()

In `ADNIMERGE2` R data package, the following selected derived analysis ready dataset will be created for illustration purpose. 

* Subject-Level Analysis Dataset - ADSL
* Analysis Dataset of Adverse Events - ADAE
* Analysis Dataset of Questionnaire - ADQS
* Analysis Dataset of Clinical Classification - ADRS

**NOTE** 

* *These derived analysis ready dataset may not be fully complied with the [CDISC-ADaM](https://www.cdisc.org/standards/foundational/adam) standardization.*

# Load Required R Packages

```{r setup-libraries}
library(tidyverse)
library(assertr)
library(metacore)
library(metatools)
library(admiral)
library(admiraldev)
library(xportr)
```

```{r study-library, purl = FALSE}
# ADNI study R data package
library(ADNIMERGE2)
```

The following r chuck loads some of study specific wrapper functions that are modified from [`metacore`](https://atorus-research.github.io/metacore/) and [`metatools`](https://pharmaverse.github.io/metatools/) R packages and stored in the `ADNIMERGE2` R package as system file.

```{r system-utils-summary, echo = FALSE}
bind_rows(
  c(
    fun_name = "convert_metacore_to_tibble",
    source = "metacore",
    text = "Convert R6-wrapper metacore object to tibble form"
  ),
  c(
    fun_name = "single_build_from_derived",
    source = "metatools",
    text = "Build from single derived data that required multiple datasets as input"
  ),
  c(
    fun_name = "convert_var_to_fct_wrapper",
    source = "metatools",
    text = "Convert multiple variables into factor - a wrapper function of `convert_var_to_fct`"
  ),
  c(
    fun_name = "create_amystatus_cols",
    source = "internal",
    text = "Create a common baseline amyloid status from multiple image tracer and resolutions"
  )
) %>%
  select(
    `Internal Function` = fun_name,
    `Package Source` = source,
    `Description` = text
  ) %>%
  datatable(searchable = FALSE, bInfo = FALSE)
```

```{r system-utils, purl = FALSE, echo = TRUE, eval = TRUE}
# Load utils function from package system file
utils_file_path <- system.file(
  "analysis-dataset-utils.R",
  package = "ADNIMERGE2",
  mustWork = TRUE
)
source(utils_file_path)
```

# Building Derived Analysis Ready Dataset

```{r prep-meta-data, echo = FALSE}
# Convert `R6-class` object into a data.frame
meta_data_dic <- convert_metacore_to_tibble(METACORES) %>%
  mutate(
    TYPE = str_to_upper(string = type),
    DERIVED = TRUE,
    FLDNAME = variable
  ) %>%
  select(
    TBLNAME = dataset, CRFNAME = dataset.label, structure,
    FLDNAME, LABEL = label, TYPE, DERIVED, type, variable
  )
```

## Subject-Level Analysis Dataset - ADSL

[`ADSL`](../reference/ADSL.html) dataset contains one record per subject. In the ADNI study, [`ADSL`](../reference/ADSL.html) dataset includes only the first record/baseline characteristics of subject when they were enrolled in the study as newly-enrollee or screened to the study for the first time.

<details>
<summary>Show `ADSL` Variables</summary>

```{r adsl-data-dic-print, class.source = NULL, echo = FALSE}
adsl_data_dic <- meta_data_dic %>%
  filter(TBLNAME == "ADSL")

adsl_data_dic %>%
  select(FLDNAME, LABEL, TYPE) %>%
  datatable(data, paging = TRUE)
```

</details>

```{r adsl-data-dic, echo = FALSE}
adsl_metacore <- METACORES %>%
  select_dataset(dataset = "ADSL", simplify = FALSE)
```

```{r generate-adsl0}
ADSL <- build_from_derived(
  metacore = adsl_metacore,
  ds_list = list("DM" = DM, "GF" = GF),
  predecessor_only = TRUE,
  keep = "ALL"
) %>%
  assert_uniq(USUBJID) %>%
  # Create AGE Group - creating categorical group from numeric
  create_cat_var(
    data = .,
    metacore = adsl_metacore,
    ref_var = AGE,
    grp_var = AGEGR1,
    num_grp_var = AGEGR1N
  ) %>%
  # Adjust for multiple race groups
  mutate(RACE = case_when(
    str_detect(RACE, "\\|") ~ "More than one race",
    str_detect(RACE, "Native Hawaiian$|^Other Pacific Islander") ~ "Native Hawaiian or Other Pacific Islander",
    TRUE ~ as.character(RACE)
  )) %>%
  # Re-code RACE
  create_var_from_codelist(
    data = .,
    metacore = adsl_metacore,
    input_var = RACE,
    out_var = RACEN,
    decode_to_code = TRUE
  ) %>%
  # Add baseline DX
  derive_vars_merged(
    dataset = .,
    dataset_add = RS,
    filter_add = RSTESTCD == "DX" & RSBLFL == "Y",
    new_vars = exprs(DX = RSORRES),
    order = exprs(VISITNUM),
    mode = "last",
    by_vars = exprs(STUDYID, USUBJID),
    check_type = "error",
    relationship = "one-to-one"
  ) %>%
  # Convert to factor with Control Terms (CT)
  convert_var_to_fct_wrapper(
    .data = .,
    metacore = adsl_metacore,
    var = c("ORIGPROT", "DX", "RACE")
  ) %>%
  # Add baseline education level and marital status
  call_derivation(
    dataset = .,
    dataset_add = SC,
    derivation = derive_vars_merged,
    variable_params = list(
      params(
        filter_add = SCTESTCD == "PTEDUCAT" & SCBLFL == "Y",
        new_vars = exprs(EDUC = SCSTRESN)
      ),
      params(
        filter_add = SCTESTCD == "PTMARRY" & SCBLFL == "Y",
        new_vars = exprs(MARISTAT = SCORRES)
      )
    ),
    order = exprs(VISITNUM),
    mode = "first",
    by_vars = exprs(STUDYID, USUBJID),
    check_type = "error",
    relationship = "one-to-one"
  ) %>%
  # Add baseline BMI derived from weight and height
  call_derivation(
    dataset = .,
    dataset_add = VS,
    derivation = derive_vars_merged,
    variable_params = list(
      params(
        filter_add = VSTESTCD == "WEIGHT" & VSBLFL == "Y",
        new_vars = exprs(WEIGHT = VSSTRESN)
      ),
      params(
        filter_add = VSTESTCD == "HEIGHT" & VSBLFL == "Y",
        new_vars = exprs(HEIGHT = VSSTRESN)
      )
    ),
    order = exprs(VISITNUM),
    mode = "first",
    by_vars = exprs(USUBJID),
    check_type = "error",
    relationship = "one-to-one"
  ) %>%
  mutate(
    BMI = round(compute_bmi(height = HEIGHT, weight = WEIGHT), 2),
    BMIBLU = ifelse(!is.na(BMI), "kg/m^2", NA_character_),
    AGEU = ifelse(!is.na(AGEU), "Years", NA_character_)
  )
```


```{r generate-adsl1}
# Add baseline amyloid status
NV_AMYSTAT_COLS <- NV %>%
  select(NVTESTCD, NVTEST, NVCAT, NVSCAT) %>%
  distinct() %>%
  filter(str_detect(NVTESTCD, "AMYSTAT")) %>%
  mutate(IMAGE_RES = str_extract_all(NVSCAT, "8MM$|6MM$")) %>%
  filter(str_detect(IMAGE_RES, "8|6")) %>%
  mutate(NVTESTCD_NVSCAT = paste0(NVTESTCD, "0", str_replace_all(NVSCAT, "-", "_")))

nv_params <- lapply(NV_AMYSTAT_COLS$NVTESTCD_NVSCAT, function(param) {
  params_cond <- admiral::params(
    filter_add = NVCAT == "AMYLOIDPET" &
      NVSCAT == stringr::str_replace(stringr::str_split(!!param, "0", simplify = TRUE)[2], "_", "-") &
      NVTESTCD == stringr::str_split(!!param, "0", simplify = TRUE)[1] &
      NVBLFL == "Y",
    new_vars = exprs(!!!exprs(param = NVSTRESC))
  )
  params_cond$new_vars <- substitute(
    exprs(param := NVSTRESC),
    list(param = as.name(param))
  )
  return(params_cond)
})

ADSL <- ADSL %>%
  call_derivation(
    dataset = .,
    dataset_add = NV,
    derivation = derive_vars_merged,
    variable_params = nv_params,
    order = exprs(VISITNUM), # For duplicated cases
    mode = "first",
    by_vars = exprs(USUBJID),
    check_type = "error",
    relationship = "one-to-one"
  )

# Required to create a common amyloid status variable from
# multiple image tracers and resolutions
ADSL <- ADSL %>%
  create_amystatus_cols(.) %>%
  select(-any_of(NV_AMYSTAT_COLS$NVTESTCD_NVSCAT))
```

```{r generate-adsl2}
# Add baseline questionnaire/assessment score
qs_params <- lapply(unique(QS$QSTESTCD), function(param) {
  params_cond <- admiral::params(
    filter_add = QSTESTCD == !!param & QSBLFL == "Y",
    new_vars = exprs(!!!exprs(param = QSORRES))
  )
  params_cond$new_vars <- substitute(
    exprs(param := as.numeric(QSORRES)),
    list(param = as.name(param))
  )
  return(params_cond)
})

ADSL <- ADSL %>%
  call_derivation(
    dataset = .,
    dataset_add = QS,
    derivation = derive_vars_merged,
    variable_params = qs_params,
    order = exprs(VISITNUM),
    mode = "first",
    by_vars = exprs(USUBJID),
    check_type = "error",
    relationship = "one-to-one"
  )
```

```{r generate-adsl3}
ADSL <- ADSL %>%
  # Create enrollment flag
  derive_var_merged_ef_msrc(
    dataset = .,
    by_vars = exprs(STUDYID, USUBJID),
    flag_events = list(
      flag_event(dataset_name = "DM", condition = !is.na(RFSTDTC))
    ),
    source_datasets = list(DM = DM),
    new_var = ENRLFL,
    true_value = "Y",
    false_value = NA_character_
  ) %>%
  # Imputation for death date and enrollment end date
  # Required partial date/time format/character format
  rename("EOSDTHC" = EOSDT) %>%
  select(-DTHDT) %>%
  mutate(
    ENRLDT = as.Date(ENRLDT),
    EOSDTHC = as.character(EOSDTHC),
    DTHDTC = as.character(DTHDTC)
  ) %>%
  call_derivation(
    dataset = .,
    derivation = derive_vars_dtm,
    variable_params = list(
      params(dtc = DTHDTC, new_vars_prefix = "DTH"),
      params(dtc = EOSDTHC, new_vars_prefix = "EOS")
    ),
    highest_imputation = "M",
    date_imputation = "mid",
    flag_imputation = "auto",
    min_dates = exprs(ENRLDT)
  ) %>%
  derive_vars_dtm_to_dt(
    dataset = .,
    source_vars = exprs(DTHDTM, EOSDTM)
  ) %>%
  # Remove any columns that are not specified in the meta-specs
  drop_unspec_vars(
    dataset = .,
    metacore = adsl_metacore
  ) %>%
  # Add variable labels
  metatools::set_variable_labels(
    data = .,
    metacore = adsl_metacore
  )
```

```{r checks-adsl1}
ADSL <- ADSL %>%
  # Check all variables specified are present and no more
  check_variables(
    data = .,
    metacore = adsl_metacore
  ) %>%
  # Checks all variables with CT only contain values within the CT
  check_ct_data(
    data = .,
    metacore = adsl_metacore,
    na_acceptable = TRUE,
    omit_vars = c("ENRLFL", "DTHFL")
  ) %>%
  # Orders the columns according to the spec
  order_cols(
    data = .,
    metacore = adsl_metacore
  ) %>%
  # Sort the dataset based on key columns
  sort_by_key(
    data = .,
    metacore = adsl_metacore
  ) %>%
  # Check uniqueness of record by key columns
  check_unique_keys(
    data = .,
    metacore = adsl_metacore
  )
```

```{r checks-adsl0}
ADSL <- ADSL %>%
  xportr_metadata(
    .df = .,
    metadata = adsl_metacore,
    domain = "ADSL",
    verbose = "stop"
  ) %>%
  # Check variable type with specs match
  # xportr_type() %>%
  # Assign variable value length from the meta specs
  xportr_length() %>%
  # Assign variable label from meta specs and
  # checks variable label length for max of 40 characters
  xportr_label() %>%
  # Checks variable format
  xportr_format()
```

## Analysis Dataset of Adverse Events - ADAE

[`ADAE`](../reference/ADAE.html) dataset contains one record per adverse events per subject with the following characteristics.

<details>
<summary>Show `ADAE` Variables</summary>

```{r adae-data-dic-print, class.source = NULL, echo = FALSE}
adae_data_dic <- meta_data_dic %>%
  filter(TBLNAME == "ADAE")

adae_data_dic %>%
  select(FLDNAME, LABEL, TYPE) %>%
  datatable(data, paging = TRUE)
```

</details>

```{r adae-data-dic}
adae_metacore <- METACORES %>%
  select_dataset(.data = ., dataset = "ADAE", simplify = FALSE)
```

```{r generate-adae}
# Modified user-defined function
ADAE <- build_from_derived(
  metacore = adae_metacore,
  ds_list = list("ADSL" = ADSL, "AE" = AE),
  predecessor_only = TRUE,
  keep = FALSE
) %>%
  # Only subject that had at least one adverse events experience
  filter(USUBJID %in% c(AE$USUBJID)) %>%
  verify(nrow(.) == nrow(AE)) %>%
  # Add adverse events onset and ended date
  derive_vars_merged(
    dataset = .,
    dataset_add = AE %>%
      select(USUBJID, AETERM, AESEQ, AESTDTC, AEENDTC),
    by_vars = exprs(USUBJID, AETERM, AESEQ),
    new_vars = NULL,
    check_type = "error",
    relationship = "one-to-one"
  ) %>%
  # AESTDTC and AEENDTC required to be a character variable
  mutate(
    ENRLDT = as.Date(ENRLDT),
    AESTDTC = as.character(AESTDTC),
    AEENDTC = as.character(AEENDTC)
  ) %>%
  # Add imputed date death date from ADSL
  derive_vars_merged(
    dataset = .,
    dataset_add = ADSL %>%
      select(USUBJID, STUDYID, DTHDT),
    by_vars = exprs(USUBJID, STUDYID),
    new_vars = NULL,
    check_type = "error",
    relationship = "many-to-one"
  ) %>%
  # Derive analysis start and end date
  call_derivation(
    dataset = .,
    derivation = derive_vars_dtm,
    variable_params = list(
      params(
        dtc = AESTDTC, new_vars_prefix = "AST", highest_imputation = "M",
        date_imputation = "first", time_imputation = "first"
      ),
      params(
        dtc = AEENDTC, new_vars_prefix = "AEN", highest_imputation = "M",
        date_imputation = "last", time_imputation = "last",
        max_dates = exprs(DTHDT)
      )
    ),
    flag_imputation = "auto",
    min_dates = exprs(ENRLDT)
  ) %>%
  # Convert into date format
  derive_vars_dtm_to_dt(
    dataset = .,
    source_vars = exprs(ASTDTM, AENDTM)
  ) %>%
  # Derive analysis start/end relative day and
  derive_vars_dy(
    reference_date = ENRLDT,
    source_vars = exprs(ASTDT, AENDT)
  ) %>%
  # Derive analysis duration (value and unit)
  derive_vars_duration(
    new_var = ADURN,
    new_var_unit = ADURU,
    start_date = ASTDT,
    end_date = AENDT,
    in_unit = "days",
    out_unit = "days",
    add_one = TRUE,
    trunc_out = FALSE
  ) %>%
  drop_unspec_vars(
    dataset = .,
    metacore = adae_metacore
  ) %>%
  metatools::set_variable_labels(
    data = .,
    metacore = adae_metacore
  )
```

```{r check-adae}
ADAE <- ADAE %>%
  check_variables(
    data = .,
    metacore = adae_metacore
  ) %>%
  # check_ct_data(
  #   data = .,
  #   metacore = adae_metacore,
  #   na_acceptable = TRUE
  # ) %>%
  order_cols(
    data = .,
    metacore = adae_metacore
  ) %>%
  sort_by_key(
    data = .,
    metacore = adae_metacore
  ) %>%
  check_unique_keys(
    data = .,
    metacore = adae_metacore
  ) %>%
  xportr_metadata(
    .df = .,
    metadata = adae_metacore,
    verbose = "stop"
  ) %>%
  # xportr_type() %>%
  xportr_length() %>%
  xportr_label() %>%
  xportr_format()
```

## Analysis Dataset of Questionnaires - ADQS

[`ADQS`](../reference/ADQS.html) dataset contains one record per questionnaire parameter per visit per subject with the following characteristics. 

<details>
<summary>Show `ADQS` Variables</summary>

```{r adqs-data-dic-print, class.source = NULL, echo = FALSE}
adqs_data_dic <- meta_data_dic %>%
  filter(TBLNAME == "ADQS")

adqs_data_dic %>%
  select(FLDNAME, LABEL, TYPE) %>%
  datatable(data, paging = TRUE)
```

</details>

```{r adqs-data-dic}
adqs_metacore <- METACORES %>%
  select_dataset(dataset = "ADQS", simplify = FALSE)
```

```{r adqs-data-prep}
# Assumed all questionnaire parameters are continuous (numeric)
ADQS <- single_build_from_derived(
  metacore = adqs_metacore,
  dataset_name = "ADQS",
  ds_list = list("QS" = QS),
  predecessor_only = TRUE,
  keep = "ALL"
) %>%
  # Add variables from ADSL dataset
  derive_vars_merged(
    dataset = .,
    dataset_add = single_build_from_derived(
      metacore = adqs_metacore,
      dataset_name = "ADQS",
      ds_list = list("ADSL" = ADSL),
      predecessor_only = TRUE,
      keep = "ALL"
    ),
    by_vars = exprs(STUDYID, USUBJID),
    new_vars = NULL,
    check_type = "error",
    relationship = "many-to-one"
  ) %>%
  verify(nrow(.) == nrow(QS)) %>%
  mutate(ADT = as.Date(ADT)) %>%
  # Re-code PARAMN
  # Assumed all PARAMCD are coded in the meta-specs
  create_var_from_codelist(
    data = .,
    metacore = adqs_metacore,
    input_var = PARAMCD,
    out_var = PARAMN,
    decode_to_code = TRUE
  ) %>%
  # Add analysis timing variable
  derive_vars_dy(
    reference_date = ENRLDT,
    source_vars = exprs(ADT)
  ) %>%
  mutate(
    AVISIT = case_when(
      ADT <= ENRLDT | !is.na(ABLFL) ~ "BASELINE",
      TRUE ~ VISIT
    ),
    AVISITN = case_when(
      ADT <= ENRLDT | !is.na(ABLFL) ~ 0,
      TRUE ~ ADY
    )
  ) %>%
  # # Flag baseline records if the flag is not presented in QS derived dataset
  # restrict_derivation(
  #   derivation = derive_var_extreme_flag,
  #   args = params(
  #     by_vars = exprs(STUDYID, USUBJID, PARAMCD),
  #     order = exprs(ADT),
  #     new_var = ABLFL,
  #     mode = "last"
  #   ),
  #   filter = !is.na(AVAL) & ADT <= ENRLDT
  # ) %>%
  # Derive baseline and change from baseline variables
  # Only applicable for continuous parameters
  derive_var_base(
    by_vars = exprs(STUDYID, USUBJID, PARAMCD),
    source_var = AVAL,
    new_var = BASE
  ) %>%
  # Derive change for post-baseline records
  restrict_derivation(
    derivation = derive_var_chg,
    filter = AVISITN > 0
  ) %>%
  # Derive percentage change for post-baseline records
  restrict_derivation(
    derivation = derive_var_pchg,
    filter = AVISITN > 0
  ) %>%
  # Derive sequence number
  derive_var_obs_number(
    by_vars = exprs(STUDYID, USUBJID, PARAMCD),
    order = exprs(PARAMCD, ADT),
    new_var = ASEQ,
    check_type = "none" # "error"
  ) %>%
  drop_unspec_vars(
    dataset = .,
    metacore = adqs_metacore
  ) %>%
  metatools::set_variable_labels(
    data = .,
    metacore = adqs_metacore
  )
```

```{r check-adqs}
ADQS <- ADQS %>%
  check_variables(
    data = .,
    metacore = adqs_metacore
  ) %>%
  check_ct_data(
    data = .,
    metacore = adqs_metacore,
    na_acceptable = TRUE,
    omit_vars = c("ENRLFL", "ABLFL")
  ) %>%
  order_cols(
    data = .,
    metacore = adqs_metacore
  ) %>%
  sort_by_key(
    data = .,
    metacore = adqs_metacore
  ) %>%
  # check_unique_keys(
  #   data = .,
  #   metacore = adqs_metacore
  # ) %>%
  xportr_metadata(
    .df = .,
    metadata = adqs_metacore,
    verbose = "stop"
  ) %>%
  # xportr_type() %>%
  xportr_length() %>%
  xportr_label()
```

### Analysis Dataset of ADAS-Cog Behaviour - ADADAS

Suppose the ADAS Cognitive Behaviors item-13 total score (`ADASTT13`) was the primary outcome in the analysis, then ADAS-Cog Behavior Analysis Dataset (`ADADAS`) can be created either directly from [`ADQS`](../reference/ADQS.html) or with a dataset-specific metadata-specs (i.e. [adding here](../reference/METACORES.html)). The following chuck illustrate how the `ADADAS` can be created directly from pre-generated [`ADQS`](../reference/ADQS.html). More information is available in the [Longitudinal Clinical Cognitive Outcome Summaries]() article.

```{r generate-adadas}
ADADAS <- ADQS %>%
  filter(PARAMCD == "ADASTT13")
```

## Response Analysis Dataset (ADRS) 

[`ADRS`](../reference/ADRS.html) dataset contains subject's clinical diagnostics summary per visits including death records. 

<details>
<summary>Show `ADRS` Variables</summary>

```{r adrs-data-dic-print, class.source = NULL, echo = FALSE}
adrs_data_dic <- meta_data_dic %>%
  filter(TBLNAME == "ADRS")

adrs_data_dic %>%
  select(FLDNAME, LABEL, TYPE) %>%
  datatable(data, paging = TRUE)
```

</details>

```{r adrs-data-dic}
adrs_metacore <- METACORES %>%
  select_dataset(dataset = "ADRS", simplify = FALSE)
```


```{r adrs-data-prep}
# Merge ADSL to RS
adsl_vars <- exprs(STUDYID, USUBJID, ENRLFL, ENRLDT, DTHFL)

ADRS_PREP <- RS %>%
  filter(RSTESTCD %in% "DX") %>%
  # Adjusting BLFL for subjects that have more than one flag
  group_by(STUDYID, USUBJID, RSTESTCD) %>%
  mutate(NUM_BLFL = sum(!is.na(RSBLFL))) %>%
  ungroup() %>%
  mutate(RSBLFL = case_when(
    NUM_BLFL > 1 & VISITNUM != 1 ~ NA_character_,
    TRUE ~ RSBLFL
  )) %>%
  select(-NUM_BLFL) %>%
  mutate(
    PARAMCD = RSTESTCD,
    PARAM = RSTEST,
    AVALC = RSSTRESC,
    AVAL = get_numeric_dx_status(RSSTRESC),
    ABLFL = RSBLFL,
    ADT = as.Date(RSDTC),
    ADY = RSDY,
    COLPROT = RSGRPID
  ) %>%
  derive_vars_merged(
    dataset = .,
    dataset_add = ADSL,
    new_vars = adsl_vars,
    by_vars = exprs(STUDYID, USUBJID)
  ) %>%
  mutate(
    AVISIT = case_when(
      ADT <= ENRLDT & !is.na(ABLFL) ~ "BASELINE",
      TRUE ~ VISIT
    ),
    AVISITN = case_when(
      ADT <= ENRLDT & !is.na(ABLFL) ~ 0,
      TRUE ~ ADY
    )
  ) %>%
  derive_var_base(
    dataset = .,
    by_vars = exprs(STUDYID, USUBJID, PARAMCD),
    source_var = AVALC
  ) %>%
  select(
    STUDYID, USUBJID, RSSEQ, COLPROT, PARAMCD, PARAM, AVISIT, AVISITN,
    AVALC, AVAL, ABLFL, BASE, ADT, ADY, VISITNUM, VISIT, EPOCH, ENRLFL, DTHFL
  )
```


```{r adrs-flags}
# Flag subjects that have baseline diagnostics summary and
## populate the flag per subject's records
ADRS_PREP <- ADRS_PREP %>%
  call_derivation(
    dataset = .,
    dataset_add = .,
    derivation = derive_vars_merged,
    variable_params = list(
      params(
        filter_add = ABLFL == "Y",
        new_vars = exprs(BLBFL = ABLFL)
      )
    ),
    by_vars = exprs(STUDYID, USUBJID),
    check_type = "error",
    relationship = "many-to-one"
  )

# Derive death parameter ----
DAETH_PARAM <- ADSL %>%
  select(!!!adsl_vars, DTHDT) %>%
  # Since month imputations was performed,
  # floor the date to the last date of a year
  mutate(
    DTHDT = ceiling_date(DTHDT, unit = "year") - days(1),
    DHDY = as.numeric(DTHDT - ENRLDT),
    DHDY = ifelse(DHDY == 0, 1, DHDY)
  ) %>%
  derive_var_merged_ef_msrc(
    dataset = .,
    by_vars = exprs(STUDYID, USUBJID),
    flag_events = list(
      flag_event(dataset_name = "ADRS_PREP", condition = BLBFL == "Y")
    ),
    source_datasets = list(ADRS_PREP = ADRS_PREP),
    new_var = BLBFL,
    true_value = "Y",
  )

# Add DEATH parameter ----
ADRS_PREP <- ADRS_PREP %>%
  # Add DEATH parameter
  derive_extreme_records(
    dataset = .,
    dataset_add = DAETH_PARAM,
    dataset_ref = DAETH_PARAM,
    by_vars = exprs(STUDYID, USUBJID),
    filter_add = !is.na(DTHDT),
    check_type = "error",
    exist_flag = AVALC,
    true_value = "DEATH",
    false_value = NA_character_,
    set_values_to = exprs(
      PARAMCD = "DEATH",
      PARAM = "Death",
      AVAL = ifelse(!is.na(DTHDT), 4, NA_real_),
      ADT = DTHDT,
      ADY = DHDY,
      AVISITN = DHDY
    )
  ) %>%
  select(-DTHDT, -DHDY) %>%
  # Remove missing DEATH param
  filter(!(PARAMCD %in% "DEATH" & is.na(DTHFL))) %>%
  mutate(PARAMN = case_when(
    PARAMCD == "DX" ~ 1,
    PARAMCD == "DEATH" ~ 2
  ))

# Flag subjects that have at least one follow-up diagnostics summary after baseline visit
#  Or had death records
ADRS_PREP <- ADRS_PREP %>%
  derive_var_merged_exist_flag(
    dataset = .,
    dataset_add = .,
    by_vars = exprs(STUDYID, USUBJID),
    filter_add = !ABLFL %in% "Y" & !is.na(AVALC) & (!EPOCH %in% "Screening" | AVISITN < 0),
    condition = BLBFL %in% "Y" | is.na(BLBFL),
    new_var = FOLLOWPFL,
    false_value = "N",
    missing_value = "M"
  )

# Identify records that will be used for the analysis: -----
# Enrolled subjects (ENRLFL: "Y"),
# Have a baseline diagnostics summary (BLBFL: "Y"), and
# have at least one followup diagnostics summary/death records (FOLLOWPFL: "Y")
ADRS_PREP <- ADRS_PREP %>%
  mutate(
    ANL01FL = case_when(
      ENRLFL == "Y" & BLBFL == "Y" & FOLLOWPFL == "Y" & !EPOCH %in% "Screening" ~ "Y",
      ENRLFL == "Y" & BLBFL == "Y" & FOLLOWPFL == "Y" & EPOCH %in% "Screening" & ABLFL %in% "Y" ~ "Y"
    )
  ) %>%
  select(-ENRLFL, -ENRLDT, -DTHFL)
```

```{r generate-adrs}
ADRS <- ADRS_PREP %>%
  # Merge ADSL ADSL dataset
  derive_vars_merged(
    dataset = .,
    dataset_add = single_build_from_derived(
      metacore = adrs_metacore,
      dataset_name = "ADRS",
      ds_list = list("ADSL" = ADSL),
      predecessor_only = TRUE,
      keep = "ALL"
    ),
    by_vars = exprs(STUDYID, USUBJID),
    check_type = "error",
    relationship = "many-to-one"
  ) %>%
  drop_unspec_vars(
    dataset = .,
    metacore = adrs_metacore
  ) %>%
  metatools::set_variable_labels(
    data = .,
    metacore = adrs_metacore
  )
```

```{r check-adrs}
ADRS <- ADRS %>%
  check_variables(
    data = .,
    metacore = adrs_metacore
  ) %>%
  check_ct_data(
    data = .,
    metacore = adrs_metacore,
    na_acceptable = TRUE,
    omit_vars = c("ENRLFL", "ABLFL", "PARAMCD")
  ) %>%
  order_cols(
    data = .,
    metacore = adrs_metacore
  ) %>%
  sort_by_key(
    data = .,
    metacore = adrs_metacore
  ) %>%
  xportr_metadata(
    .df = .,
    metadata = adrs_metacore,
    verbose = "stop"
  ) %>%
  # xportr_type() %>%
  xportr_length() %>%
  xportr_label()
```
