---
title: "ADNIMERGE2-Derived-Data"
output: 
 rmarkdown::html_vignette:
  toc: true
  code_folding: hide
params: 
   INCLUDE_PACC: true
date: "Last Updated: `r format(Sys.Date(), '%B %d, %Y')`"
vignette: >
  %\VignetteIndexEntry{ADNIMERGE2-Derived-Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  class.source = "fold-show"
)
# Data print function
datatable <- function(data, paging = FALSE, searchable = TRUE, bInfo = FALSE, ...) {
  DT::datatable(
    data = data, ...,
    options = list(paging = paging, searchable = searchable, bInfo = bInfo, ...)
  )
}
```

# Introduction {#sec-intro}

This article describes creating derived datasets from raw data (eCRFs) for ADNI study. The purpose of generating such standardized dataset is to create analysis ready dataset using the [PHARMERVERSE](https://pharmaverse.org/) workflow, which required standardized dataset as input. In `ADNIMERGE2` R data package, the following derived datasets will be created for illustration purpose, and more information about of each dataset is presented in the corresponding subsections.

* Demographic (DM)
* Subject Characteristics (SC)
* Adverse Event (AE)
* Questionnaires (QS)
* Clinical Classification (RS)
* Nervous System Finding (NV)
* Laboratory Test Result (LB)
* Genomics Findings (GF)
* Vital Sign (VS)

**NOTE: ** 

* *ORIGPROT* variable is included in the DM dataset to identify the first study protocol/phase of subjects' enrollment in the ADNI study. 

* *COLPROT* or *---GRPID* variables are also included in some of the derived dataset to identify the data collection study protocol/phase in the ADNI study.

* *VISITNUM* variable is included only for creating study epoch across the study phases and mainly used for data merging and sorting purpose. The *VISITNUM* variable should not be used in analysis that related to these derived dataset. Learn more about how the study visit number *VISITNUM* is created in `Data Preparation` section below.

* *These derived dataset may not be fully complied with the [CDISC-SDTM](https://www.cdisc.org/standards/foundational/sdtm) standardization.*  

# Load Required R Packages {#sec-load-pkg}

```{r setup-libraries}
library(tidyverse)
library(assertr)
library(labelled)
library(DT)
library(measurements)
library(sdtm.oak)
```

```{r study-library, purl = FALSE}
# ADNI study R data package
library(ADNIMERGE2)
```

```{r params-check, include = FALSE, echo = FALSE}
check_object_type(params$INCLUDE_PACC, "logical")
```

# Data Preparation {#sec-data-prep}

In the study, subjects are categorized into two groups:
  
  + *New Subject: *Subjects that did not participate/enroll in any of the previous ADNI study phases prior to a given study phase.
  
  + *Rollover Subject: *Subjects that participated/enrolled at least in one of the previous study phases prior to a given study phase.

## Study Visit and Visit Number {#sec-visitnum .unlisted}

Some data wrangling were performed prior to creating the pre-specified derived datasets. First, study epoch (*EPOCH*) with corresponding study visit number (*VISITNUM*) was created based on the [`VISITS`](../reference/VISITS.html) data, and stored as `EPOCH_LIST_LONG`. Next, subject-specific study visit with corresponding study epoch was created based on the [`REGISTRY`](../reference/REGISTRY.html) and [`ROSTER`](../reference/ROSTER.html) eCRFs along with the `EPOCH_LIST_LONG` data. Then, the result output named as `ADNI_VISIT_RECORD`. Here, rollover subjects may not have sequential *VISITNUM* due to the study design which allows rollovers to enroll in the next study phase before completing a given study phase. As a result, *VISITNUM* records can be used for data merging and sorting purpose with corresponding study visit date/form completion date.

For more information, please refer to the source script of package vignettes `vignette(topic = 'ADNIMERGE2-Derived-Data', package = 'ADNIMERGE2')`.

## Assessement Completion Date and Status {#sec-status .unlisted}

A study visit date from the [`REGISTRY`](../reference/REGISTRY.html) eCRF will be used if the assessment specific completion date is missing with known assessment result. Furthermore, any assessments with missing/unknown results will be considered as not completed/not done.

## Baseline Flag {#sec-bflg .unlisted}

A baseline flag is included in some of the derived dataset to identify records that are closest to subject enrollment date or collected at baseline visit. A record is flagged as baseline if one of the following criteria is met: 

* Record collected at baseline visits regardless the baseline visit/collection date.

* Any record collected prior to the baseline visit and closest to the enrollment window period end date if the baseline record value is missing or not collected. Enrollment window period is defined as 90 days after the actual enrollment date.

The assessment baseline flag is created using the [`derive_blfl_adni`](https://github.com/atri-biostats/ADNIMERGE2/tree/inst/derived-dataset-sdtmoak-utils.R) function which is based the `sdtm.oak::derive_blfl` function with minor study-specific modification. 
*NOTE: * However, there might be more than one baseline flags per assessment per subject in a given derived dataset. For future work, `derive_blfl_adni` function will be updated to account such cases.

## Subject Identifier {#sec-subjid .unlisted}

In ADNI derived dataset, subject identifier variable (`USUBJID`) was created based on subject identification number in the raw dataset with corresponding site id. The subject identifier variable will have `ADNI-XXX-XXXXX` format. The first three digits represents site id (`SITEID`) variable that obtained either from the [`REGISTRY`](../reference/REGISTRY.html) or [`PTDEMOG`](../reference/PTDEMOG.html) or [`ROSTER`](../reference/ROSTER.html) eCRFs, and the last five digits represents the subject identification number in the raw dataset (`RID`). A preceding `0` value was added for any values that have less number of digits. 

*NOTE: * Please see `create_usubjid()` and [`derive_usubjid()`](https://github.com/atri-biostats/ADNIMERGE2/tree/main/inst/derived-dataset-utils.R) functions to learn more about how the subject identifier variable was created.

```{r system-utils, purl = FALSE, echo = TRUE, eval = TRUE}
# Load utils function from package system file
utils_file_list <- c(
  "derived-dataset-sdtmoak-utils.R",
  "derived-dataset-utils.R",
  "derived-labdata-utils.R"
)
utils_file_path <- system.file(
  utils_file_list,
  package = "ADNIMERGE2",
  mustWork = TRUE
)
load_utils_funs <- lapply(utils_file_path, source)
```

```{r phase-visit-epoch, class.source = NULL, echo = FALSE}
# Create study epoch stage- overall and phase-specific
EPOCH_LIST <- VISITS %>%
  filter(!VISCODE %in% c("reg", "AUT", "4_disp")) %>%
  mutate(EPOCH = case_when(
    VISCODE %in% c("sc", "4_sc", "scmri", "f", "v01", "v02") ~ "Screening",
    VISCODE %in% c("bl", "4_bl", "init", "4_init", "v03", "v06") ~ "Baseline",
    VISCODE %in% c("reg", "f") ~ NA_character_,
    TRUE ~ "Follow-up" # Includes uns1, tau and nv visits
  )) %>%
  mutate(PHASE_EPOCH = paste0(PHASE, " - ", EPOCH)) %>%
  select(PHASE, VISCODE, PHASE_EPOCH, EPOCH, VISORDER, VISNAME) %>%
  group_by(PHASE) %>%
  arrange(VISORDER) %>%
  mutate(PHASE_VISITNUM = row_number() - 2) %>%
  ungroup() %>%
  # Create phase-specific visit number/order
  # 999 - represents unscheduled visit
  mutate(PHASE_VISITNUM = case_when(
    VISCODE %in% c("nv", "uns1", "tau") ~ 999,
    PHASE %in% "ADNI1" & VISCODE %in% c("f", "sc") ~ -1,
    PHASE %in% "ADNI1" & VISCODE %in% "bl" ~ 1,
    PHASE %in% "ADNI1" & str_detect(VISCODE, "m") ~ PHASE_VISITNUM + 1,
    PHASE %in% "ADNIGO" & VISCODE %in% c("sc", "scmri") ~ PHASE_VISITNUM - 1,
    PHASE %in% "ADNI2" & VISCODE %in% c("v01", "v02") ~ PHASE_VISITNUM - 1,
    PHASE %in% "ADNI2" & VISCODE %in% c("v03", "v06") ~ 1,
    PHASE %in% "ADNI2" & VISCODE %in% c("v04", "v07") ~ 2,
    PHASE %in% "ADNI2" & VISCODE %in% "v05" ~ 3,
    PHASE %in% "ADNI2" & !VISCODE %in% str_c("v0", 1:7) ~ PHASE_VISITNUM - 2,
    PHASE %in% c("ADNI3", "ADNI4") & VISCODE %in% c("sc", "4_sc") ~ -1,
    PHASE %in% c("ADNI3", "ADNI4") & VISCODE %in% c("bl", "4_bl") ~ 1,
    TRUE ~ PHASE_VISITNUM
  )) %>%
  mutate(PTTYPE = case_when(
    PHASE %in% c("ADNI1", "ADNIGO") ~ "Both",
    str_detect(VISNAME, "Continuing ") == TRUE ~ "Rollover",
    str_detect(VISNAME, "New Pt|Screening") == TRUE ~ "New",
    TRUE ~ "Both"
  )) %>%
  arrange(PHASE, PHASE_VISITNUM) %>%
  mutate(VISIT = case_when(
    !str_detect(VISNAME, paste0(adni_phase(), collapse = "|")) ~ paste0(PHASE, " ", VISNAME),
    TRUE ~ VISNAME
  )) %>%
  assert_non_missing(-VISORDER)

# Separate by participant type
EPOCH_LIST_LONG <- EPOCH_LIST %>%
  filter(PTTYPE %in% c("New", "Both")) %>%
  mutate(PTTYPE = "New") %>%
  bind_rows(
    EPOCH_LIST %>%
      filter(PTTYPE %in% c("Rollover", "Both")) %>%
      mutate(PTTYPE = "Rollover") %>%
      # Adjust EPOCH for rollovers
      mutate(EPOCH = case_when(
        EPOCH %in% c("Baseline", "Screening") &
          PHASE %in% c(adni_phase()[-1]) ~ "Follow-up",
        TRUE ~ EPOCH
      ))
  )
```

```{r include-origprot-colprot, class.source = NULL, echo = TRUE}
# Add study track (i.e., New or Rollover)
REGISTRY <- REGISTRY %>%
  mutate(PTTYPE = adni_study_track(COLPROT, ORIGPROT))

ROSTER <- ROSTER %>%
  mutate(PTTYPE = adni_study_track(COLPROT, ORIGPROT))
```

```{r study-visit-flows, class.source = NULL, echo = FALSE}
# Unique participant IDs
unique_rid <- unique(c(ROSTER$RID, REGISTRY$RID, PTDEMOG$RID))

# Generate VISITNUM and VISITDY
PHASE_VISIT_RECORD <- tibble(RID = unique_rid) %>%
  # Add ORIGPOL
  mutate(ORIGPROT = original_study_protocol(RID)) %>%
  # Last known study phases based ROSTER/REGISTRY data
  left_join(
    ROSTER %>%
      # Add study phase order number
      mutate(PHASE_NUM = adni_phase_order_num(COLPROT)) %>%
      assert_non_missing(PHASE_NUM) %>%
      distinct(RID, COLPROT, PHASE_NUM) %>%
      group_by(RID) %>%
      filter(PHASE_NUM == max(PHASE_NUM)) %>%
      ungroup() %>%
      assert_uniq(RID) %>%
      select(RID, LASTPROT = COLPROT, LAST_PHASE_NUM = PHASE_NUM),
    by = "RID"
  ) %>%
  assert_uniq(RID) %>%
  assert_non_missing(LASTPROT) %>%
  mutate(FIRST_PHASE_NUM = adni_phase_order_num(ORIGPROT)) %>%
  # Get possible study phases between the `FIRST` and `LAST` known study phase
  expand_grid(PHASE_NUM_INTERVAL = adni_phase_order_num(adni_phase())) %>%
  assert_non_missing(PHASE_NUM_INTERVAL) %>%
  # Possible study participation period
  filter(PHASE_NUM_INTERVAL <= LAST_PHASE_NUM & PHASE_NUM_INTERVAL >= FIRST_PHASE_NUM) %>%
  group_by(RID) %>%
  mutate(MAX_PHASE = n()) %>%
  ungroup() %>%
  filter(MAX_PHASE <= LAST_PHASE_NUM) %>%
  # Convert PHASE_NUM_INTERVAL to ADNI phase name
  mutate(COLPROT = convert_adni_phase_order_num(PHASE_NUM_INTERVAL)) %>%
  # Add study track
  mutate(PTTYPE = adni_study_track(COLPROT, ORIGPROT)) %>%
  # filter(RID %in% 1412) %>%
  # Add phase-specific visit code and order
  left_join(
    EPOCH_LIST_LONG %>%
      filter(PTTYPE %in% c("New", "Rollover")) %>%
      select(COLPROT = PHASE, VISCODE, PHASE_VISITNUM, VISIT, PTTYPE, EPOCH),
    relationship = "many-to-many",
    by = c("COLPROT", "PTTYPE")
  ) %>%
  assert_uniq(RID, COLPROT, VISCODE)

PHASE_VISIT_RECORD <- PHASE_VISIT_RECORD %>%
  left_join(
    REGISTRY %>%
      select(RID, ORIGPROT, COLPROT, VISCODE, EXAMDATE) %>%
      mutate(REGISTRY_RECORD = "Yes"),
    by = c("RID", "COLPROT", "ORIGPROT", "VISCODE")
  ) %>%
  verify(nrow(.) == nrow(PHASE_VISIT_RECORD))

# Overall VISITNUM (Planned Visits)
ADNI_VISIT_RECORD <- PHASE_VISIT_RECORD %>%
  filter(!c(PHASE_VISITNUM == 999 | VISCODE == "f")) %>%
  mutate(COLPROT = factor(COLPROT, levels = adni_phase())) %>%
  arrange(RID, COLPROT) %>%
  # Adjusting for rollover screening visits in ADNIGO
  mutate(PHASE_VISITNUM = case_when(
    PTTYPE %in% "Rollover" & COLPROT %in% "ADNIGO" ~ PHASE_VISITNUM + 2,
    TRUE ~ PHASE_VISITNUM
  )) %>%
  group_by(RID, COLPROT) %>%
  mutate(NUM_RECORDS_PHASE = n()) %>%
  nest(data = everything() & !c(RID, COLPROT, NUM_RECORDS_PHASE)) %>%
  ungroup() %>%
  group_by(RID) %>%
  mutate(CUM_NUM_RECORDS_PHASE = cumsum(NUM_RECORDS_PHASE)) %>%
  mutate(LAG_NUM_RECORDS_PHASE = lag(CUM_NUM_RECORDS_PHASE, n = 1L)) %>%
  ungroup() %>%
  # Adjusting screening for rollovers
  mutate(data = map2(
    .x = data,
    .y = LAG_NUM_RECORDS_PHASE,
    ~ mutate(.x,
      VISITNUM = case_when(
        is.na(.y) ~ PHASE_VISITNUM,
        !is.na(.y) ~ PHASE_VISITNUM + .y
      )
    )
  )) %>%
  unnest(cols = data)

# This mapping may overestimate the planned
#  VISITNUM for rollovers due to the study design
ADNI_VISIT_RECORD <- bind_rows(
  ADNI_VISIT_RECORD,
  PHASE_VISIT_RECORD %>%
    filter((PHASE_VISITNUM == 999 | VISCODE == "f")) %>%
    mutate(VISITNUM = PHASE_VISITNUM)
) %>%
  select(
    RID, ORIGPROT, COLPROT, LASTPROT, PTTYPE,
    VISCODE, VISIT, VISITNUM, EPOCH
  )

ADNI_VISIT_RECORD <- ADNI_VISIT_RECORD %>%
  # Add VISITDATE and VISTATUS
  left_join(
    REGISTRY %>%
      mutate(
        VISTAT = case_when(
          RGCONDCT %in% "No" | VISTYPE %in% "Not done" ~ "NOT DONE"
        ),
        VISDTC = create_iso8601(as.character(EXAMDATE), .format = "y-m-d")
      ) %>%
      select(RID, ORIGPROT, COLPROT, VISCODE, VISDTC, VISTAT) %>%
      distinct(),
    by = c("RID", "ORIGPROT", "COLPROT", "VISCODE")
  ) %>%
  verify(nrow(.) == nrow(ADNI_VISIT_RECORD))
```

# Building Derived Dataset {#sec-build-data}

## Demographic (DM) {#sec-dm}

[`DM`](../reference/DM.html) dataset contains one records per subject when subjects were enrolled or screened in the ADNI study for the first time (i.e. as new-enrollee).

```{r dm-data-dic, echo = FALSE}
dm_data_dic <- bind_rows(
  c(
    FLDNAME = "STUDYID",
    LABEL = "Study Identifier"
  ),
  c(
    FLDNAME = "DOMAIN",
    LABEL = "Domain Abbreviation"
  ),
  c(
    FLDNAME = "USUBJID",
    LABEL = "Unique Subject Identifier"
  ),
  c(
    FLDNAME = "SUBJID",
    LABEL = "Subject Identifier for the Study"
  ),
  c(
    FLDNAME = "ORIGPROT",
    LABEL = "Original study protocol",
    TEXT = "First time participation in ADNI study phase"
  ),
  c(
    FLDNAME = "SESTDTC",
    LABEL = "Screening Start Date/Time",
    TEXT = "First screening start date/time"
  ),
  c(
    FLDNAME = "RFSTDTC",
    LABEL = "Subject Enrolled Date/Time *"
  ),
  # c(FLDNAME = "RFENDTC", LABEL = "Subject Reference End Date/Time"),
  # c(FLDNAME = "RFICDTC", LABEL = "Date/Time of Informed Consent"),
  c(
    FLDNAME = "RFPENDTC",
    LABEL = "Date/Time of End of Participation*",
    TEXT = paste0(
      "Last known date/time when participant ended thier participation or",
      " follow-up in ADNI study. As character format for unknown MM-DD values."
    )
  ),
  c(
    FLDNAME = "DTHDTC",
    LABEL = "Date/Time of Death"
  ),
  c(
    FLDNAME = "DTHFL",
    LABEL = "Subject Death Flag"
  ),
  c(
    FLDNAME = "SITEID",
    LABEL = "Study Site Identifier"
  ),
  c(
    FLDNAME = "AGE",
    LABEL = "Age",
    TEXT = "Age at first screening visit (in years)"
  ),
  c(
    FLDNAME = "AGEU",
    LABEL = "Age Units"
  ),
  c(
    FLDNAME = "SEX",
    LABEL = "Sex at Birth"
  ),
  c(
    FLDNAME = "RACE",
    LABEL = "Race"
  ),
  c(
    FLDNAME = "ETHNIC",
    LABEL = "Ethnicity"
  ),
  # These fields are included to illustrate the PHARMAVERSE approach
  c(
    FLDNAME = "ARMCD",
    LABEL = "Planned Arm Code"
  ),
  c(
    FLDNAME = "ARM",
    LABEL = "Description of Planned Arm"
  ),
  c(
    FLDNAME = "ACTARMCD",
    LABEL = "Actual Arm Code"
  ),
  c(
    FLDNAME = "ACTARM",
    LABEL = "Actual Treatment Description"
  ),
  c(
    FLDNAME = "ARMNRS",
    LABEL = "Reason Arm and/or Actual Arm is Null",
    TEXT = "Non-Interventional Study"
  ),
  # Missing variables
  c(
    FLDNAME = "COUNTRY",
    LABEL = "Country",
    TEXT = "Country of the investigational site located"
  ),
  c(
    FLDNAME = "DMDTC",
    LABEL = "Date/Time of Collection"
  ),
  c(
    FLDNAME = "DMDY",
    LABEL = "Study Day of Collection"
  )
) %>%
  tibble::as_tibble() %>%
  relocate(c(FLDNAME, LABEL, TEXT)) %>%
  mutate(
    TEXT = replace_na(TEXT, " "),
    DERIVED = TRUE,
    TBLNAME = "DM",
    CRFNAME = "[ Derived ] Demographic Characteristics"
  )
```

<details>
<summary>Show `DM` Variables</summary>

```{r dm-data-dic-print, echo = FALSE}
dm_data_dic %>%
  select(FLDNAME, LABEL, TEXT) %>%
  datatable(., paging = TRUE)
```

</details>

```{r generate-dm0}
# Join columns
dm_join_var <- c("RID", "ORIGPROT")

# Demographic data columns
dm_cols <- c("PTGENDER", "PTRACCAT", "PTETHCAT", "PTDOB")
names(dm_cols) <- c("SEX", "RACE", "ETHNIC", "BRTHDTC")

DM <- tibble(RID = unique_rid) %>%
  mutate(ORIGPROT = original_study_protocol(RID = RID)) %>%
  # Add PTDEMOG record
  left_join(
    PTDEMOG %>%
      assert_non_missing(ORIGPROT, COLPROT) %>%
      filter(ORIGPROT == COLPROT) %>%
      group_by(RID) %>%
      filter(
        (any(!is.na(VISDATE)) & VISDATE == min(VISDATE, na.rm = TRUE)) |
          (all(is.na(VISDATE)) & row_number() == 1)
      ) %>%
      ungroup() %>%
      assert_uniq(RID) %>%
      select(RID, ORIGPROT, VISDATE, all_of(as.character(dm_cols))),
    by = dm_join_var
  ) %>%
  rename(all_of(dm_cols)) %>%
  generate_oak_id_vars_adni(raw_src = "PTDEMOG")
```

```{r generate-dm1}
DM <- DM %>%
  # Add screening visit date
  left_join(
    get_adni_screen_date(
      .registry = REGISTRY,
      phase = "Overall",
      multiple_screen_visit = FALSE
    ) %>%
      select(RID, ORIGPROT, SESTDTC = SCREENDATE) %>%
      assert(is_uniq, RID),
    by = dm_join_var
  ) %>%
  # Add enrollment/RFSTDTC date
  create_rfstdtc(.registry = REGISTRY) %>%
  # Compute age based on first screening date
  mutate(
    AGE = round(as.numeric(SESTDTC - my(BRTHDTC)) / 365.25, 1),
    AGEU = "Years",
    SUBJID = as.character(RID),
    ARMCD = NA_character_,
    ACTARM = NA_character_,
    ARM = NA_character_,
    ACTARMCD = NA_character_,
    ACTARM = NA_character_,
    ARMNRS = "Non-Interventional Study",
    COUNTRY = NA_character_,
    DMDTC = create_iso8601(as.character(VISDATE), .format = "y-m-d")
  )
```

```{r generate-dm2}
# Add death flag
DM <- DM %>%
  left_join(
    get_death_flag(
      .studysum = STUDYSUM,
      .adverse = ADVERSE,
      .recadv = RECADV
    ) %>%
      verify(all(DTHFL == "Yes")) %>%
      select(RID, ORIGPROT, DTHFL, DTHDTC),
    by = dm_join_var
  )
```

```{r generate-dm3}
# Add last known disposition date
DM <- DM %>%
  left_join(
    get_disposition_flag(
      .registry = REGISTRY,
      .studysum = STUDYSUM
    ) %>%
      mutate(COLPROT = factor(COLPROT, levels = adni_phase())) %>%
      group_by(RID) %>%
      arrange(COLPROT) %>%
      # Last known discontinuation/disposition date
      filter(row_number() == n()) %>%
      ungroup() %>%
      select(RID, ORIGPROT, SDSTATUS, RFPENDTC = SDDATE),
    by = dm_join_var
  ) %>%
  mutate(RFPENDTC = case_when(
    !is.na(DTHDTC) & is.na(RFPENDTC) ~ as.character(DTHDTC),
    TRUE ~ as.character(RFPENDTC)
  ))
```

```{r generate-dm4}
DM <- DM %>%
  # Derive USUBJID and SITIED
  derive_usubjid(
    .data = .,
    .registry = REGISTRY,
    .roster = ROSTER,
    .ptdemog = PTDEMOG,
    varList = c("USUBJID", "SITEID")
  ) %>%
  derive_study_day_adni(
    sdtm_in = .,
    domain = "DM",
    dm_domain = .,
    refdt = "RFSTDTC"
  ) %>%
  assign_studyid_domain(
    studyid = "ADNI",
    domain = "DM"
  ) %>%
  assign_vars_label(
    .data = .,
    data_dict = dm_data_dic,
    .strict = TRUE
  ) %>%
  assert_non_missing(SITEID, USUBJID, SUBJID)
```

## Subject Characteristics (SC) {#sec-sc}

[`SC`](../reference/SC.html) dataset contains subject-related data that are not collected in [`DM`](../reference/DM.html) domain and/or characteristics that are collected over time. In ADNI study, subject characteristics that are collected across the study phases: ADNI1, ADNIGO, ADNI2, ADNI3, and ADNI4 can be stored in [`SC`](../reference/SC.html). The [`SC`](../reference/SC.html) dataset will contains characteristics collected from three source datasets: [`PTDEMOG`](../reference/PTDEMOG.html), [`ADI`](../reference/ADI.html) 
 and [`RURALITY`](../reference/RURALITY.html).

```{r sc-data-dic, echo = FALSE}
sc_data_dic <- bind_rows(
  c(
    FLDNAME = "STUDYID",
    LABEL = "Study Identifier",
    TEXT = ""
  ),
  c(
    FLDNAME = "DOMAIN",
    LABEL = "Domain Abbrevation"
  ),
  c(
    FLDNAME = "USUBJID",
    LABEL = "Unique Subject Identifier"
  ),
  c(
    FLDNAME = "SCSEQ",
    LABEL = "Sequence Number"
  ),
  c(
    FLDNAME = "SCGRPID",
    LABEL = "Group ID",
    TEXT = "Study Protocol of Data Collection"
  ),
  c(
    FLDNAME = "SCTESTCD",
    LABEL = "Subject Characteristic Short Name"
  ),
  c(
    FLDNAME = "SCTEST",
    LABEL = "Subject Characteristic"
  ),
  c(
    FLDNAME = "SCCAT",
    LABEL = "Category for Subject Characteristic"
  ),
  c(
    FLDNAME = "SCORRES",
    LABEL = "Result or Finding in Original Units"
  ),
  c(
    FLDNAME = "SCSTRESN",
    LABEL = "Numeric Result/Finding in Standard Units"
  ),
  c(
    FLDNAME = "SCBLFL",
    LABEL = "Baseline Flag"
  ),
  c(
    FLDNAME = "VISITNUM",
    LABEL = "Visit Number"
  ),
  c(
    FLDNAME = "VISIT",
    LABEL = "Visit Name"
  ),
  # c(
  #   FLDNAME = "VISITDY",
  #   LABEL = "(Planned*) Study Day of Visit"
  # ),
  c(
    FLDNAME = "EPOCH",
    LABEL = "Epoch"
  ),
  c(
    FLDNAME = "SCDTC",
    LABEL = "Date/Time of Collection"
  ),
  c(
    FLDNAME = "SCDY",
    LABEL = "Study Day of Collection"
  )
) %>%
  tibble::as_tibble() %>%
  relocate(c(FLDNAME, LABEL, TEXT)) %>%
  mutate(TEXT = replace_na(TEXT, " ")) %>%
  mutate(
    DERIVED = TRUE,
    TBLNAME = "SC",
    CRFNAME = "[ Derived ] Subject Characteristics"
  )
```

<details>
<summary>Show `SC` Variables</summary>

```{r sc-data-dic-print, eval = TRUE, echo = FALSE}
sc_data_dic %>%
  select(FLDNAME, LABEL) %>%
  datatable(., paging = TRUE)
```

</details>

```{r generate-sc0}
# Full demographic records ----
sc_common_cols <- c("ORIGPROT", "COLPROT", "RID", "VISCODE", "VISDATE")

SC_PTDEMOG <- PTDEMOG %>%
  select(-all_of(
    c(
      "PTID", "VISCODE2", "ID", "SITEID", "USERDATE", "USERDATE2",
      "DD_CRF_VERSION_LABEL", "LANGUAGE_CODE", "HAS_QC_ERROR", "update_stamp"
    )
  )) %>%
  mutate(COLPROT = factor(COLPROT, levels = adni_phase())) %>%
  assert_non_missing(COLPROT) %>%
  group_by(RID, ORIGPROT, COLPROT) %>%
  # Since it is observational study
  arrange(COLPROT, VISDATE) %>%
  fill(-all_of(sc_common_cols), .direction = "down") %>%
  fill(-all_of(sc_common_cols), .direction = "up") %>%
  ungroup() %>%
  mutate(across(-all_of(sc_common_cols), as.character)) %>%
  mutate(VISCODE = factor(VISCODE, levels = c(unique(EPOCH_LIST$VISCODE)))) %>%
  assert_non_missing(VISCODE)

SC_DEMOG <- tibble(RID = unique_rid) %>%
  mutate(ORIGPROT = original_study_protocol(RID = RID)) %>%
  left_join(
    SC_PTDEMOG,
    by = dm_join_var
  ) %>%
  pivot_longer(
    cols = !all_of(sc_common_cols),
    names_to = "SCTESTCD",
    values_to = "SCORRES"
  ) %>%
  # Remove missing values
  drop_na(SCORRES) %>%
  mutate(
    SCCAT = "Demographic Records",
    SCDTC = as.character(VISDATE)
  ) %>%
  # Required a long format dataset
  generate_oak_id_vars_adni(raw_src = "PTDEMOG")

# Area Deprivation Index (ADI) for ADNI4 phase only ----
SC_ADI <- ADI %>%
  verify(all(COLPROT == adni_phase()[5])) %>%
  select(
    RID, ORIGPROT, COLPROT, VISCODE, ADISTATE, ADINATIONAL, ADIREV,
    SCDTC = ADIDATE
  ) %>%
  generate_oak_id_vars_adni(raw_src = "ADI") %>%
  mutate(across(c(ADISTATE, ADINATIONAL, ADIREV, SCDTC), as.character)) %>%
  pivot_longer(
    cols = c(ADISTATE, ADINATIONAL, ADIREV),
    names_to = "SCTESTCD",
    values_to = "SCORRES"
  ) %>%
  mutate(SCCAT = "Area Deprivation Index")

# RUCA and RUCC from ADNI4 phase only ----
rurality_cols <- c("RUCA", "RUCC", "RUCA_2010", "RUCC_2023")
SC_RURALITY <- RURALITY %>%
  verify(all(COLPROT == adni_phase()[5])) %>%
  select(
    RID, ORIGPROT, COLPROT, VISCODE, all_of(rurality_cols),
    SCDTC = RURDATE
  ) %>%
  generate_oak_id_vars_adni(raw_src = "RURALITY") %>%
  mutate(across(all_of(c(rurality_cols, "SCDTC")), as.character)) %>%
  pivot_longer(
    cols = all_of(rurality_cols),
    names_to = "SCTESTCD",
    values_to = "SCORRES"
  ) %>%
  mutate(SCCAT = "Rurality")
```

```{r generate-sc1}
SC <- bind_rows(SC_DEMOG, SC_ADI, SC_RURALITY) %>%
  assert_uniq(RID, ORIGPROT, COLPROT, VISCODE, SCTESTCD) %>%
  mutate(
    COLPROT = factor(COLPROT, levels = adni_phase()),
    SCGRPID = COLPROT,
    SCSTAT = case_when(is.na(SCDTC) ~ "NOT DONE"),
    SCSTRESN = as.numeric(SCORRES),
    SCDTC = create_iso8601(SCDTC, .format = "y-m-d")
  ) %>%
  assert_uniq(RID, ORIGPROT, COLPROT, VISCODE, SCTESTCD) %>%
  derive_usubjid(varList = "USUBJID") %>%
  assign_studyid_domain(
    .data = .,
    studyid = "ADNI",
    domain = "SC"
  ) %>%
  assign_visit_attr(
    .data = .,
    visit_record_data = ADNI_VISIT_RECORD,
    domain = "SC"
  ) %>%
  assign_epoch(
    .data = .,
    .epoch = EPOCH_LIST_LONG
  ) %>%
  derive_blfl_adni(
    sdtm_in = .,
    dm_domain = DM,
    tgt_var = "SCBLFL"
  ) %>%
  derive_study_day_adni(
    sdtm_in = .,
    dm_domain = DM,
    domain = "SC",
    refdt = "RFSTDTC"
  ) %>%
  derive_seq(
    tgt_var = "SCSEQ",
    rec_vars = c("USUBJID", "SCTESTCD", "SCGRPID")
  )
```

```{r scrtest-detail, eval = TRUE, echo = FALSE}
# Field labels from data dictionary
# Some labels might have more than 40 characters
sc_string_pattern <- paste0(
  "^[0-9]\\.|^[0-9]{2}\\.|^[0-9][a-z]\\.|",
  "[0-9]{2}[a-z]\\.|Participant|Participant\\'s",
  collapse = "|"
)

SC_TESTCD_LIST <- DATADIC %>%
  # filter(TBLNAME %in% c("PTDEMOG", "RURALITY", "ADI")) %>%
  filter(FLDNAME %in% SC$SCTESTCD) %>%
  distinct(FLDNAME, TEXT) %>%
  mutate(
    TEXT = str_remove_all(TEXT, sc_string_pattern),
    TEXT = trimws(TEXT)
  ) %>%
  select(SCTESTCD = FLDNAME, SCTEST = TEXT) %>%
  distinct() %>%
  group_by(SCTESTCD) %>%
  filter(row_number() == n()) %>%
  ungroup()
```

```{r add-sctest}
SC <- SC %>%
  # Add characteristics test name and check coded value
  set_dom_test(
    .data = .,
    .data_list = SC_TESTCD_LIST %>%
      select(SCTESTCD, SCTEST),
    merge_by = "SCTESTCD"
  ) %>%
  assert_non_missing(SCTESTCD, SCTEST) %>%
  assign_vars_label(data_dict = sc_data_dic)
```

## Adverse Events (AE) {#sec-ae}

[`AE`](../reference/AE.html) dataset contains one records per adverse event per subject. The [`AE`](../reference/AE.html) dataset will includes the following characteristics for subjects that have at least one adverse events experience during the study.

*NOTE: * Currently, `AE` only includes records that are collected in *ADNI3* and *ADNI4* study phases.

```{r ae-data-dic, echo = FALSE}
ae_data_dic <- bind_rows(
  c(
    FLDNAME = "STUDYID",
    LABEL = "Study Identifier"
  ),
  c(
    FLDNAME = "DOMAIN",
    LABEL = "Domain Abbreviation"
  ),
  c(
    FLDNAME = "USUBJID",
    LABEL = "Unique Subject Identifier"
  ),
  c(
    FLDNAME = "AEGRPID",
    LABEL = "Group ID",
    TEXT = "Study Protocol of Data Collection"
  ),
  c(
    FLDNAME = "AESEQ",
    LABEL = "Sequence Number"
  ),
  c(
    FLDNAME = "AETERM",
    LABEL = "Reported Term for the Adverse Event"
  ),
  c(
    FLDNAME = "AELLT",
    LABEL = "Lowest Level Term",
    TEXT = "Based on MeDRA Coding"
  ),
  c(
    FLDNAME = "AELLTCD",
    LABEL = "Lowest Level Term Code",
    TEXT = "Based on MeDRA Coding"
  ),
  c(
    FLDNAME = "AEDECOD",
    LABEL = "Dictionary-Derived Term",
    TEXT = "Based on MeDRA Coding"
  ),
  c(
    FLDNAME = "AEPTCD",
    LABEL = "Preferred Term Code",
    TEXT = "Based on MeDRA Coding"
  ),
  c(
    FLDNAME = "AEHLT",
    LABEL = "High Level Term",
    TEXT = "Based on MeDRA Coding"
  ),
  c(
    FLDNAME = "AEHLTCD",
    LABEL = "High Level Term Code",
    TEXT = "Based on MeDRA Coding"
  ),
  c(
    FLDNAME = "AEHLGT",
    LABEL = "High Level Group Term",
    TEXT = "Based on MeDRA Coding"
  ),
  c(
    FLDNAME = "AEHLGTCD",
    LABEL = "High Level Group Term Code",
    TEXT = "Based on MeDRA Coding"
  ),
  c(
    FLDNAME = "AESOC",
    LABEL = "Primary Sytem Organ Class",
    TEXT = "Based on MeDRA Coding"
  ),
  c(
    FLDNAME = "AESOCCD",
    LABEL = "Primary Sytem Organ Class Code",
    TEXT = "Based on MeDRA Coding"
  ),
  c(
    FLDNAME = "AESEV",
    LABEL = "Severity/Intensity"
  ),
  c(
    FLDNAME = "AESER",
    LABEL = "Serious Event"
  ),
  c(
    FLDNAME = "AEOUT",
    LABEL = "Outcome of Adverse Event"
  ),
  c(
    FLDNAME = "AESCONG",
    LABEL = "Congential Anomaly or Birth Defect"
  ),
  c(
    FLDNAME = "AESDISAB",
    LABEL = "Persist or Signif Disability/Incapacity"
  ),
  c(
    FLDNAME = "AESDTH",
    LABEL = "Results in Death"
  ),
  c(
    FLDNAME = "AESHOSP",
    LABEL = "Requires or Prologns Hospitalization"
  ),
  c(
    FLDNAME = "AESLIFE",
    LABEL = "Is Life Threating"
  ),
  c(
    FLDNAME = "AESMIE",
    LABEL = "Other Medically Important Serious Event"
  ),
  c(
    FLDNAME = "AECONTRT",
    LABEL = "Concomitant or Additional Trtmnt Given"
  ),
  c(
    FLDNAME = "AERELAD",
    LABEL = "Related to Early Stage or Progression of AD"
  ),
  c(
    FLDNAME = "AERELCM",
    LABEL = "Related to Concomitant Therapy"
  ),
  c(
    FLDNAME = "AERELFLRBTBN",
    LABEL = "Related to Florbetaben Tracer"
  ),
  c(
    FLDNAME = "AERELFLRBPR",
    LABEL = "Related to Florbetapir Tracer"
  ),
  c(
    FLDNAME = "AEHIMG",
    LABEL = "Related to PET/MRI Imaging Procedure"
  ),
  c(
    FLDNAME = "AERELTAU",
    LABEL = "Related to Flotaucipir Tracer "
  ),
  c(
    FLDNAME = "AERELNAV",
    LABEL = "Related to NAV-4694 Tracer"
  ),
  c(
    FLDNAME = "AERELMK",
    LABEL = "Related to MK-6240 Tracer"
  ),
  c(
    FLDNAME = "AERELPI",
    LABEL = "Related to PI-2620 Tracer"
  ),
  c(
    FLDNAME = "AEHLUMB",
    LABEL = "Related to Lumbar Puncture"
  ),
  c(
    FLDNAME = "AERELCOVID",
    LABEL = "Related to COVID-19 Illness"
  ),
  c(
    FLDNAME = "AERELPAN",
    LABEL = "Related to COVID-19 Pandemic Disruption"
  ),
  c(
    FLDNAME = "AERELATESP",
    LABEL = "Related to Other Study Procedure(s)"
  ),
  c(
    FLDNAME = "VISITNUM",
    LABEL = "Visit Number"
  ),
  c(
    FLDNAME = "VISIT",
    LABEL = "Visit Name"
  ),
  c(
    FLDNAME = "EPOCH",
    LABEL = "Epoch"
  ),
  c(
    FLDNAME = "AESTDTC",
    LABEL = "Start Date/Time of Adverse Event"
  ),
  c(
    FLDNAME = "AEENDTC",
    LABEL = "End Date/Time of Adverse Event"
  )
) %>%
  tibble::as_tibble() %>%
  relocate(c(FLDNAME, LABEL)) %>%
  mutate(
    TEXT = " ",
    DERIVED = TRUE,
    TBLNAME = "AE",
    CRFNAME = "[ Derived ] Adverse Events"
  )
```

<details>
<summary>Show `AE` Variables</summary>

```{r ae-data-dic-print, eval = TRUE, echo = FALSE}
ae_data_dic %>%
  select(FLDNAME, LABEL) %>%
  datatable(., paging = TRUE)
```

</details>

```{r generate-ae-adni34}
# Adverse Event for ADNI3 and ADNI4
AE_ADNI34 <- ADVERSE %>%
  select(
    ID, RID, ORIGPROT, COLPROT, VISCODE, SITEID, AENUMBER, AEOUTCOME,
    AEHONSDT, AEHCSDT, AEHDTHDT, AERELAD, AERELCM, AERELFLRBTBN, AERELFLRBPR,
    AEHIMG, AERELTAU, AERELNAV, AERELMK, AERELPI, AEHLUMB, AERELCOVID, AERELPAN,
    AERELATESP, AESERIOUS, AESERDATE, SAELIFE, SAEHOSPIT, SAEPROLONG, SAEDEATH,
    SAECONGEN, SAEDISAB, SAEOTHER, AEHCMEDS,
    AESEV0 = AEHSEVR, all_of(paste0("AESEV", 1:10))
  ) %>%
  generate_oak_id_vars_adni(raw_src = "ADVERSE") %>%
  assert_non_missing(AENUMBER) %>%
  # Filter the worst severity level per RID, AENUMBER, COLPROT
  mutate(across(all_of(paste0("AESEV", 0:10)), as.character)) %>%
  pivot_longer(
    cols = all_of(paste0("AESEV", 0:10)),
    names_to = "SEVERITY_COL",
    values_to = "AESEV"
  ) %>%
  mutate(AESEV_NUM = case_when(
    AESEV == "Mild" ~ 1,
    AESEV == "Moderate" ~ 2,
    AESEV == "Severe" ~ 3
  )) %>%
  group_by(RID, ORIGPROT, COLPROT, SITEID, AENUMBER, VISCODE) %>%
  filter(
    (all(is.na(AESEV)) & row_number() == 1) |
      (any(!is.na(AESEV)) & AESEV_NUM == max(AESEV_NUM, na.rm = TRUE))
  ) %>%
  filter(
    (n() > 1 & row_number() == n()) |
      (n() == 1 & row_number() == 1)
  ) %>%
  ungroup() %>%
  verify(nrow(.) == nrow(ADVERSE)) %>%
  assert_uniq(RID, ORIGPROT, COLPROT, AENUMBER, VISCODE) %>%
  rename(
    "AESER" = AESERIOUS, "AEOUT" = AEOUTCOME,
    "AESCONG" = SAECONGEN, "AESDISAB" = SAEDISAB, "AESDTH" = SAEDEATH,
    "AESLIFE" = SAELIFE, "AESMIE" = SAEOTHER, "AECONTRT" = AEHCMEDS,
    "AESTDTC" = AEHONSDT, "AEENDTC" = AEHCSDT
  ) %>%
  # Adverse events for `required or prolongs hospitalization`
  mutate(
    AESHOSP = case_when(
      SAEHOSPIT == "Yes" | SAEPROLONG == "Yes" ~ "Yes",
      SAEHOSPIT == "No" & SAEPROLONG == "No" ~ "No",
      SAEHOSPIT == "No" & is.na(SAEPROLONG) ~ "No",
      is.na(SAEHOSPIT) & SAEPROLONG == "No" ~ "No"
    )
  ) %>%
  select(-c(SAEHOSPIT, SAEPROLONG, AESEV_NUM, SEVERITY_COL))
```

```{r generate-ae-adni12go}
# Required checking for missing AENUMBER in RECADV
AE_ADNI12GO <- tibble(PHASE = NA_character_) %>%
  na.omit()
```

```{r generate-ae}
AE <- AE_ADNI34 %>%
  bind_rows(AE_ADNI12GO) %>%
  assert_non_missing(RID) %>%
  mutate(
    AESTDTC = as.character(AESTDTC),
    AEENDTC = as.character(AEENDTC)
  ) %>%
  group_by(RID) %>%
  arrange(AESTDTC) %>%
  mutate(AESEQ = row_number()) %>%
  ungroup()

AE <- AE %>%
  derive_usubjid() %>%
  assign_studyid_domain(domain = "AE") %>%
  assign_visit_attr() %>%
  assign_epoch() %>%
  mutate(
    AEGRPID = COLPROT,
    AETERM = NA_character_,
    AELLT = NA_character_,
    AELLTCD = NA_character_,
    AEDECOD = NA_character_,
    AEPTCD = NA_character_,
    AEHLT = NA_character_,
    AEHLTCD = NA_character_,
    AEHLGT = NA_character_,
    AEHLGTCD = NA_character_,
    AESOC = NA_character_,
    AESOCCD = NA_character_
  ) %>%
  assign_vars_label(data_dict = ae_data_dic)
```

## Questionnaires (QS) {#sec-qs}

[`QS`](../reference/QS.html) dataset contains one record per parameter finding (i.e. total score) per visit per subject.

```{r qs-data-dic, echo = FALSE}
qs_data_dic <- bind_rows(
  c(
    FLDNAME = "STUDYID",
    LABEL = "Study Identifier",
    TEXT = ""
  ),
  c(
    FLDNAME = "DOMAIN",
    LABEL = "Domain Abbrevation"
  ),
  c(
    FLDNAME = "USUBJID",
    LABEL = "Unique Subject Identifier"
  ),
  c(
    FLDNAME = "QSSEQ",
    LABEL = "Sequence Number"
  ),
  c(
    FLDNAME = "QSGRPID",
    LABEL = "Group ID",
    TEXT = "Study Protocol of Data Collection"
  ),
  c(
    FLDNAME = "QSTESTCD",
    LABEL = "Question Short Name"
  ),
  c(
    FLDNAME = "QSTEST",
    LABEL = "Question Name"
  ),
  c(
    FLDNAME = "QSCAT",
    LABEL = "Category of Question"
  ),
  c(
    FLDNAME = "QSORRES",
    LABEL = "Finding in Original Units"
  ),
  c(
    FLDNAME = "QSSTRESC",
    LABEL = "Character Result/Finding in Std Format"
  ),
  c(
    FLDNAME = "QSSTRESN",
    LABEL = "Numeric Finding in Standard Units"
  ),
  c(
    FLDNAME = "QSSTAT",
    LABEL = "Completion Status"
  ),
  c(
    FLDNAME = "QSREASND",
    LABEL = "Reason Not Performed"
  ),
  c(
    FLDNAME = "QSBLFL",
    LABEL = "Baseline Flag"
  ),
  c(
    FLDNAME = "QSDRVFL",
    LABEL = "Derived Flag"
  ),
  c(
    FLDNAME = "VISITNUM",
    LABEL = "Visit Number"
  ),
  c(
    FLDNAME = "VISIT",
    LABEL = "Visit Name"
  ),
  # c(
  #   FLDNAME = "VISITDY",
  #   LABEL = "(Planned*) Study Day of Visit"
  # ),
  c(
    FLDNAME = "EPOCH",
    LABEL = "Epoch"
  ),
  c(
    FLDNAME = "QSDTC",
    LABEL = "Date/Time of Finding"
  ),
  c(
    FLDNAME = "QSDY",
    LABEL = "Study Day of Finding"
  )
) %>%
  tibble::as_tibble() %>%
  relocate(c(FLDNAME, LABEL, TEXT)) %>%
  mutate(TEXT = replace_na(TEXT, " ")) %>%
  mutate(
    DERIVED = TRUE,
    TBLNAME = "QS",
    CRFNAME = "[ Derived ] Questionnaires"
  )
```

<details>
<summary>Show `QS` Variables</summary>

```{r qs-data-dic-print, eval = TRUE, echo = FALSE}
qs_data_dic %>%
  select(FLDNAME, LABEL) %>%
  datatable(., paging = TRUE)
```

</details>

The following cognitive/functional assessment scores are included in the [`QS`](../reference/QS.html) dataset.

```{r qs-testcd-list, eval = TRUE, echo = FALSE}
OTHER_QS_TESTCD_LIST <- bind_rows(
  c(
    QSCAT = "ADAS-COG",
    QSTESTCD = "ADASTT11",
    QSTEST = "ADAS-Cognitive Behavior 11-Item Total Score",
    SOURCE = "ADAS"
  ),
  c(
    QSCAT = "ADAS-COG",
    QSTESTCD = "ADASTT13",
    QSTEST = "ADAS-Cognitive Behavior 13-Item Total Score",
    SOURCE = "ADAS"
  ),
  c(
    QSCAT = "CDR",
    QSTESTCD = "CDGLOBAL",
    QSTEST = "Clinical Dementia Rating Global Score",
    SOURCE = "CDR"
  ),
  c(
    QSCAT = "CDR",
    QSTESTCD = "CDRSB",
    QSTEST = "Clinical Dementia Rating Sum of Boxes",
    SOURCE = "CDR"
  ),
  c(
    QSCAT = "ECOG SHORT FORM",
    QSTESTCD = "ECOGPTTT",
    QSTEST = "Everyday Cognition - Participant",
    SOURCE = "ECOGPT"
  ),
  c(
    QSCAT = "ECOG SHORT FORM",
    QSTESTCD = "ECOGSPTT",
    QSTEST = "Everyday Cognition - Study Partner",
    SOURCE = "ECOGSP"
  ),
  c(
    QSCAT = "FAQ",
    QSTESTCD = "FAQTOTAL",
    QSTEST = "Functional Assessments Questionnaires",
    SOURCE = "FAQ"
  ),
  c(
    QSCAT = "FCI",
    QSTESTCD = "FCISCORE",
    QSTEST = "Financial Capacity Instrument Short Form",
    SOURCE = "FCI"
  ),
  c(
    QSCAT = "GDS",
    QSTESTCD = "GDTOTAL",
    QSTEST = "Geriatric Depression Scale",
    SOURCE = "GDS"
  ),
  c(
    QSCAT = "MMSE",
    QSTESTCD = "MMSCORE",
    QSTEST = "Mini Mental State Exam",
    SOURCE = "MMSE"
  ),
  c(
    QSCAT = "MOCA",
    QSTESTCD = "MOCA",
    QSTEST = "Montreal Cognitive Assessments",
    SOURCE = "MOCA"
  ),
  c(
    QSCAT = "NPI",
    QSTESTCD = "NPITOTAL",
    QSTEST = "Neuropsychiatric Inventory",
    SOURCE = "NPI"
  ),
  c(
    QSCAT = "NPIQ",
    QSTESTCD = "NPIQTOTL",
    QSTEST = "Neuropsychiatric Inventory Q",
    SOURCE = "NPIQ"
  )
)

NEUROBAT_TESTCD <- bind_rows(
  c(
    QSTESTCD = "DIGITSCR",
    QSTEST = "Digit Symbol Substitution"
  ),
  c(
    QSTESTCD = "LDELTOTL",
    QSTEST = "Logical Memory - Delayed Recall"
  ),
  c(
    QSTESTCD = "LIMMTOTL",
    QSTEST = "Logical Memory - Immediate Recall"
  ),
  c(
    QSTESTCD = "RAVLTFG",
    QSTEST = "RAVLT Forgetting (Trial 5 - Delayed)"
  ),
  c(
    QSTESTCD = "RAVLTFGP",
    QSTEST = "RAVLT Percent Forgetting"
  ),
  c(
    QSTESTCD = "RAVLTIMM",
    QSTEST = "RVAL Immediate (Sum of 5 Trials)"
  ),
  c(
    QSTESTCD = "RAVLTLRN",
    QSTEST = "RVAL Learning (Trial 5 - Trial 1)"
  ),
  c(
    QSTESTCD = "TRABSCOR",
    QSTEST = "Trails B"
  )
) %>%
  mutate(
    QSCAT = "NEUROBAT",
    SOURCE = "NEUROBAT"
  )

if (params$INCLUDE_PACC) {
  PACC_TESTCD <- bind_rows(
    c(
      QSTESTCD = "mPACCdigit",
      QSTEST = "Modified PACC Digit Score"
    ),
    c(
      QSTESTCD = "mPACCtrailsB",
      QSTEST = "Modified PACC Trials B Score"
    ),
  ) %>%
    mutate(
      QSTESTCD = str_to_upper(QSTESTCD),
      QSCAT = "PACC",
      SOURCE = "PACC"
    )
}

QS_TESTCD_LIST <- bind_rows(
  OTHER_QS_TESTCD_LIST,
  NEUROBAT_TESTCD
) %>%
  {
    if (params$INCLUDE_PACC) bind_rows(., PACC_TESTCD) else (.)
  }
```


```{r print-qs-testcd-list, eval = TRUE, echo = FALSE}
QS_TESTCD_LIST %>%
  arrange(SOURCE, QSCAT, QSTESTCD) %>%
  select(SOURCE, QSCAT, QSTESTCD, QSTEST) %>%
  datatable(., paging = TRUE)
```

```{r qs-input}
qs_com_cols <- c(
  "RID", "ORIGPROT", "COLPROT", "VISCODE",
  "VISCODE2", "VISDATE", "SITEID"
)

# ADAS Cognitive Behavior Total Score ----
ADAS_SCORE_DATA <- ADAS %>%
  select(all_of(qs_com_cols), ADASTT11 = TOTSCORE, ADASTT13 = TOTAL13) %>%
  generate_oak_id_vars_adni(raw_src = "ADAS") %>%
  pivot_longer(
    cols = c(ADASTT11, ADASTT13),
    names_to = "QSTESTCD",
    values_to = "QSSTRESC"
  ) %>%
  mutate(
    QSSTRESC = as.character(QSSTRESC),
    QSDRVFL = "Yes",
    QSSTAT = ifelse(is.na(QSSTRESC), "NOT DONE", NA_character_)
  )

# Clinical Dementia Rating Score ----
CDR_SCORE_DATA <- CDR %>%
  select(all_of(qs_com_cols), CDGLOBAL, CDRSB) %>%
  generate_oak_id_vars_adni(raw_src = "CDR") %>%
  pivot_longer(
    cols = c(CDGLOBAL, CDRSB),
    names_to = "QSTESTCD",
    values_to = "QSSTRESC"
  ) %>%
  mutate(
    QSSTRESC = as.character(QSSTRESC),
    QSSTAT = ifelse(is.na(QSSTRESC), "NOT DONE", NA_character_)
  )

# Everyday Cognition Total Score ----
ECOG_SCORE_DATA <- ECOGPT %>%
  mutate(QSTESTCD = "ECOGPTTT") %>%
  select(all_of(qs_com_cols), QSTESTCD, QSSTRESC = EcogPtTotal) %>%
  generate_oak_id_vars_adni(raw_src = "ECOGPT") %>%
  bind_rows(
    ECOGSP %>%
      mutate(QSTESTCD = "ECOGSPTT") %>%
      select(all_of(qs_com_cols), QSTESTCD, QSSTRESC = EcogSPTotal) %>%
      generate_oak_id_vars_adni(raw_src = "ECOGSP")
  ) %>%
  mutate(
    QSSTRESC = as.character(QSSTRESC),
    QSDRVFL = "Yes",
    QSSTAT = ifelse(is.na(QSSTRESC), "NOT DONE", NA_character_)
  )

# Financial Capacity Instrument Short Form - Score ----
FCI_SCORE_DATA <- FCI %>%
  select(all_of(qs_com_cols),
    QSSTRESC = FCISCORE,
    QSSTAT = DONE, QSREASND = NDREASON
  ) %>%
  mutate(
    QSSTRESC = as.character(QSSTRESC),
    QSTESTCD = "FCISCORE"
  ) %>%
  generate_oak_id_vars_adni(raw_src = "FCI")

# Functional Assessments Questionnaires - Score ----
FAQ_SCORE_DATA <- FAQ %>%
  select(all_of(qs_com_cols), QSSTRESC = FAQTOTAL) %>%
  mutate(
    QSSTRESC = as.character(QSSTRESC),
    QSTESTCD = "FAQTOTAL",
  ) %>%
  generate_oak_id_vars_adni(raw_src = "FAQ")

# Geriatric Depression Scale ----
GDS_SCORE_DATA <- GDSCALE %>%
  select(all_of(qs_com_cols), QSSTRESC = GDTOTAL, QSREASND = GDUNABL) %>%
  mutate(
    QSSTRESC = as.character(QSSTRESC),
    QSTESTCD = "GDTOTAL",
    QSSTAT = ifelse(!is.na(QSREASND), "NOT DONE", NA_character_)
  ) %>%
  generate_oak_id_vars_adni(raw_src = "GDSCALE")

# Mini Mental State Exam Score ----
MMSE_SCORE_DATA <- MMSE %>%
  select(all_of(qs_com_cols),
    QSSTRESC = MMSCORE,
    QSSTAT = DONE, QSREASND = NDREASON
  ) %>%
  mutate(
    QSSTRESC = as.character(QSSTRESC),
    QSTESTCD = "MMSCORE",
    QSDRVFL = "Yes"
  ) %>%
  generate_oak_id_vars_adni(raw_src = "MMSE")

# Montreal Cognitive Assessments ----
MOCA_SCORE_DATA <- MOCA %>%
  select(all_of(qs_com_cols), QSSTRESC = MOCA) %>%
  mutate(
    QSSTRESC = as.character(QSSTRESC),
    QSTESTCD = "MOCA",
    QSDRVFL = "Yes"
  ) %>%
  generate_oak_id_vars_adni(raw_src = "MOCA")

# Neuropsychiatric Inventory ----
NPI_SCORE_DATA <- NPI %>%
  rename("VISDATE" = EXAMDATE) %>%
  select(all_of(qs_com_cols), QSSTRESC = NPITOTAL) %>%
  mutate(
    QSSTRESC = as.character(QSSTRESC),
    QSTESTCD = "NPITOTAL",
    QSDRVFL = "Yes",
    QSSTAT = ifelse(is.na(QSSTRESC), "NOT DONE", NA_character_)
  ) %>%
  generate_oak_id_vars_adni(raw_src = "NPI")

# Neuropsychiatric Inventory Q ----
NPIQ_SCORE_DATA <- NPIQ %>%
  select(all_of(qs_com_cols), QSSTRESC = NPISCORE) %>%
  mutate(
    QSSTRESC = as.character(QSSTRESC),
    QSTESTCD = "NPIQTOTL",
    QSDRVFL = "Yes",
    QSSTAT = ifelse(is.na(QSSTRESC), "NOT DONE", NA_character_)
  ) %>%
  generate_oak_id_vars_adni(raw_src = "NPIQ")

# Logical Memory - Immediate/Delayed Recall ----
neurobat_cols <- c(
  "LIMMTOTL", "LDELTOTL", "DIGITSCR", "TRABSCOR",
  "RAVLTIMM", "RAVLTLRN", "RAVLTFG", "RAVLTFGP"
)
NEUROBAT_SCORE_DATA <- NEUROBAT %>%
  compute_neurobat_subscore(.neurobat = .) %>%
  select(all_of(c(qs_com_cols, neurobat_cols))) %>%
  generate_oak_id_vars_adni(raw_src = "NEUROBAT") %>%
  pivot_longer(
    cols = all_of(neurobat_cols),
    names_to = "QSTESTCD",
    values_to = "QSSTRESC"
  ) %>%
  mutate(
    QSSTRESC = as.character(QSSTRESC),
    QSDRVFL = case_when(
      QSTESTCD %in% c(
        "RAVLTIMM", "RAVLTRN", "RAVLTFG", "RAVLTFGP"
      ) ~ "Yes"
    ),
    QSSTAT = ifelse(is.na(QSSTRESC), "NOT DONE", NA_character_)
  )

# Composite modified PACC score
if (params$INCLUDE_PACC) {
  pacc_cols <- c("mPACCdigit", "mPACCtrailsB")

  PACC_SCORE_DATA <- PACC %>%
    rename("VISDATE" = PACC_VISDATE) %>%
    select(any_of(qs_com_cols), all_of(pacc_cols)) %>%
    mutate(RID = as.numeric(RID)) %>%
    generate_oak_id_vars_adni(raw_src = "PACC") %>%
    pivot_longer(
      cols = all_of(pacc_cols),
      names_to = "QSTESTCD",
      values_to = "QSSTRESC"
    ) %>%
    mutate(
      QSTESTCD = str_to_upper(QSTESTCD),
      QSSTRESC = as.character(QSSTRESC),
      QSDRVFL = "Yes",
      QSSTAT = ifelse(is.na(QSSTRESC), "NOT DONE", NA_character_)
    )
}

# Score data names
score_data_names <- ls()[str_detect(ls(), "SCORE\\_DATA")]
```

```{r generate-qs}
QS <- mget(score_data_names) %>%
  bind_rows() %>%
  assert_non_missing(RID, COLPROT, VISCODE, QSTESTCD, VISCODE) %>%
  mutate(
    QSGRPID = COLPROT,
    QSDTC = create_iso8601(as.character(VISDATE), .format = "y-m-d"),
    QSSTRESN = as.numeric(QSSTRESC),
    QSORRES = as.character(QSSTRESN),
    QSSTAT = case_when(
      QSSTAT %in% c("No", "NOT DONE") ~ "NOT DONE",
      TRUE ~ NA_character_
    )
  ) %>%
  set_dom_test(
    .data_list = QS_TESTCD_LIST %>%
      select(QSTESTCD, QSTEST, QSCAT),
    merge_by = "QSTESTCD"
  ) %>%
  assert_uniq(RID, COLPROT, VISCODE, QSTESTCD) %>%
  derive_usubjid() %>%
  assign_studyid_domain(domain = "QS") %>%
  assign_visit_attr() %>%
  assign_epoch() %>%
  derive_blfl_adni(
    dm_domain = DM,
    tgt_var = "QSBLFL"
  ) %>%
  derive_study_day_adni(
    dm_domain = DM,
    domain = "QS"
  ) %>%
  derive_seq(
    tgt_var = "QSSEQ",
    rec_vars = c("USUBJID", "QSTESTCD", "COLPROT")
  ) %>%
  # # Update QSORRES for derived scores
  # mutate(QSORRES = case_when(is.na(QSDRVFL) ~ QSSTRESC)) %>%
  assign_vars_label(data_dict = qs_data_dic)
```

## Clinical Classification (RS) {#sec-rs}

[`RS`](../reference/RS.html) dataset will contains one record per instrument status per visit per subject.

```{r rs-data-dic, echo = FALSE}
rs_data_dic <- bind_rows(
  c(
    FLDNAME = "STUDYID",
    LABEL = "Study Identifier",
    TEXT = ""
  ),
  c(
    FLDNAME = "DOMAIN",
    LABEL = "Domain Abbrevation"
  ),
  c(
    FLDNAME = "USUBJID",
    LABEL = "Unique Subject Identifier"
  ),
  c(
    FLDNAME = "RSSEQ",
    LABEL = "Sequence Number"
  ),
  c(
    FLDNAME = "RSGRPID",
    LABEL = "Group ID",
    TEXT = "Study Protocol of Data Collection"
  ),
  c(
    FLDNAME = "RSTESTCD",
    LABEL = "Assessment Short Name"
  ),
  c(
    FLDNAME = "RSTEST",
    LABEL = "Assessment Name"
  ),
  c(
    FLDNAME = "RSCAT",
    LABEL = "Category for Assessment"
  ),
  c(
    FLDNAME = "RSORRES",
    LABEL = "Result or Finding in Original Units"
  ),
  c(
    FLDNAME = "RSSTRESC",
    LABEL = "Character Result/Finding in Std Format"
  ),
  c(
    FLDNAME = "RSBLFL",
    LABEL = "Baseline Flag"
  ),
  c(
    FLDNAME = "RSEVAL",
    LABEL = "Evaluator",
    TEXT = "SITEID will be used instead of actual evaluator initial."
  ),
  c(
    FLDNAME = "VISITNUM",
    LABEL = "Visit Number"
  ),
  c(
    FLDNAME = "VISIT",
    LABEL = "Visit Name"
  ),
  # c(
  #   FLDNAME = "VISITDY",
  #   LABEL = "(Planned*) Study Day of Visit"
  # ),
  c(
    FLDNAME = "EPOCH",
    LABEL = "Epoch"
  ),
  c(
    FLDNAME = "RSDTC",
    LABEL = "Date/Time of Assessment"
  ),
  c(
    FLDNAME = "RSDY",
    LABEL = "Study Day of Assessment"
  )
) %>%
  tibble::as_tibble() %>%
  relocate(FLDNAME, LABEL, TEXT) %>%
  mutate(TEXT = replace_na(TEXT, " ")) %>%
  mutate(
    DERIVED = TRUE,
    TBLNAME = "RS",
    CRFNAME = "[ Derived ] Clinical Classification"
  )
```

<details>
<summary>Show `RS` Variables</summary>

```{r rs-data-dic-print, eval = TRUE, echo = FALSE}
rs_data_dic %>%
  select(FLDNAME, LABEL, TEXT) %>%
  datatable(., paging = TRUE)
```

</details>

The [`RS`](../reference/RS.html) dataset will contains clinical diagnostics summary of subject per study visit. The clinical diagnostics status either Cognitive Normal (CN), Mild Cognitive Impairment (MCI) or Dementia/Alzheimer's (DEM/AD) of a subject was determined by clinicians' judgment.

```{r rs-testcd-list, eval = TRUE, echo = FALSE}
# Recommended to check variable name length and test labels
RS_TESTCD_LIST <- bind_rows(
  c(
    RSCAT = "DIAGNOSTICS",
    RSTESTCD = "DX",
    RSTEST = "Clinical Diagnostics Status",
    SOURCE = "DXSUM"
  )
)

RS_TESTCD_LIST %>%
  arrange(SOURCE, RSCAT, RSTESTCD) %>%
  select(SOURCE, RSCAT, RSTESTCD, RSTEST) %>%
  datatable()
```

```{r generate-rs0}
# Clinical diagnostics status
RS <- DXSUM %>%
  generate_oak_id_vars_adni(raw_src = "DXSUM") %>%
  mutate(
    DIAGNOSIS = case_when(
      DIAGNOSIS %in% "Dementia" ~ "DEM",
      TRUE ~ as.character(DIAGNOSIS)
    ),
    RSTESTCD = "DX",
    RSORRES = as.character(DIAGNOSIS),
    RSSTRESC = as.character(DIAGNOSIS),
    RSEVAL = SITEID,
    RSDTC = create_iso8601(as.character(EXAMDATE), .format = "y-m-d"),
    RSGRPID = COLPROT,
    RSSTAT = NA_character_
  ) %>%
  assert_non_missing(COLPROT)
```

```{r generate-rs1}
RS <- RS %>%
  set_dom_test(
    .data_list = RS_TESTCD_LIST %>%
      select(RSCAT, RSTEST, RSTESTCD),
    merge_by = "RSTESTCD"
  ) %>%
  derive_usubjid() %>%
  assign_studyid_domain(domain = "RS") %>%
  assign_visit_attr() %>%
  assign_epoch() %>%
  derive_blfl_adni(
    dm_domain = DM,
    tgt_var = "RSBLFL"
  ) %>%
  derive_study_day_adni(
    dm_domain = DM,
    domain = "RS"
  ) %>%
  derive_seq(
    tgt_var = "RSSEQ",
    rec_vars = c("USUBJID", "RSTESTCD", "COLPROT")
  ) %>%
  assign_vars_label(data_dict = rs_data_dic)
```

## Nervous System Finding (NV) {#sec-nv}

[`NV`](../reference/NV.html) dataset contains one record of physiological and morphological finding related to the nervous system (including brain) per visit per subject.

**NOTE: **

* Records in the raw dataset of [`Amyloid PET`](../reference/UCBERKELEY_AMY_6MM.html) and [`Tau PET`](../reference/UCBERKELEY_TAU_6MM.html) may include a QC flag. Records that contain `Fail`, `Not assessed` or `Cannot be processed` QC flag will be indicated as `NOT DONE` status in the `NV` derived dataset with those flags as a reason. We suggest to check this variable value before using `ADNIMERGE2::NV` data in any analysis. For examples, records that were collected at baseline but contained a QC flag of `Fail`, `Not assessed` or `Cannot be processed` will not be flagged as a baseline record in `NV`. To learn more about about the QC flag, please refer to the the manual documentation at data sharing platform. Moreover, please see the following r chunks about how these QC flags were translated in the `NV` derived data.

```{r nv-data-dic, echo = FALSE}
nv_data_dic <- bind_rows(
  c(
    FLDNAME = "STUDYID",
    LABEL = "Study Identifier",
    TEXT = ""
  ),
  c(
    FLDNAME = "DOMAIN",
    LABEL = "Domain Abbrevation"
  ),
  c(
    FLDNAME = "USUBJID",
    LABEL = "Unique Subject Identifier"
  ),
  c(
    FLDNAME = "NVSEQ",
    LABEL = "Sequence Number"
  ),
  c(
    FLDNAME = "NVGRPID",
    LABEL = "Group ID",
    TEXT = "Study Protocol of Data Collection"
  ),
  c(
    FLDNAME = "NVLNKID",
    LABEL = "Link ID"
  ),
  c(
    FLDNAME = "NVTESTCD",
    LABEL = "Short Name of Nervous System Test"
  ),
  c(
    FLDNAME = "NVTEST",
    LABEL = "Name of Nervous System Test"
  ),
  c(
    FLDNAME = "NVCAT",
    LABEL = "Category for Nervous System"
  ),
  c(
    FLDNAME = "NVSCAT",
    LABEL = "Subcategory for Nervous System"
  ),
  c(
    FLDNAME = "NVORRES",
    LABEL = "Result or Finding in Original Units"
  ),
  c(
    FLDNAME = "NVSTRESC",
    LABEL = "Character Result/Finding in Std Format"
  ),
  c(
    FLDNAME = "NVSTRESN",
    LABEL = "Numeric Result/Finding in Standard Units"
  ),
  # c(
  #   FLDNAME = "NVSTRESU",
  #   LABEL = "Standard Units"
  # ),
  c(
    FLDNAME = "NVSTAT",
    LABEL = "Completion Status"
  ),
  c(
    FLDNAME = "NVREASND",
    LABEL = "Reason Not Done"
  ),
  c(
    FLDNAME = "NVMETHOD",
    LABEL = "Method of Test or Examination"
  ),
  c(
    FLDNAME = "NVBLFL",
    LABEL = "Baseline Flag"
  ),
  c(
    FLDNAME = "NVDRVFL",
    LABEL = "Derived Flag"
  ),
  c(
    FLDNAME = "VISITNUM",
    LABEL = "Visit Number"
  ),
  c(
    FLDNAME = "VISIT",
    LABEL = "Visit Name"
  ),
  # c(
  #   FLDNAME = "VISITDY",
  #   LABEL = "(Planned*) Study Day of Visit"
  # ),
  c(
    FLDNAME = "EPOCH",
    LABEL = "Epoch"
  ),
  c(
    FLDNAME = "NVDTC",
    LABEL = "Date/Time of Collection"
  ),
  c(
    FLDNAME = "NVDY",
    LABEL = "Study Day of Collection"
  )
) %>%
  tibble::as_tibble() %>%
  relocate(c(FLDNAME, LABEL, TEXT)) %>%
  mutate(TEXT = replace_na(TEXT, " ")) %>%
  mutate(
    DERIVED = TRUE,
    TBLNAME = "NV",
    CRFNAME = "[ Derived ] Nervous System Finding"
  )
```

<details>
<summary>Show `NV` Variables</summary>

```{r nv-data-dic-print, eval = TRUE, echo = FALSE}
nv_data_dic %>%
  select(FLDNAME, LABEL, TEXT) %>%
  datatable(., paging = TRUE)
```

</details>

```{r nv-testcd-list, eval = TRUE, echo = FALSE}
# Recommended to check variable name length and test labels
AMYPET_TESTCD <- bind_rows(
  c(
    NVCAT = "AMYLOIDPET",
    NVTESTCD = "AMYSTAT",
    NVTEST = "Amyloid Status - Centiloids",
    TEXT = paste0(
      "Amyloid status determined based on centiloids that normalized by",
      "whole cerebellum. ",
      "UC Berkeley - Amyloid PET 6mm Res analysis."
    )
  ),
  c(
    NVCAT = "AMYLOIDPET",
    NVTESTCD = "AMYSTATC",
    NVTEST = "Amyloid Status - Centiloids Composite Ref",
    TEXT = paste0(
      "Amyloid status determined based on centiloids that normalized by ",
      "composite reference region. ",
      "UC Berkeley - Amyloid PET 6mm Res analysis."
    )
  ),
  c(
    NVCAT = "AMYLOIDPET",
    NVTESTCD = "CENTILOIDS",
    NVTEST = "Centiloids",
    TEXT = paste0(
      "Summary cortical SUVR normalized by whole cerebellum and transformed",
      " to Centiloids. ",
      "UC Berkeley - Amyloid PET 6mm Res analysis."
    )
  ),
  c(
    NVCAT = "AMYLOIDPET",
    NVTESTCD = "SUVRSM",
    NVTEST = "Summary Cortical SUVR",
    TEXT = paste0(
      "Summary cortical SUVR normalized by whole cerebellum. ",
      "UC Berkeley - Amyloid PET 6mm Res analysis."
    )
  ),
  c(
    NVCAT = "AMYLOIDPET",
    NVTESTCD = "SUVRCR",
    NVTEST = "Summary Cortical SUVR Composite Ref",
    TEXT = paste0(
      "Summary cortical SUVR normalized by composite reference region. ",
      "UC Berkeley - Amyloid PET 6mm Res analysis."
    )
  )
) %>%
  mutate(SOURCE = "UCBERKELEY_AMY_6MM") %>%
  expand_grid(NVSCAT = c("FBB-6MM", "FBB-8MM", "FBP-6MM", "FBP-8MM", "NAV-6MM"))

TAUPET_TESTCD <- bind_rows(
  c(
    NVCAT = "TAUPET",
    NVTESTCD = "SUVRINFE",
    NVTEST = "Standard Uptake Value - Inferior CGM",
    TEXT = paste0(
      "Reference region - mean SUVR of inferior cerebellar grey matter. ",
      "UC Berkeley - Tau PET 6mm Res analysis."
    )
  ),
  c(
    NVCAT = "TAUPET",
    NVTESTCD = "SUVRSC",
    NVTEST = "Standard Uptake Value - SCWM",
    TEXT = paste0(
      "Reference region - mean SUVR of subcortical white matter eroded",
      "away from cortex. ",
      "UC Berkeley - Tau PET 6mm Res analysis."
    )
  ),
  c(
    NVCAT = "TAUPET",
    NVTESTCD = "SUVRMETA",
    NVTEST = "Weighted Standard Uptake Value",
    TEXT = paste0(
      "Weighted mean SUVR of Jack et al temporal meta-ROI. ",
      "UC Berkeley - Tau PET 6mm Res analysis."
    )
  )
) %>%
  mutate(SOURCE = "UCBERKELEY_TAU_6MM") %>%
  expand_grid(NVSCAT = c("FTP-6MM", "FTP-8MM", "MK6240-6MM", "PI2620-6MM"))

TAUPETPVC_TESTCD <- bind_rows(
  c(
    NVCAT = "TAUPETPVC",
    NVTESTCD = "SUVRINFE",
    NVTEST = "Standard Uptake Value - Inferior CGM",
    TEXT = paste0(
      "Reference region - mean SUVR of inferior cerebellar grey matter. ",
      "UC Berkeley - Tau PET PVC 6mm Res analysis."
    )
  ),
  c(
    NVCAT = "TAUPETPVC",
    NVTESTCD = "SUVRCWM",
    NVTEST = "Standard Uptake Value - CWM",
    TEXT = paste0(
      "Bilateral cerebral white matter.",
      "UC Berkeley - Tau PET PVC 6mm Res analysis."
    )
  ),
  c(
    NVCAT = "TAUPETPVC",
    NVTESTCD = "SUVRMETA",
    NVTEST = "Weighted Standard Uptake Value",
    TEXT = paste0(
      "Weighted mean SUVR of Jack et al temporal meta-ROI. ",
      "UC Berkeley - Tau PET PVC 6mm Res analysis."
    )
  )
) %>%
  mutate(SOURCE = "UCBERKELEY_TAUPVC_6MM") %>%
  expand_grid(NVSCAT = c("FTP-NONE", "MK6240-NONE", "PI2620-NONE"))

NV_TESTCD_LIST <- bind_rows(
  c(
    NVCAT = "FDG-8MM",
    NVTESTCD = "FDGMROI",
    NVTEST = "FDG-PET metaROI 8mm",
    TEXT = "UC Berkeley - FDG analysis: 2023-02-17. Available for ADNI1-3.",
    SOURCE = "UCBERKELEYFDG_8mm"
  ),
  c(
    NVCAT = "PIBPETSV",
    NVTESTCD = "PIB",
    NVTEST = "Average PIB PET SUVr",
    TEXT = "U Pitt PIB PET Analysis. Only available for ADNI1.",
    SOURCE = "PIBPETSUVR"
  ),
  c(
    NVCAT = "AMYLOID-STATUS",
    NVTESTCD = "AMYSTAT",
    NVTEST = "Amyloid Status - Clinician Judgment",
    TEXT = "Avaiable for ADNI4 phase.",
    SOURCE = "AMYREAD"
  ),
  AMYPET_TESTCD,
  TAUPET_TESTCD,
  TAUPETPVC_TESTCD
) %>%
  mutate(NVSCAT = ifelse(is.na(NVSCAT), NVCAT, NVSCAT)) %>%
  separate(NVSCAT,
    into = c("TRACER", "IMAGE_RESOLUTION"),
    sep = "-",
    remove = FALSE
  ) %>%
  mutate(across(
    all_of(c("TRACER", "IMAGE_RESOLUTION")),
    ~ case_when(
      is.na(.x) | .x %in% c("PIBPETSV", "AMYLOID", "STATUS") ~ " - ",
      TRUE ~ as.character(.x)
    )
  ))
```

```{r nv-testcd-list-print, eval = TRUE, echo = FALSE}
NV_TESTCD_LIST %>%
  arrange(NVCAT, NVTESTCD) %>%
  select(
    SOURCE, NVTESTCD, NVTEST, NVCAT, NVSCAT,
    `TRACER TYPE` = TRACER,
    `IMAGE RESOLUTION` = IMAGE_RESOLUTION, DESCRIPTION = TEXT,
  ) %>%
  datatable(paging = TRUE, filter = "top") %>%
  formatStyle("DESCRIPTION", fontSize = "10px")
```

```{r nv-input1}
# FDG-PET Analysis Results Data ----
FDG_PET_DATA <- UCBERKELEYFDG_8mm %>%
  # Based on method manual document
  select(ORIGPROT, RID, VISCODE, VISCODE2, EXAMDATE, ROINAME, MEAN) %>%
  filter(!is.na(VISCODE)) %>%
  assert_uniq(RID, VISCODE, VISCODE2, ROINAME) %>%
  verify(all(ORIGPROT %in% adni_phase()[-5])) %>%
  pivot_wider(
    id_cols = everything(),
    names_from = "ROINAME",
    values_from = "MEAN"
  ) %>%
  mutate(FDGMROI = MetaROI / Top50PonsVermis) %>%
  select(-MetaROI, -Top50PonsVermis) %>%
  check_duplicate_records(col_names = c("RID", "EXAMDATE", "VISCODE")) %>%
  mutate(
    NVMETHOD = "FDG PET",
    NVTESTCD = "FDGMROI",
    NVORRES = as.character(FDGMROI),
    NVSTRESC = as.character(FDGMROI),
    NVSTRESN = FDGMROI,
    NVDRVFL = "Yes",
    NVDTC = as.character(EXAMDATE)
  ) %>%
  generate_oak_id_vars_adni(raw_src = "UCBERKELEYFDG_8mm")

# Mapping phase-specific visit code based opn registry
FDG_PET_DATA <- FDG_PET_DATA %>%
  # Trying to map visit code from registry
  use_dtplyr() %>%
  left_join(
    REGISTRY %>%
      mutate(EXAMDATE = as.character(EXAMDATE)) %>%
      select(RID, ORIGPROT, COLPROT, VISCODE, VISCODE2,
        REGISTRY.EXAMDATE = EXAMDATE
      ) %>%
      filter(!is.na(REGISTRY.EXAMDATE)) %>%
      filter(RID %in% unique(FDG_PET_DATA$RID)) %>%
      distinct() %>%
      check_duplicate_records(
        col_names = c("RID", "ORIGPROT", "VISCODE", "REGISTRY.EXAMDATE")
      ),
    by = c("RID", "ORIGPROT", "VISCODE", "VISCODE2")
  ) %>%
  as_tibble() %>%
  mutate(COLPROT = case_when(
    is.na(COLPROT) & VISCODE == VISCODE2 ~ ORIGPROT,
    is.na(COLPROT) & VISCODE != VISCODE2 & VISCODE2 %in% "bl" ~ ORIGPROT,
    TRUE ~ COLPROT
  )) %>%
  verify(nrow(.) == nrow(FDG_PET_DATA)) %>%
  assert_non_missing(COLPROT) %>%
  assert_uniq(RID, COLPROT, VISCODE, NVTESTCD)
```

```{r nv-input2}
# PIB PET Analysis Results Data -----
# Only for ADNI1 phase
PIB_PET_DATA <- PIBPETSUVR %>%
  verify(all(ORIGPROT == adni_phase()[1])) %>%
  # Based on previously generated dataset
  mutate(PIB = rowMeans(across(c("ACG", "FRC", "PAR", "PRC")), na.rm = FALSE)) %>%
  select(RID, ORIGPROT, VISCODE, EXAMDATE, NVSTRESN = PIB, LONIUID) %>%
  check_duplicate_records(col_names = c("RID", "EXAMDATE")) %>%
  mutate(
    NVMETHOD = "PET",
    NVTESTCD = "PIB",
    NVSTRESU = NA_character_,
    NVORRES = as.character(NVSTRESN),
    NVSTRESC = as.character(NVSTRESN),
    NVDRVFL = "Yes",
    NVDTC = as.character(EXAMDATE),
    NVLNKID = as.character(LONIUID),
    COLPROT = adni_phase()[1]
  ) %>%
  assert_non_missing(VISCODE) %>%
  assert_uniq(RID, VISCODE, NVTESTCD) %>%
  generate_oak_id_vars_adni(raw_src = "PIBPETSUVR")
```

```{r nv-input3}
# Amyloid status ----
amystatus_lvls <- c("Non Elevated", "Elevated")

AMYREAD_DATA <- AMYREAD %>%
  verify(all(COLPROT == adni_phase()[5])) %>%
  # Based on clinicians' decision
  mutate(AMYSTAT = case_when(
    str_detect(CONSENS, "No, visual read and quantification") ~ as.character(CONGRU),
    str_detect(CONSENS, "Yes, this scan should be reviewed") ~ as.character(CONSENSRES)
  )) %>%
  mutate(
    AMYSTAT = str_remove(AMYSTAT, " scan"),
    TRACERTYPE = as.character(TRACERTYPE)
  ) %>%
  verify(all(AMYSTAT %in% amystatus_lvls)) %>%
  generate_oak_id_vars_adni(raw_src = "AMYREAD") %>%
  pivot_longer(
    cols = AMYSTAT,
    names_to = "NVTESTCD",
    values_to = "NVORRES"
  ) %>%
  mutate(
    NVDTC = as.character(SCANDATE),
    NVSTRESC = as.character(NVORRES),
    NVMETHOD = "PET",
    NVDRVFL = "Yes"
  )
```

```{r nv-imaging-mapping-list, echo=FALSE}
# Imaging mapping List ----
imaging_cols <- c(
  "ORIGPROT", "COLPROT", "RID", "VISCODE", "SCANDATE", "LONIUID"
)
IMAGING_MAPPING_LIST <- bind_rows(
  AMYQC %>%
    select(all_of(imaging_cols)) %>%
    mutate(SOURCE = "AMYQC"),
  AV45QC %>%
    rename("SCANDATE" = EXAMDATE) %>%
    select(all_of(imaging_cols)) %>%
    mutate(SOURCE = "AV45QC"),
  PETQC %>%
    rename("SCANDATE" = EXAMDATE) %>%
    select(all_of(imaging_cols)) %>%
    mutate(SOURCE = "PETQC"),
  TAUQC %>%
    select(all_of(imaging_cols)) %>%
    mutate(SOURCE = "TAUQC"),
  PIBQC %>%
    mutate(
      COLPROT = adni_phase()[1],
      SCANDATE = EXAMDATE
    ) %>%
    select(all_of(imaging_cols)) %>%
    mutate(SOURCE = "PIBQC"),
) %>%
  filter(!is.na(LONIUID)) %>%
  use_dtplyr() %>%
  group_by(RID, SCANDATE) %>%
  filter((all(n() >= 1) & row_number() == n())) %>%
  ungroup() %>%
  as_tibble() %>%
  assert_uniq(all_of(c("RID", "SCANDATE")))
```

```{r nv-amypet}
# Common cols in PET data
pet_data_common_cols <- c(
  "ORIGPROT", "LONIUID", "RID", "VISCODE", "SCANDATE", "PROCESSDATE",
  "IMAGE_RESOLUTION", "TRACER", "qc_flag"
)

# Amyloid PET Data ----
amypet_cols <- list(
  AMYSTAT = "AMYLOID_STATUS",
  AMYSTATC = "AMYLOID_STATUS_COMPOSITE_REF",
  SUVRSM = "SUMMARY_SUVR",
  SUVRCR = "COMPOSITE_REF_SUVR",
  CENTILOIDS = "CENTILOIDS"
)

AMYPET_DATA <- UCBERKELEY_AMY_6MM %>%
  mutate(across(
    all_of(as.character(amypet_cols[1:2])),
    ~ case_when(
      .x == 1 ~ amystatus_lvls[2],
      .x == 0 ~ amystatus_lvls[1]
    )
  )) %>%
  rename_with_list(., name_char = amypet_cols, by_name = TRUE) %>%
  select(all_of(c(pet_data_common_cols, names(amypet_cols)))) %>%
  mutate(across(all_of(names(amypet_cols)), as.character)) %>%
  distinct() %>%
  generate_oak_id_vars_adni(raw_src = "UCBERKELEY_AMY_6MM") %>%
  pivot_longer(
    cols = all_of(names(amypet_cols)),
    names_to = "NVTESTCD",
    values_to = "NVSTRESC"
  )
```

```{r nv-taupet}
# Tau PET Data -----
taupet_cols <- list(
  SUVRINFE = "INFERIORCEREBELLUM_SUVR",
  SUVRSC = "ERODED_SUBCORTICALWM_SUVR",
  SUVRMETA = "META_TEMPORAL_SUVR"
)
TAUPET_DATA <- UCBERKELEY_TAU_6MM %>%
  # filter(!is.na(VISCODE)) %>%
  rename_with_list(., name_char = taupet_cols, by_name = TRUE) %>%
  select(all_of(c(pet_data_common_cols, names(taupet_cols)))) %>%
  mutate(across(all_of(names(taupet_cols)), as.character)) %>%
  generate_oak_id_vars_adni(raw_src = "UCBERKELEY_TAU_6MM") %>%
  pivot_longer(
    cols = all_of(names(taupet_cols)),
    names_to = "NVTESTCD",
    values_to = "NVSTRESC"
  )

# Tau PET - PVC Data ----
taupet_pvc_cols <- list(
  SUVRINFE = "INFERIORCEREBELLUM_SUVR",
  SUVRCWM = "CEREBRAL_WHITE_MATTER_SUVR",
  SUVRMETA = "META_TEMPORAL_SUVR"
)
TAUPET_PVC_DATA <- UCBERKELEY_TAUPVC_6MM %>%
  mutate(
    IMAGE_RESOLUTION = "None",
    qc_flag = NA_real_
  ) %>%
  rename_with_list(., name_char = taupet_pvc_cols, by_name = TRUE) %>%
  select(all_of(c(pet_data_common_cols, names(taupet_pvc_cols)))) %>%
  mutate(across(all_of(names(taupet_pvc_cols)), as.character)) %>%
  generate_oak_id_vars_adni(raw_src = "UCBERKELEY_TAUPVC_6MM") %>%
  pivot_longer(
    cols = all_of(names(taupet_pvc_cols)),
    names_to = "NVTESTCD",
    values_to = "NVSTRESC"
  )
```

```{r nv-petdata}
# PET dataset
PET_join_var <- paste0(c("ORIGPROT", "RID", "SCANDATE"), "_MPL")
names(PET_join_var) <- str_remove_all(PET_join_var, "\\_MPL")

PET_DATA <- bind_rows(AMYPET_DATA, TAUPET_DATA, TAUPET_PVC_DATA) %>%
  select(-VISCODE) %>%
  # Fuzzy join for actual study phase and visits
  left_fuzzy_join(
    data1 = .,
    data2 = IMAGING_MAPPING_LIST %>%
      select(ORIGPROT, COLPROT, RID, VISCODE, SCANDATE, SOURCE) %>%
      rename_with_list(., name_char = PET_join_var, by_name = FALSE),
    join_by = PET_join_var,
    check_cols = "COLPROT",
    main_cols = "SCANDATE",
    date_col = "SCANDATE"
  ) %>%
  mutate(
    LONIUID = as.character(LONIUID),
    NVMETHOD = "PET",
    NVDTC = as.character(SCANDATE),
    NVLNKID = as.character(LONIUID),
    NVORRES = as.character(NVSTRESC),
    NVSTRESN = as.numeric(NVSTRESC),
    NVSTAT = case_when(qc_flag %in% -2:0 ~ "NOT DONE"),
    NVREASND = case_when(
      qc_flag == -2 ~ "CANNOT BE PROCESSED",
      qc_flag == -1 ~ "NOT ASSESSED",
      qc_flag == 0 ~ "FAIL"
    )
  ) %>%
  assert_non_missing(COLPROT)
```

```{r generate-nv}
NV <- bind_rows(FDG_PET_DATA, PIB_PET_DATA, AMYREAD_DATA, PET_DATA) %>%
  mutate(
    NVGRPID = COLPROT,
    NVDTC = create_iso8601(NVDTC, .format = "y-m-d")
  ) %>%
  mutate(NVSCAT_MAPID = case_when(
    str_detect(raw_source, "^UCBERKELEY_") ~ toupper(paste0(TRACER, "-", IMAGE_RESOLUTION)),
    !str_detect(raw_source, "^UCBERKELEY_") ~ "-"
  ))

NV <- NV %>%
  left_join(
    NV_TESTCD_LIST %>%
      select(NVCAT, NVSCAT, NVTESTCD, NVTEST, SOURCE) %>%
      mutate(NVSCAT_MAPID = case_when(
        str_detect(SOURCE, "^UCBERKELEY_") ~ NVSCAT,
        !str_detect(SOURCE, "^UCBERKELEY_") ~ "-"
      )),
    by = c("NVTESTCD" = "NVTESTCD", "raw_source" = "SOURCE", "NVSCAT_MAPID" = "NVSCAT_MAPID"),
    relationship = "many-to-many"
  ) %>%
  verify(nrow(.) == nrow(NV)) %>%
  assert_non_missing(NVTEST, NVSCAT, NVCAT) %>%
  derive_usubjid() %>%
  assign_studyid_domain(domain = "NV") %>%
  assign_visit_attr() %>%
  assign_epoch()

# Adjusting for the same NVTESTCD that comes from multiple raw sources
nv_grp_cols <- c("NVTESTCD", "NVCAT", "NVSCAT", "IMAGE_RESOLUTION", "TRACER")
NV <- NV %>%
  unite("NVTESTCD", all_of(nv_grp_cols), sep = "-/") %>%
  derive_blfl_adni(
    sdtm_in = .,
    dm_domain = DM,
    tgt_var = "NVBLFL"
  ) %>%
  separate(NVTESTCD, into = nv_grp_cols, sep = "-/") %>%
  assert_non_missing(NVTESTCD)

NV <- NV %>%
  derive_study_day_adni(
    dm_domain = DM,
    domain = "NV"
  ) %>%
  derive_seq(
    tgt_var = "NVSEQ",
    rec_vars = c("USUBJID", "NVCAT", "NVSCAT", "NVTESTCD", "NVGRPID")
  ) %>%
  assign_vars_label(nv_data_dic)
```

## Laboratory Test Results (LB) {#sec-lb}

[`LB`](../reference/LB.html) dataset contains laboratory test data such as hematology, clinical chemistry and urinalysis per visit per subject. Additionally, the [`LB`](../reference/LB.html) dataset contains blood plasma biomarkers results.

```{r lb-data-dic, echo = FALSE}
lb_data_dic <- bind_rows(
  c(
    FLDNAME = "STUDYID",
    LABEL = "Study Identifier",
    TEXT = ""
  ),
  c(
    FLDNAME = "DOMAIN",
    LABEL = "Domain Abbrevation"
  ),
  c(
    FLDNAME = "USUBJID",
    LABEL = "Unique Subject Identifier"
  ),
  c(
    FLDNAME = "ORIGPROT",
    LABEL = "Original Study Protocol"
  ),
  c(
    FLDNAME = "LBGRPID",
    LABEL = "Group ID",
    TEXT = "Study Protocol of Data Collection"
  ),
  c(
    FLDNAME = "LBSEQ",
    LABEL = "Sequence Number"
  ),
  c(
    FLDNAME = "LBTESTCD",
    LABEL = "Lab Test or Examination Short Name"
  ),
  c(
    FLDNAME = "LBTEST",
    LABEL = "Lab Test or Examination Name"
  ),
  # c(
  #   FLDNAME = "LBCAT",
  #   LABEL = "Category for Lab Test"
  # ),
  # c(
  #   FLDNAME = "LBSCAT",
  #   LABEL = "Subcategory for Lab Test"
  # ),
  c(
    FLDNAME = "LBORRES",
    LABEL = "Result or Finding in Original Units"
  ),
  c(
    FLDNAME = "LBORRESU",
    LABEL = "Original Units"
  ),
  c(
    FLDNAME = "LBSTRESC",
    LABEL = "Character Result/Finding in Std Format"
  ),
  c(
    FLDNAME = "LBSTRESN",
    LABEL = "Numeric Result/Finding in Standard Units"
  ),
  c(
    FLDNAME = "LBSTRESU",
    LABEL = "Standard Units"
  ),
  c(
    FLDNAME = "LBSTAT",
    LABEL = "Completion Status"
  ),
  c(
    FLDNAME = "LBREASND",
    LABEL = "Reason Test Not Done"
  ),
  c(
    FLDNAME = "LBSPEC",
    LABEL = "Specimen Type"
  ),
  c(
    FLDNAME = "LBSPCCND",
    LABEL = "Specimen Condition"
  ),
  # c(
  #   FLDNAME = "LBMETHOD",
  #   LABEL = "Method of Test or Examination"
  # ),
  c(
    FLDNAME = "LBNAM",
    LABEL = "Vendor Name"
  ),
  c(
    FLDNAME = "LBBLFL",
    LABEL = "Baseline Flag"
  ),
  c(
    FLDNAME = "VISITNUM",
    LABEL = "Visit Number"
  ),
  c(
    FLDNAME = "VISIT",
    LABEL = "Visit Name"
  ),
  c(
    FLDNAME = "VISCODE",
    LABEL = "Visit Code"
  ),
  # c(
  #   FLDNAME = "VISITDY",
  #   LABEL = "(Planned*) Study Day of Visit"
  # ),
  c(
    FLDNAME = "EPOCH",
    LABEL = "Epoch"
  ),
  c(
    FLDNAME = "LBDTC",
    LABEL = "Date/Time of Specimen Collection"
  ),
  c(
    FLDNAME = "LBDY",
    LABEL = "Study Day of Specimen Collection"
  )
) %>%
  tibble::as_tibble() %>%
  relocate(c(FLDNAME, LABEL, TEXT)) %>%
  mutate(TEXT = replace_na(TEXT, " ")) %>%
  mutate(
    DERIVED = TRUE,
    TBLNAME = "LB",
    CRFNAME = "[ Derived ] Laboratory Test Results"
  )
```

<details>
<summary>Show `LB` Variables</summary>

```{r lb-data-dic-print, eval = TRUE, echo = FALSE}
lb_data_dic %>%
  select(FLDNAME, LABEL, TEXT) %>%
  datatable(., paging = TRUE)
```

</details>

```{r lb-prepare-labdata-1go2, eval = TRUE, echo = FALSE}
common_cols <- c(
  "ORIGPROT", "COLPROT", "ID", "PTID", "RID", "SITEID", "VISCODE", "VISCODE2",
  "USERDATE", "USERDATE2", "RECNO", "ACCNO", "COVVIS", "EXAMDATE", "update_stamp"
)
lab_cols <- c(
  "ORIGPROT", "COLPROT", "RID", "SITEID",
  "VISCODE", "VISCODE2", "EXAMDATE"
)

# Lab data for ANDI1-GO-2 phases
ADNI1GO2_CLINICAL_LABDATA <- LABDATA %>%
  generate_oak_id_vars_adni(raw_src = "LABDATA") %>%
  mutate(across(-all_of(c(common_cols, oak_id_vars())), as.character)) %>%
  pivot_longer(
    cols = everything() & -all_of(c(common_cols, oak_id_vars())),
    names_to = "LBTESTCD",
    values_to = "LBORRES"
  ) %>%
  select(-all_of(common_cols[!common_cols %in% lab_cols])) %>%
  filter(LBORRES != -1) %>%
  mutate(LBDTC = create_iso8601(as.character(EXAMDATE), .format = "y-m-d")) %>%
  adjust_lab_visitcode()
```

```{r lb-prepare-labdata-34, eval = TRUE, echo = FALSE, message = FALSE, results = 'hold'}
# Lab data for ANDI3-4 phases
ADNI34_CLINICAL_LABDATA <- URMC_LABDATA %>%
  generate_oak_id_vars_adni(raw_src = "URMC_LABDATA") %>%
  mutate(
    LBNAM = "URMC",
    TestID = ifelse(TestName %in% "Sodium", "NA", TestID)
  ) %>%
  mutate(across(ends_with(c("Date", "Time")), as.character))

# Create LBDTC
ADNI34_CLINICAL_LABDATA <- sdtm.oak::assign_datetime(
  tgt_dat = ADNI34_CLINICAL_LABDATA %>%
    select(-SampleDate, -SampleTime),
  raw_dat = ADNI34_CLINICAL_LABDATA,
  tgt_var = "LBDTC",
  raw_var = c("SampleDate", "SampleTime"),
  raw_fmt = c("y-m-d", "H:M:S")
)
```

```{r lb-prepare-labdata-34-cont, eval = TRUE, echo = FALSE}
ADNI34_CLINICAL_LABDATA <- ADNI34_CLINICAL_LABDATA %>%
  # Adjust for lab test that were considered as 'not completed/done'
  adjust_lab_status() %>%
  adjust_lab_visitcode() %>%
  mutate(
    LBGRPID = COLPROT,
    LBTESTCD = TestID,
    LBTEST = TestName,
    LBORRES = ResultValueConv_translated,
    LBORRESU = UnitsConv,
    LBORNRLO = LowerRangeConv,
    LBORNRHI = UpperRangeConv,
    LBSTRESC = ResultValueSI_translated,
    LBSTRESN = as.numeric(ResultValueSI_translated),
    LBSTRESU = UnitsSI,
    LBSTNRLO = LowerRangeSI,
    LBSTNRHI = UpperRangeSI,
    LBFAST = Fasting,
    LBSPCCND = Comments
  )
```

```{r lb-prepare-labdata-input, eval = TRUE, echo = FALSE}
CLINICAL_LB_TESTCD_LIST <- DATADIC %>%
  filter(TBLNAME %in% "LABDATA") %>%
  filter(!FLDNAME %in% common_cols) %>%
  distinct(FLDNAME, TEXT) %>%
  assert_uniq(FLDNAME) %>%
  mutate(
    LBTESTCD = FLDNAME,
    LBTEST = str_remove_all(TEXT, paste0("Test |", FLDNAME, ";|\\(\\%\\)")),
    LBTEST = str_trim(LBTEST),
    LBORRESU = str_extract_all(TEXT, "\\(\\%\\)", simplify = TRUE),
    LBORRESU = str_remove_all(LBORRESU, "[\\(\\)]"),
    LBSTRESU = LBORRESU
  ) %>%
  select(LBTESTCD, LBTEST, LBORRESU, LBSTRESU) %>%
  mutate(across(everything(), ~ ifelse(is.na(.x), "", .x))) %>%
  filter(LBTESTCD %in% ADNI34_CLINICAL_LABDATA$LBTESTCD)
```

The [`LB`](../reference/LB.html) dataset contains the following blood plasma biomarker results in addition to the safety laboratory data. 

*NOTE: *

* As a future work, to include the sample batch of `UPENNBIOMK_MASTER` and `UPENNBIOMK_ROCHE_ELECSYS`.

* Also to harmonized the safety laboratory data across different study phase.

```{r lb-testcd-list, eval = TRUE, echo = FALSE}
LB_TESTCD_LIST <- bind_rows(
  CLINICAL_LB_TESTCD_LIST,
  get_biomarker_details(assay = "C2N"),
  get_biomarker_details(assay = "Roche"),
  get_biomarker_details(assay = "AlzBio3"),
  get_biomarker_details(assay = "FQ")
) %>%
  select(SOURCE, LBNAM, LBTESTCD, LBTEST, LBORRESU, LBSTRESU, LBMETHOD)

LB_TESTCD_LIST %>%
  arrange(LBTESTCD) %>%
  filter(!is.na(LBNAM)) %>%
  arrange(SOURCE, LBNAM, LBMETHOD) %>%
  select(SOURCE, LBNAM, LBMETHOD, LBTESTCD, LBTEST, LBORRESU, LBSTRESU) %>%
  datatable(paging = TRUE, filter = "top")
```

```{r lb-prepare-labdata-print, echo = TRUE, warning = FALSE, message = FALSE, results = 'asis', ref.label = c("lb-prepare-labdata-1go2", "lb-prepare-labdata-34", "lb-prepare-labdata-34-cont")}
```

```{r lb-prepare-upenbiomk1, eval = TRUE, echo = FALSE}
# Roche Elecsys Immunioassay
upenn_roche_cols <- c("ABETA40", "ABETA42", "TAU", "PTAU")
names(upenn_roche_cols) <- c("ABETA40", "ABETA42", "TPROT", "PTAU181")
join_var_upenn <- paste0(c("ORIGPROT", "COLPROT", "RID", "EXAMDATE"), "_BM")
names(join_var_upenn) <- str_remove_all(join_var_upenn, "\\_BM")

UPENNBIOMK_ROCHE_DATA <- UPENNBIOMK_ROCHE_ELECSYS %>%
  select(
    ORIGPROT, COLPROT, RID, VISCODE2, EXAMDATE, RUNDATE, BATCH, COMMENT,
    all_of(as.character(upenn_roche_cols))
  ) %>%
  generate_oak_id_vars_adni(raw_src = "UPENNBIOMK_ROCHE_ELECSYS") %>%
  # Fuzzy join for actual study visit name/code
  left_fuzzy_join(
    data1 = .,
    data2 = BIOMARK %>%
      select(ORIGPROT, COLPROT, RID, EXAMDATE, VISCODE) %>%
      filter(!is.na(EXAMDATE)) %>%
      rename_with_list(., name_char = join_var_upenn, by_name = FALSE),
    join_by = join_var_upenn,
    main_cols = "EXAMDATE",
    date_col = "EXAMDATE",
    check_cols = "VISCODE",
    relation = "one-to-many"
  ) %>%
  pivot_longer(
    cols = all_of(as.character(upenn_roche_cols)),
    names_to = "LBTESTCD",
    values_to = "LBORRES"
  ) %>%
  mutate(
    LBORRES = as.character(LBORRES),
    LBDTC = create_iso8601(as.character(EXAMDATE), .format = "y-m-d")
  )

UPENNBIOMK_ROCHE_DATA <- UPENNBIOMK_ROCHE_DATA %>%
  # Adjust comment by LBTETSCD
  adjust_lab_comment(lab_data = .) %>%
  # Create lab completion status
  mutate(
    LBSTAT = case_when(!is.na(COMMENT) & is.na(LBORRES) ~ "NOT DONE"),
    LBREASND = case_when(!is.na(COMMENT) & is.na(LBORRES) ~ COMMENT),
    LBTESTCD = factor(LBTESTCD,
      levels = as.character(upenn_roche_cols),
      labels = names(upenn_roche_cols)
    ),
    LBTESTCD = as.character(LBTESTCD)
  ) %>%
  # Add lab result limit values
  left_join(
    get_biomarker_details(assay = "Roche"),
    by = c("BATCH", "LBTESTCD")
  ) %>%
  assert_non_missing(LBTEST)
```

```{r lb-prepare-upennbiomk2, eval = TRUE, echo = FALSE}
# INNO_BIA AlzBio3 Immunoassay - Research Use Only (ROU)
upenn_alzbio3_cols <- c("ABETA", "TAU", "PTAU")
upenn_alzbio3_cols_label <- c("ABETA42", "TPROT", "PTAU181")
upenn_alzbio3_cols <- c(
  upenn_alzbio3_cols,
  paste0(upenn_alzbio3_cols, "_RAW")
)
names(upenn_alzbio3_cols) <- c(
  paste0(upenn_alzbio3_cols_label, "S"),
  paste0(upenn_alzbio3_cols_label, "R")
)

UPENNBIOMK_ALZBIO3_DATA <- UPENNBIOMK_MASTER %>%
  rename(all_of(c(
    upenn_alzbio3_cols,
    c("EXAMDATE" = "DRAWDTE", "VISCODE2" = "VISCODE")
  ))) %>%
  assert_uniq(all_of(c("ORIGPROT", "RID", "EXAMDATE", "BATCH"))) %>%
  generate_oak_id_vars_adni(raw_src = "UPENNBIOMK_MASTER") %>%
  left_fuzzy_join(
    data1 = .,
    data2 = BIOMARK %>%
      select(ORIGPROT, RID, EXAMDATE, VISCODE, COLPROT) %>%
      filter(!is.na(EXAMDATE)) %>%
      rename_with_list(., name_char = join_var_upenn[-2], by_name = FALSE),
    join_by = join_var_upenn[-2],
    main_cols = "EXAMDATE",
    date_col = "EXAMDATE",
    check_cols = c("COLPROT", "VISCODE"),
    relation = "one-to-many"
  )

UPENNBIOMK_ALZBIO3_DATA <- UPENNBIOMK_ALZBIO3_DATA %>%
  pivot_longer(
    cols = all_of(names(upenn_alzbio3_cols)),
    names_to = "LBTESTCD",
    values_to = "LBORRES"
  ) %>%
  mutate(
    LBORRES = as.character(LBORRES),
    LBDTC = create_iso8601(as.character(EXAMDATE), .format = "y-m-d")
  ) %>%
  left_join(
    get_biomarker_details(assay = "AlzBio3"),
    by = "LBTESTCD"
  ) %>%
  assert_non_missing(LBTEST)
```

```{r lb-prepare-upennbiomk3, eval = TRUE, echo = FALSE}
# Fujirebio Lumipulse G1200 and Quanterix HD-X
fq_cols <- c(
  "pT217_F", "AB42_F", "AB40_F", "AB42_AB40_F",
  "pT217_AB42_F", "NfL_Q", "GFAP_Q"
)
names(fq_cols) <- c(
  "PT217", "AB42", "AB40", "AB42AB40",
  "PT217AB42", "NFL", "GFAP"
)

UPENN_PLASMA_FQ_DATA <- UPENN_PLASMA_FUJIREBIO_QUANTERIX %>%
  generate_oak_id_vars_adni(raw_src = "UPENN_PLASMA_FUJIREBIO_QUANTERIX") %>%
  select(-update_stamp) %>%
  rename(all_of(fq_cols)) %>%
  pivot_longer(
    cols = all_of(names(fq_cols)),
    names_to = "LBTESTCD",
    values_to = "LBORRES"
  ) %>%
  left_join(
    get_biomarker_details(assay = "FQ"),
    by = "LBTESTCD"
  ) %>%
  assert_non_missing(LBTEST) %>%
  mutate(
    LBORRES = as.character(as.numeric(LBORRES)),
    LBDTC = create_iso8601(as.character(EXAMDATE), .format = "y-m-d"),
    LBANTREG = Primary
  ) %>%
  filter(!is.na(VISCODE))
```

```{r lb-prepare-c2n-data}
# C2N Blood Plasma Result ----
c2n_cols <- c(
  "pT217_C2N", "npT217_C2N", "AB42_C2N", "AB40_C2N", "AB42_AB40_C2N",
  "pT217_npT217_C2N", "APS2_C2N"
)
names(c2n_cols) <- c(
  "PT217", "NPT217", "AB42", "AB40",
  "AB42AB40", "PTNPT217", "APS2"
)

C2N_PLASMA_DATA <- C2N_PRECIVITYAD2_PLASMA %>%
  generate_oak_id_vars_adni(raw_src = "C2N_PRECIVITYAD2_PLASMA") %>%
  rename(all_of(c2n_cols)) %>%
  select(
    all_of(c(oak_id_vars(), names(c2n_cols))),
    ORIGPROT, COLPROT, RID, VISCODE,
    EXAMDATE,
    LBANTREG = Primary, LBSPCCND = Comments
  ) %>%
  mutate(across(all_of(names(c2n_cols)), as.character)) %>%
  pivot_longer(
    cols = all_of(names(c2n_cols)),
    names_to = "LBTESTCD",
    values_to = "LBORRES"
  ) %>%
  mutate(
    LBSPEC = "PLASMA",
    LBDTC = create_iso8601(as.character(EXAMDATE), .format = "y-m-d")
  ) %>%
  left_join(
    get_biomarker_details(assay = "C2N"),
    by = "LBTESTCD"
  ) %>%
  assert_non_missing(LBTEST)
```

```{r generate-lb}
LB <- bind_rows(
  C2N_PLASMA_DATA, ADNI1GO2_CLINICAL_LABDATA,
  UPENNBIOMK_ROCHE_DATA, UPENNBIOMK_ALZBIO3_DATA,
  UPENN_PLASMA_FQ_DATA
) %>%
  mutate(
    LBGRPID = COLPROT,
    LBSTRESC = as.character(LBORRES),
    LBSTRESN = as.numeric(LBORRES)
  ) %>%
  bind_rows(ADNI34_CLINICAL_LABDATA) %>%
  assert_non_missing(LBGRPID) %>%
  # Future work to add lab result category
  # set_dom_test(
  #    .data_list = LB_TESTCD_LIST %>%
  #     select(LBTESTCD, LBTEST, LBORRESU, LBSTRESU),
  #   merge_by = "LBTESTCD"
  # ) %>%
  derive_usubjid() %>%
  assign_studyid_domain(domain = "LB") %>%
  assign_visit_attr(check_missing_visnum = FALSE) %>%
  assign_epoch() %>%
  derive_blfl_adni(
    dm_domain = DM,
    tgt_var = "LBBLFL",
    .strict = FALSE
  ) %>%
  derive_study_day_adni(
    dm_domain = DM,
    domain = "LB"
  ) %>%
  derive_seq(
    tgt_var = "LBSEQ",
    rec_vars = c("USUBJID", "LBTESTCD", "LBGRPID")
  ) %>%
  assign_vars_label(data_dict = lb_data_dic, .strict = FALSE)
```

## Genomics Findings (GF) {#sec-gf}

[`GF`](../reference/GF.html) dataset contains data related to genomic material of interest. The [`GF`](../reference/GF.html) dataset will contains subjects' APOE genotype that collected once during the study period. 

*NOTE: * There might be some instance with duplicated APOE genotype records per subject. 
```{r gf-data-dic, echo = FALSE}
gf_data_dic <- bind_rows(
  c(
    FLDNAME = "STUDYID",
    LABEL = "Study Identifier",
    TEXT = ""
  ),
  c(
    FLDNAME = "DOMAIN",
    LABEL = "Domain Abbrevation"
  ),
  c(
    FLDNAME = "USUBJID",
    LABEL = "Unique Subject Identifier"
  ),
  c(
    FLDNAME = "GFSEQ",
    LABEL = "Sequence Number"
  ),
  c(
    FLDNAME = "GFGRPID",
    LABEL = "Group ID",
    TEXT = "Study Protocol of Data Collection"
  ),
  c(
    FLDNAME = "GFTESTCD",
    LABEL = "Short Name of Genomic Measurement"
  ),
  c(
    FLDNAME = "GFTEST",
    LABEL = "Name of Genomic Measurement"
  ),
  c(
    FLDNAME = "GFSTDTL",
    LABEL = "Measurement, Test, or Examination Detail"
  ), # GENOTYPE
  c(
    FLDNAME = "GFORRES",
    LABEL = "Result or Finding in Original Units"
  ),
  c(
    FLDNAME = "GFSTRESC",
    LABEL = "Character Result/Finding in Std Format"
  ),
  c(
    FLDNAME = "VISITNUM",
    LABEL = "Visit Number"
  ),
  c(
    FLDNAME = "VISIT",
    LABEL = "Visit Name"
  ),
  # c(
  #   FLDNAME = "VISITDY",
  #   LABEL = "(Planned*) Study Day of Visit"
  # ),
  c(
    FLDNAME = "GFDTC",
    LABEL = "Date/Time of Specimen Collection"
  ),
  c(
    FLDNAME = "GFDY",
    LABEL = "Study Day of Specimen Collection"
  )
) %>%
  tibble::as_tibble() %>%
  relocate(c(FLDNAME, LABEL, TEXT)) %>%
  mutate(TEXT = replace_na(TEXT, " ")) %>%
  mutate(
    DERIVED = TRUE,
    TBLNAME = "GF",
    CRFNAME = "[ Derived ] Genomics Findings"
  )
```

<details>
<summary>Show `GF` Variables</summary>

```{r gf-data-dic-print, eval = TRUE, echo = FALSE}
gf_data_dic %>%
  select(FLDNAME, LABEL, TEXT) %>%
  datatable(., paging = TRUE)
```

</details>

```{r generate-gf}
GF <- APOERES %>%
  assert_uniq(RID) %>%
  generate_oak_id_vars_adni(raw_src = "APOERES") %>%
  mutate(
    GFTESTCD = "APOE",
    GFTEST = "Apolipoprotein E",
    GFGRPID = COLPROT,
    GFSTDTL = "GENOTYPE",
    GFORRES = GENOTYPE,
    GFSTRESC = str_replace_all(GENOTYPE, "(\\d+)", "\\1"),
    GFDTC = create_iso8601(as.character(APTESTDT), .format = "y-m-d"),
  ) %>%
  derive_usubjid() %>%
  assign_studyid_domain(domain = "GF") %>%
  assign_visit_attr(check_missing_visnum = FALSE) %>%
  derive_study_day_adni(
    dm_domain = DM,
    domain = "GF"
  ) %>%
  derive_seq(
    tgt_var = "GFSEQ",
    rec_vars = c("USUBJID", "GFTESTCD", "GFGRPID")
  ) %>%
  assign_vars_label(data_dict = gf_data_dic)
```

## Vital Sign (VS) {#sec-vs}

[`VS`](../reference/VS.html) dataset contains measurement finding of blood pressure, heart beat, respiratory rate, temperature, weight and height per visit per subject.

```{r vs-data-dic, echo = FALSE}
vs_data_dic <- bind_rows(
  c(
    FLDNAME = "STUDYID",
    LABEL = "Study Identifier",
    TEXT = ""
  ),
  c(
    FLDNAME = "DOMAIN",
    LABEL = "Domain Abbrevation"
  ),
  c(
    FLDNAME = "USUBJID",
    LABEL = "Unique Subject Identifier"
  ),
  c(
    FLDNAME = "VSSEQ",
    LABEL = "Sequence Number"
  ),
  c(
    FLDNAME = "VSGRPID",
    LABEL = "Group ID",
    TEXT = "Study Protocol of Data Collection"
  ),
  c(
    FLDNAME = "VSTESTCD",
    LABEL = "Vital Signs Test Short Name"
  ),
  c(
    FLDNAME = "VSTEST",
    LABEL = "Vital Signs Test Name"
  ),
  c(
    FLDNAME = "VSORRES",
    LABEL = "Result or Finding in Original Units"
  ),
  c(
    FLDNAME = "VSORRESU",
    LABEL = "Original Units"
  ),
  c(
    FLDNAME = "VSSTRESC",
    LABEL = "Character Result/Finding in Std Format"
  ),
  c(
    FLDNAME = "VSSTRESN",
    LABEL = "Numeric Result/Finding in Standard Units"
  ),
  c(
    FLDNAME = "VSSTRESU",
    LABEL = "Standard Units"
  ),
  c(
    FLDNAME = "VSLOC",
    LABEL = "Location of Vital Signs Measurment"
  ),
  c(
    FLDNAME = "VSBLFL",
    LABEL = "Baseline Flag"
  ),
  c(
    FLDNAME = "VSDRVFL",
    LABEL = "Derived Flag"
  ),
  c(
    FLDNAME = "VISITNUM",
    LABEL = "Visit Number"
  ),
  c(
    FLDNAME = "VISIT",
    LABEL = "Visit Name"
  ),
  # c(
  #   FLDNAME = "VISITDY",
  #   LABEL = "(Planned*) Study Day of Visit"
  # ),
  c(
    FLDNAME = "EPOCH",
    LABEL = "Epoch"
  ),
  c(
    FLDNAME = "VSDTC",
    LABEL = "Date/Time of Measurements"
  ),
  c(
    FLDNAME = "VSDY",
    LABEL = "Study Day of Vital Signs"
  )
) %>%
  tibble::as_tibble() %>%
  relocate(c(FLDNAME, LABEL, TEXT)) %>%
  mutate(TEXT = replace_na(TEXT, " ")) %>%
  mutate(
    DERIVED = TRUE,
    TBLNAME = "VS",
    CRFNAME = "[ Derived ] Vital Signs"
  )
```

<details>
<summary>Show `VS` Variables</summary>

```{r vs-data-dic-print, eval = TRUE, echo = FALSE}
vs_data_dic %>%
  select(FLDNAME, LABEL, TEXT) %>%
  datatable(., paging = TRUE)
```

</details>

```{r vs-testcd-list, eval = TRUE, echo = FALSE}
# Required checking variable name length and test labels
VS_TESTCD_LIST <- bind_rows(
  c(
    VSTESTCD = "WEIGHT",
    VSTEST = "Weight",
    VSORRESU = "Pound (LB) or Kilogram (kg)",
    VSSTRESU = "kg"
  ),
  c(
    VSTESTCD = "HEIGHT",
    VSTEST = "Height",
    VSORRESU = "Centimeter (cm) or Inches (inch)",
    VSSTRESU = "m"
  ),
  c(
    VSTESTCD = "TEMP",
    VSTEST = "Temperature",
    VSORRESU = "Fahrenheit (F) or Celsius (C)",
    VSSTRESU = "C"
  ),
  c(
    VSTESTCD = "DIABP",
    VSTEST = "Diastolic Blood Pressure",
    VSORRESU = "mmHg",
    VSSTRESU = "mmHg"
  ),
  c(
    VSTESTCD = "SYSBP",
    VSTEST = "Diastolic Blood Pressure",
    VSORRESU = "mmHg",
    VSSTRESU = "mmHg"
  ),
  c(
    VSTESTCD = "PLUSE",
    VSTEST = "Pluse Rate",
    VSORRESU = "beats/min",
    VSSTRESU = "beats/min"
  ),
  c(
    VSTESTCD = "RESP",
    VSTEST = "Respiratory Rate",
    VSORRESU = "breaths/min",
    VSSTRESU = "breaths/min"
  ),
) %>%
  mutate(SOURCE = "VITALS")

VS_TESTCD_LIST %>%
  arrange(SOURCE, VSTESTCD) %>%
  select(SOURCE, VSTESTCD, VSTEST, VSORRESU, VSSTRESU) %>%
  datatable()
```

```{r generate-vs0}
# Wide format
VS <- VITALS %>%
  mutate(VSDTC = create_iso8601(as.character(VISDATE), .format = "y-m-d")) %>%
  select(
    ORIGPROT, COLPROT, RID, VISCODE, VSDTC,
    VSWEIGHT, VSWTUNIT, VSHEIGHT, VSHTUNIT, VSBPSYS, VSBPDIA, VSPULSE,
    VSRESP, VSTEMP, VSTMPSRC, VSTMPUNT, VSHGTSC
  ) %>%
  generate_oak_id_vars_adni(raw_src = "VITALS") %>%
  mutate(across(c(VSWEIGHT, VSHEIGHT, VSTEMP), ~ ifelse(.x == -1, NA, .x))) %>%
  verify(all(VSWTUNIT %in% c("kilograms", "pounds") | is.na(VSWTUNIT))) %>%
  verify(all(VSHTUNIT %in% c("centimeters", "inches") | is.na(VSHTUNIT))) %>%
  verify(all(VSTMPUNT %in% c("Fahrenheit", "Celsius") | is.na(VSTMPUNT))) %>%
  mutate(
    VSWTUNIT_TRANSLATED = case_when(
      VSWTUNIT %in% "kilograms" ~ "kg",
      VSWTUNIT %in% "pounds" ~ "LB"
    ),
    VSHTUNIT_TRANSLATED = case_when(
      VSHTUNIT %in% "centimeters" ~ "cm",
      VSHTUNIT %in% "inches" ~ "inch"
    ),
    VSTMPUNT_TRANSLATED = case_when(
      VSTMPUNT %in% "Fahrenheit" ~ "F",
      VSTMPUNT %in% "Celsius" ~ "C"
    ),
  ) %>%
  unite("WEIGHT", VSWEIGHT, VSWTUNIT_TRANSLATED, sep = "-") %>%
  unite("HEIGHT", VSHEIGHT, VSHTUNIT_TRANSLATED, sep = "-") %>%
  unite("TEMP", VSTEMP, VSTMPUNT_TRANSLATED, VSTMPSRC, sep = "-") %>%
  mutate(
    DIABP = ifelse(!is.na(VSBPDIA), paste0(VSBPDIA, "-", "mmHg"), NA),
    SYSBP = ifelse(!is.na(VSBPSYS), paste0(VSBPSYS, "-", "mmHg"), NA),
    PLUSE = ifelse(!is.na(VSPULSE), paste0(VSPULSE, "-", "beats/min"), NA),
    RESP = ifelse(!is.na(VSRESP), paste0(VSRESP, "-", "breaths/min"), NA)
  )
```

```{r generate-vs1}
# Long format
VS <- VS %>%
  pivot_longer(
    cols = c(WEIGHT, HEIGHT, TEMP, DIABP, SYSBP, PLUSE, RESP),
    names_to = "VSTESTCD",
    values_to = "VALUE"
  ) %>%
  set_dom_test(
    .data_list = VS_TESTCD_LIST %>%
      select(VSTESTCD, VSTEST),
    merge_by = "VSTESTCD"
  ) %>%
  separate(VALUE, into = c("VSORRES", "VSORRESU", "VSLOC"), sep = "-") %>%
  mutate(VSLOC = str_to_upper(VSLOC)) %>%
  mutate(VSSTRESU = case_when(
    VSORRESU %in% c("F", "C") ~ "C",
    VSORRESU %in% c("cm", "inch") ~ "cm",
    VSORRESU %in% c("kg", "LB") ~ "kg",
    VSORRESU %in% c("mmHg", "beats/min", "breaths/min") ~ VSORRESU
  )) %>%
  # Unit conversion: conv_unit
  mutate(
    FROM_UNIT = case_when(
      VSORRESU %in% "LB" ~ "lbs",
      VSORRESU %in% c("C", "F", "cm", "inch", "kg") ~ VSORRESU,
      TRUE ~ NA_character_
    ),
    TO_UNIT = case_when(
      VSSTRESU %in% c("C", "kg", "cm") ~ VSSTRESU,
      TRUE ~ NA_character_
    )
  ) %>%
  rowwise() %>%
  mutate(
    VSSTRESN = ifelse(
      VSTESTCD %in% c("WEIGHT", "HEIGHT", "TEMP") & !is.na(FROM_UNIT),
      conv_unit(as.numeric(VSORRES), from = FROM_UNIT, to = TO_UNIT),
      ifelse(VSTESTCD %in% c("DIABP", "SYSBP", "PLUSE", "RESP"),
        as.numeric(VSORRES), NA_real_
      )
    )
  ) %>%
  mutate(
    VSGRPID = COLPROT,
    VSSTRESN = round(VSSTRESN, digits = 1),
    VSSTRESC = as.character(VSSTRESN),
    VSDRVFL = case_when(!is.na(FROM_UNIT) & FROM_UNIT != TO_UNIT ~ "Yes")
  ) %>%
  as_tibble() %>%
  assert_non_missing(COLPROT, VISCODE, VSTEST)
```

```{r generate-vs2}
VS <- VS %>%
  derive_usubjid() %>%
  assign_studyid_domain(domain = "VS") %>%
  assign_visit_attr() %>%
  assign_epoch() %>%
  derive_blfl_adni(
    dm_domain = DM,
    tgt_var = "VSBLFL"
  ) %>%
  derive_study_day_adni(
    dm_domain = DM,
    domain = "VS"
  ) %>%
  derive_seq(
    tgt_var = "VSSEQ",
    rec_vars = c("USUBJID", "VSTESTCD", "VSGRPID")
  ) %>%
  assign_vars_label(data_dict = vs_data_dic)
```

```{r}
```
