---
title: "ADNI-Longitudinal"
output: 
   rmarkdown::html_vignette:
     toc: true
     toc_depth: 3
     code_folding: hide
date: "Last Updated: `r format(Sys.Date(), '%B %d, %Y')`"
vignette: >
  %\VignetteIndexEntry{ADNI-Longitudinal}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = TRUE,
  message = TRUE,
  echo = TRUE,
  fig.width = 7,
  fig.height = 5,
  class.source = "fold-show",
  result = "asis",
  dpi = 100,
  dev = "png"
)
```

# Introduction

This article demonstrate how to use the `ADNIMERGE2` R package to generate simple longitudinal summaries of clinical cognitive outcomes. The analysis results that are presented in this article are merely for illustration purpose.

# Load Required R Packages

```{r setup-libraries, warning = FALSE, message = FALSE}
library(tidyverse)
library(labelled)
library(ggplot2)
library(splines)
library(lme4)
library(marginaleffects)
library(ADNIMERGE2)
```

```{r setup-parameters}
theme_set(theme_bw(base_size = 12))
# Color for DX group
dx_color_pal <- c("#73C186", "#F2B974", "#DF957C", "#999999")
```

# ADNI Longitudinal Summaries - Clinical Cognitive Outcomes

## ADAS-Cog Item-13 Total Score

The following r chucks are used to summarize the ADAS-Cognitive Behavior item-13 total score (ADAS-Cog) over time. Individual's score trajectory by corresponding baseline diagnostics status, summary result of a hierarchical model fit, and predicted scores overtime by baseline diagnosis status are presented in this section. 

### Data Preparation

The analysis will based on enrolled subjects in the ADNI study with known baseline diagnostics summary and that have at least one assessment score. The analysis population is summarized in the following diagram.

```{r consort-diagram, echo = FALSE, fig.width = 8.5, fig.alt = "Analysis Population Diagram"}
# Data prep for consort diagram
consort_data <- ADSL %>%
  filter(ENRLFL %in% "Y") %>%
  select(USUBJID, ENRLFL, DX) %>%
  left_join(
    ADQS %>%
      filter(PARAMCD %in% "ADASTT13") %>%
      group_by(USUBJID) %>%
      summarize(NUM_NON_MISSING = sum(!is.na(AVAL))) %>%
      ungroup(),
    by = "USUBJID"
  ) %>%
  assert_uniq(USUBJID) %>%
  mutate(NUM_NON_MISSING = ifelse(is.na(NUM_NON_MISSING), 0, NUM_NON_MISSING)) %>%
  mutate(
    enrolled = case_when(ENRLFL %in% "Y" ~ "Enrolled"),
    not_bl_dx = case_when(is.na(DX) ~ "Have no a baseline diagnostics summary"),
    bl_dx = factor(
      case_when(!is.na(DX) ~ as.character(DX)),
      levels = c("CN", "MCI", "DEM"),
      labels = c("Cognitive Normal (CN)\n", "Mild Cognitive Impairment (MCI)\n", "Dementia (DEM)")
    ),
    missing_score_value = case_when(
      NUM_NON_MISSING == 0 | is.na(NUM_NON_MISSING) ~ "Have no at least one ADAS-Cog\n Item-13 Score"
    ),
    known_score_value = ifelse(NUM_NON_MISSING != 0, "Analysis Population", NA_character_)
  )

consort::consort_plot(
  data = consort_data,
  orders = c(
    enrolled = "Enrolled Population",
    not_bl_dx = "Excluded",
    bl_dx = "Have a Baseline \n Diagnostics Summary",
    missing_score_value = "Excluded",
    known_score_value = "Analysis Population"
  ),
  side_box = c("not_bl_dx", "missing_score_value"),
  allocation = "bl_dx",
  cex = 0.8
)
```

*NOTE: * Records with missing time variable (`TIME`) are also excluded from the analysis population as shown in the following r chunk.

```{r data-prep}
# Prepare analysis dataset of ADAS-cog item-13 score
ADADAS <- ADQS %>%
  # Enrolled participant
  filter(ENRLFL %in% "Y") %>%
  # ADAS-cog item-13 total score
  filter(PARAMCD %in% "ADASTT13") %>%
  # Compute time variable in years
  mutate(
    TIME = convert_number_days(AVISITN, unit = "year"),
    DX = factor(DX, levels = levels(ADSL$DX))
  ) %>%
  # For spline term in the model
  filter(!if_any(all_of(c("TIME", "DX", "AVAL")), ~ is.na(.x)))
```

```{r check-analysis-popu, echo = FALSE}
# Total analysis population - consort data
num_final_popu_consort <- consort_data %>%
  filter(!is.na(bl_dx) & !is.na(known_score_value)) %>%
  nrow()

# Total analysis population - ADADAS data
num_final_popu_adadas <- ADADAS %>%
  distinct(USUBJID) %>%
  nrow()

if (num_final_popu_adadas != num_final_popu_consort) {
  stop("Number of analysis population is not the same!")
}
```

### Individual Profile Plot

The individual profile of ADAS-Cog item-13 total score over time by baseline diagnostics group is presented below.

```{r indiv-profile-plot, warning = FALSE, fig.alt = "Individual profile plot"}
# Individual profile (spaghetti) plot
individual_profile_plot <- ADADAS %>%
  ggplot(aes(x = TIME, y = AVAL, group = USUBJID, color = DX)) +
  geom_line(alpha = 0.75) +
  scale_x_continuous(breaks = seq(0, max(ADADAS$TIME, na.rm = TRUE), 2)) +
  scale_color_manual(values = dx_color_pal) +
  labs(
    y = "ADAS-cog Item-13 Total Score",
    x = "Timeline since baseline visit (in years)",
    color = "Baseline Diagnostics Status",
    caption = paste0(
      "Based on known baseline diagnostics status ",
      "(i.e., any missing diagnostics status is excluded)"
    ),
    plot.caption = element_text(hjust = 0)
  ) +
  theme(legend.position = "bottom")
individual_profile_plot
```

### Model Fit

A linear mixed model with spline term of continuous time variable can be used to analysis the longitudinal ADAS-Cog score using `lme4` R package. 

The following terms are included in the mixed effect model: 

* Main effect of baseline diagnostics status (`DX`): To account any baseline difference of ADAS-cog item-13 total score among diagnostics status. In a clinical trial study, it is often to assume the same subject characteristics between treatment group at baseline visit which is a strong assumption. Unlike a clinical trial study, subject characteristics might be different at baseline visit in an observation study. Thus, this term is included here due to the design of ADNI study which is an observational study.

* Interaction term between time and baseline diagnostics status (`TIME:DX`): To account for any difference of the ADAS-cog assessment item-13 total score among diagnostics status overtime. 

* Random intercept and random slope terms (`TIME|USUBJID`): To account the repeated measurements per subject. It seems subjects have different baseline ADAS-cog item-13 total score as shown in the above individual profile plot (random intercept: `|USUBJID`) and subject might also have different trajectory overtime (`TIME|`). 

* Spline terms (`ns`): To considered any non-linear progression trajectory in the model (i.e., non-linear effect of time in the model).

```{r spline-term}
# Spline term defined in the global environment
ns21 <- function(t) {
  as.numeric(predict(splines::ns(ADADAS$TIME,
    df = 2,
    Boundary.knots = c(0, max(ADADAS$TIME))
  ), t)[, 1])
}
ns22 <- function(t) {
  as.numeric(predict(splines::ns(ADADAS$TIME,
    df = 2,
    Boundary.knots = c(0, max(ADADAS$TIME))
  ), t)[, 2])
}
assign("ns21", ns21, envir = .GlobalEnv)
assign("ns22", ns22, envir = .GlobalEnv)
```

```{r model-fit, warning = TRUE, message = TRUE}
# Fit linear mixed effect model using spline terms to account
#  non-linear trend of time effect

# With a random slope term ----
lmer_mod_fit <- lme4::lmer(
  formula = AVAL ~ (I(ns21(TIME)) + I(ns22(TIME))) * DX + (TIME | USUBJID),
  data = ADADAS,
  control = lmerControl(optimizer = "Nelder_Mead")
)

lmer_mod_fit
```

* **Model Assumption Check: ** [`performance::check_model`](https://easystats.github.io/performance/reference/check_model.html) function could be used to generate model assumption check (i.e., model diagnostic) plots.

```{r model-check, fig.alt = "Posterior predictive plot - model assumptions check"}
library(performance)
# To relabel x-axis and getting ggplot2 object, set panel = FALSE
model_check <- performance::check_model(
  x = lmer_mod_fit,
  check = "pp_check",
  panel = FALSE
)
model_check <- plot(model_check)
model_check$PP_CHECK +
  labs(x = "ADAS-cog Item-13 Total Score")
```

### Population-Level Model Prediction

*NOTE: * 

 * The prediction timeline for Dementia (DEM) diagnostics status was limited to two years due to the anticipated follow-up collection of ADAS cognitive behavior assessment which is up-to 2 years in the study. However, ADAS-cog assessment might have been collected for rollover subjects with DEM baseline diagnostics status during their initial visit in the next study phase. 
      
```{r model-predict, fig.width = 7, fig.alt = "Predicted score by baseline diagnostics status"}
# Population level prediction
pred_value <- marginaleffects::predictions(
  model = lmer_mod_fit,
  conf_level = 0.95,
  re.form = NA,
  newdata = expand_grid(
    TIME = seq(0, max(ADADAS$TIME), by = 0.03),
    DX = levels(ADSL$DX),
    USUBJID = NA
  ) %>%
    filter(!(DX %in% "DEM" & TIME > 2))
)

predicted_plot <- pred_value %>%
  ggplot(aes(x = TIME, y = estimate, color = DX)) +
  geom_line() +
  geom_ribbon(
    aes(ymin = conf.low, ymax = conf.high, fill = DX),
    alpha = 0.15, linetype = 0, show.legend = FALSE
  ) +
  scale_x_continuous(breaks = seq(0, max(pred_value$TIME, na.rm = TRUE), 2)) +
  scale_fill_manual(values = dx_color_pal) +
  scale_color_manual(values = dx_color_pal) +
  labs(
    x = "Timeline since baseline visit (in years)",
    y = "Predicted ADAS-cog Item-13 \n Total Score with 95% CI",
    color = "Baseline Diagnostics Status",
    fill = NULL,
    caption = paste0(
      "The prediction timeline for Dementia (DEM) diagnostics status was limited",
      " to 2 years due to the\n anticipated follow-up assessment collection ",
      "in the study."
    )
  ) +
  theme(
    legend.position = "bottom",
    plot.caption = element_text(hjust = 0)
  )
predicted_plot
```
