---
title: "ADNIMERGE2-Analysis-Meta-Specs"
output: 
   rmarkdown::html_vignette:
     toc: true
     code_folding: hide
params: 
  INCLUDE_PACC: true
date: "Last Updated: `r format(Sys.Date(), '%B %d, %Y')`"
vignette: >
  %\VignetteIndexEntry{ADNIMERGE2-Analysis-Meta-Specs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  class.source = "fold-show"
)
```

```{r utils, echo = FALSE, include = FALSE}
# Data print function
datatable <- function(data, paging = FALSE, searchable = TRUE, bInfo = FALSE, ...) {
  require(DT)
  DT::datatable(
    data = data,
    options = list(paging = paging, searchable = searchable, bInfo = bInfo, ...),
    ...
  )
}

# Unique values
assert_uniq <- function(...) {
  require(assertr)
  assert_rows(
    row_reduction_fn = col_concat,
    predicate = is_uniq,
    ...
  )
}

# Non missing value
assert_non_missing <- function(...) {
  require(assertr)
  assert(
    predicate = not_na,
    ...
  )
}

# Check common variables in a metcore
check_inconsistent_common <- function(metacore, show_mismatch = FALSE) {
  require(tidyverse)
  require(tibble)
  require(tidyr)

  if (!is_metacore(metacore)) stop("`metacore` must be a R6-Wrapper object.")
  if (!is.logical(show_mismatch)) stop("`show_mismatch` must be a boolen value.")

  # Common variable list
  common_var <- metacore$var_spec %>%
    as_tibble() %>%
    {
      if ("common" %in% colnames(.)) {
        filter(., common == TRUE)
      } else {
        filter(., row_number() %in% 0)
      }
    } %>%
    distinct(variable) %>%
    pull()

  expected_ds_var <- expand_grid(
    dataset = as.character(metacore$ds_spec$dataset),
    variable = as.character(common_var)
  ) %>%
    mutate(dataset.var = paste0(dataset, ".", variable))

  actual_ds_var <- metacore$ds_vars %>%
    as_tibble() %>%
    verify(nrow(.) > 0) %>%
    filter(variable %in% common_var) %>%
    mutate(dataset.var = paste0(dataset, ".", variable)) %>%
    distinct(dataset.var) %>%
    pull(dataset.var)

  if (length(common_var) > 0) {
    ds_var_list <- expected_ds_var$dataset.var
    non_existed_ds_var <- ds_var_list[!ds_var_list %in% actual_ds_var]
  }

  if (all(!is.na(non_existed_ds_var))) {
    non_existed_var <- sapply(strsplit(
      x = non_existed_ds_var,
      split = "\\."
    ), function(x) {
      x[[2]]
    })
    checks <- expected_ds_var %>%
      filter(variable %in% non_existed_var) %>%
      mutate(common_var_status = case_when(
        dataset.var %in% non_existed_ds_var ~ FALSE,
        !dataset.var %in% non_existed_ds_var ~ TRUE
      )) %>%
      arrange(variable, dataset) %>%
      select(dataset, variable, common_var_status) %>%
      as_tibble()
  } else {
    checks <- NULL
  }
  if (nrow(checks) > 0 & is.data.frame(checks)) {
    if (show_mismatch) {
      return(checks)
    } else {
      stop("Mismatch common variables, set `show_mismatch = TRUE`")
    }
  }
  message("No mismatch common variables")
  return(NULL)
}
```

# Introduction

This document describes creating a metadata specification using [PHARMAVERSE](https://pharmaverse.org/) workflow. The matadata-specs can be used to create analysis ready dataset (i.e., ADaM). The matadata-specs required standardized derived dataset, such as SDTM, as input. For illustration purpose, a metadata-specs for following selected analysis ready dataset will be created.

* Subject-Level Analysis Dataset - ADSL
* Adverse Events Analysis Dataset - ADAE
* Analysis Dataset of Questionnaires - ADQS
* Analysis Dataset of Clinical Classification - ADRS

[metacore](https://atorus-research.github.io/metacore/) R package is used to create the metadata-specs.
 
# Load Required R packages

```{r setup-libraries, class.source = NULL, echo = TRUE}
library(tidyverse)
library(assertr)
library(labelled)
library(DT)
library(metacore)
```

# Building Metadata Specification

```{r prep-adsl-spec, class.source = NULL}
# ADSL Specs ----
adsl_dic <- bind_rows(
  c(variable = "STUDYID", type = "text", length = "4", label = "Study Identifier", common = TRUE, origin = "Predecessor", derivation = "DM.STUDYID"),
  c(variable = "USUBJID", type = "text", length = "20", label = "Unique Subject Identifier", common = TRUE, origin = "Predecessor", derivation = "DM.USUBJID"),
  c(variable = "SUBJID", type = "text", length = "15", label = "Subject Identifier for the Study", origin = "Predecessor", derivation = "DM.SUBJID"),
  c(variable = "SITEID", type = "text", length = "3", label = "Study Site Identifier", origin = "Predecessor", derivation = "DM.SITEID"),
  c(variable = "ORIGPROT", type = "text", length = "8", label = "Original Study Protocol", common = FALSE, code_id = "PHASE", origin = "Predecessor", derivation = "DM.ORIGPROT"),
  c(variable = "AGE", type = "float", length = "3", label = "Age (in Years)", common = TRUE, sig_dig = 1, origin = "Predecessor", derivation = "DM.AGE"),
  c(variable = "AGEU", type = "text", length = "5", label = "Age Units", common = TRUE, code_id = "AGEU", origin = "Predecessor", derivation = "DM.AGEU"),
  c(
    variable = "AGEGR1N", type = "integer", length = "1", label = "Pooled Age Group 1 (N)", common = FALSE, code_id = "AGEGR1N", origin = "Derived",
    derivation = "AGEGR1 = 1 if AGE <65. AGEGR1 = 2 if AGE >=65."
  ),
  c(
    variable = "AGEGR1", type = "text", length = "8", label = "Pooled Age Group 1", common = FALSE, code_id = "AGEGR1", origin = "Derived",
    derivation = "Character variable derived from ADSL.AGEGR1N"
  ),
  c(variable = "SEX", type = "text", length = "8", label = "Sex", common = TRUE, code_id = "SEX", origin = "Predecessor", derivation = "DM.SEX"),
  c(variable = "RACE", type = "text", length = "45", label = "Race", common = TRUE, code_id = "RACE", origin = "Predecessor", derivation = "DM.RACE"),
  c(variable = "RACEN", type = "integer", length = "3", label = "Race (N)", common = FALSE, code_id = "RACEN", origin = "Predecessor", derivation = "Numeric code for RACE"),
  c(variable = "ETHNIC", type = "text", length = "40", label = "Ethnicity", code_id = "ETHNIC", origin = "Predecessor", derivation = "DM.ETHNIC"),
  c(
    variable = "ENRLFL", type = "text", length = "3", label = "Enrolled Population Flag", common = TRUE, code_id = "YN", origin = "Derived",
    derivation = "Yes if ENRLDT is non-missing. No otherwise"
  ),
  c(variable = "DTHFL", type = "text", length = "3", label = "Death Flag", code_id = "Y_BLANK", origin = "Predecessor", derivation = "DM.DTHFL"),
  c(variable = "ARM", type = "text", length = "40", label = "Description of Planned Arm", origin = "Predecessor", derivation = "DM.ARM"),
  c(variable = "ACTARM", type = "text", length = "40", label = "Description of Actual Arm", origin = "Predecessor", derivation = "DM.ACTARM"),
  c(variable = "ENRLDT", type = "date", length = "10", label = "Date of Enrollment", format = "DATE10.", origin = "Predecessor", derivation = "DM.RFSTDTC"),
  # c(variable = "EOSSTT", type = "text", length = "30", label = "End of Study Status", code_id = "SBJTSTAT", origin = "Derived"),
  c(
    variable = "EOSDT", type = "date", length = "10", format = "DATE10.", label = "End of Study Date",
    common = TRUE, origin = "Predecessor", derivation = "DM.RFPENDTC"
  ),
  c(
    variable = "EOSDTF", type = "text", length = "1", label = "End of Study Date Imputation Flag", common = TRUE, origin = "Derived",
    derivation = "ADSL.EOSDTF is missing"
  ),
  # c(variable = "DCSREAS", type = "text", length = "60", label = "Reasons for Discontinuation from Study",   origin = "Derived"),
  # c(variable = "DCSREASP", type = "text", length = "60", label = "Reason Spec for Discont from Study",   origin = "Derived"),
  c(variable = "DTHDT", type = "date", length = "10", label = "Date of Death", format = "DATE10.", origin = "Predecessor", derivation = "DM.DTHDTC"),
  c(variable = "DTHDTF", type = "text", length = "1", label = "Date of Death Imputation Flag", origin = "Derived", derivation = "ADSL.DTHDT if days or month missing."),
  c(
    variable = "EDUC", type = "float", length = "2", label = "Education", sig_dig = 1, origin = "Derived",
    derivation = "SC.SCSTRESN where SC.SCTESTCD=EDUCAT and SCBLFL = 'Y'."
  ),
  c(
    variable = "MARISTAT", type = "text", length = "25", label = "Marital Status", origin = "Derived",
    derivation = "SC.SCORRES where SC.SCTESTCD=PTMARRY and SCBLFL = 'Y'."
  ),
  c(
    variable = "BMI", type = "float", length = "3", label = "Body Mass Index", sig_dig = 1, origin = "Derived",
    derivation = "Compute from VS.VSSTRESN where VS.VSTESTCD=WEIGHT and VS.VSTESTCD=HEIGHT, and VSBLFL = 'Y'."
  ),
  c(variable = "BMIBLU", type = "text", length = "8", label = "Body Mass Index Unit", code_id = "BMIBLU", origin = "Assigned"),
  c(
    variable = "DX", type = "text", length = "40", label = "Baseline Diagnostics Status", common = TRUE, code_id = "DX", origin = "Derived",
    derivation = "RS.RSSTRESC where RS.RSTESTCD=DX and RSBLFL = 'Y'."
  ),
  c(
    variable = "AMYSTAT", type = "text", length = "20", label = "Baseline Amyloid Status", code_id = "AMYSTAT",
    origin = "Predecessor",
    derivation = paste0(
      "Common amyloid status from NV.NVSTRESC where ",
      "NV.NVCAT=AMYLOIDPET & NV.NVTESTCD=AMYSTAT and NVBLFL = 'Y'."
    )
  ),
  c(
    variable = "AMYSTATC", type = "text", length = "20", label = "Baseline Amyloid Status - Composite Ref", code_id = "AMYSTAT",
    origin = "Predecessor",
    derivation = paste0(
      "Common amyloid status from NV.NVSTRESC where ",
      "NV.NVCAT=AMYLOIDPET & NV.NVTESTCD=AMYSTATC and NVBLFL = 'Y'."
    )
  ),
  c(
    variable = "TRACER", type = "text", length = "8", label = "Image Tracer Type",
    origin = "Predecessor", derivation = "Image tracer corresponding to both ADSL.AMYSTAT and ADSL.AMYSTATC"
  ),
  c(
    variable = "IMAGERES", type = "text", length = "8", label = "Image Resolution",
    origin = "Predecessor", derivation = "Image resolution corresponding to both ADSL.AMYSTAT and ADSL.AMYSTATC"
  ),
  c(
    variable = "ADASTT11", type = "float", length = "2", label = "Baseline ADAS-Cog Item 11 Total Score", sig_dig = 2,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=ADASTT11 and QSBLFL = 'Y'."
  ),
  c(
    variable = "ADASTT13", type = "float", length = "2", label = "Baseline ADAS-Cog Item 13 Total Score", sig_dig = 2,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=ADASTT13 and QSBLFL = 'Y'."
  ),
  c(
    variable = "CDGLOBAL", type = "float", length = "1", label = "Baseline CDR Global Score", sig_dig = 1,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=CDGLOBAL and QSBLFL = 'Y'."
  ),
  c(
    variable = "CDRSB", type = "float", length = "2", label = "Baseline CDR Sum of Boxes Score", sig_dig = 1,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=CDRSB and QSBLFL = 'Y'."
  ),
  c(
    variable = "ECOGPTTT", type = "float", length = "2", label = "Baseline Everyday Cognition - Pt Score", sig_dig = 2,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=ECOGPTTT and QSBLFL = 'Y'."
  ),
  c(
    variable = "ECOGSPTT", type = "float", length = "2", label = "Baseline Everyday Cognition - Sp Score", sig_dig = 2,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=ECOGSPTT and QSBLFL = 'Y'."
  ),
  c(
    variable = "FAQTOTAL", type = "float", length = "2", label = "Baseline FAQ Total Score", sig_dig = 2,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=FAQTOTAL and QSBLFL = 'Y'."
  ),
  c(
    variable = "FCISCORE", type = "float", length = "1", label = "Baseline FCI Score", sig_dig = 2,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=FCISCORE and QSBLFL = 'Y'."
  ),
  c(
    variable = "GDTOTAL", type = "float", length = "2", label = "Baseline GDS Score", sig_dig = 0,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=GDTOTAL and QSBLFL = 'Y'."
  ),
  c(
    variable = "MMSCORE", type = "float", length = "2", label = "Baseline MMSE Score", sig_dig = 0,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=MMSCORE and QSBLFL = 'Y'."
  ),
  c(
    variable = "MOCA", type = "float", length = "2", label = "Baseline MOCA Score", sig_dig = 0,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=MOCA and QSBLFL = 'Y'."
  ),
  c(
    variable = "NPITOTAL", type = "float", length = "2", label = "Baseline NPI Total Score", sig_dig = 2,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=NPITOTAL and QSBLFL = 'Y'."
  ),
  c(
    variable = "NPIQTOTL", type = "float", length = "1", label = "Baseline NPIQ Total Score", sig_dig = 2,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=NPIQTOTL and QSBLFL = 'Y'."
  ),
  c(
    variable = "DIGITSCR", type = "float", length = "2", label = "Baseline DIGIT Score", sig_dig = 2,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=DIGITSCR and QSBLFL = 'Y'."
  ),
  c(
    variable = "LDELTOTL", type = "float", length = "2", label = "Baseline LDEL Total Score", sig_dig = 2,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=LDELTOTL and QSBLFL = 'Y'."
  ),
  c(
    variable = "LIMMTOTL", type = "float", length = "2", label = "Baseline LIMM Total Score", sig_dig = 2,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=LIMMTOTL and QSBLFL = 'Y'."
  ),
  c(
    variable = "RAVLTFG", type = "integer", length = "2", label = "Baseline RAVLT Forgetting Score",
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=RAVLTFG and QSBLFL = 'Y'."
  ),
  c(
    variable = "RAVLTFGP", type = "float", length = "2", label = "Baseline RAVLT Percent Forgetting Score", sig_dig = 2,
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=RAVLTFG and QSBLFL = 'Y'."
  ),
  c(
    variable = "RAVLTIMM", type = "integer", length = "2", label = "Baseline RVAL Immediate Score",
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=RAVLTIMM and QSBLFL = 'Y'."
  ),
  c(
    variable = "RAVLTLRN", type = "integer", length = "2", label = "Baseline RVAL Learning Score",
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=RAVLTLRN and QSBLFL = 'Y'."
  ),
  c(
    variable = "TRABSCOR", type = "integer", length = "3", label = "Baseline Trails B Score",
    origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=TRABSCOR and QSBLFL = 'Y'."
  ),
  c(
    variable = "APOE", type = "text", length = "8", label = "APOE Genotype",
    origin = "Predecessor", derivation = "GF.GFSTRESC"
  )
)

adsl_dic <- adsl_dic %>%
  as_tibble() %>%
  {
    if (params$INCLUDE_PACC) {
      bind_rows(
        .,
        bind_rows(
          ., c(
            variable = "MPACCDIGIT", type = "float", length = "3", label = "Baseline mPACCdigit", sig_dig = 8,
            origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=MPACCDIGIT and QSBLFL = 'Y'."
          ),
          c(
            variable = "MPACCTRAILSB", type = "float", length = "3", label = "Baseline mPACCtrialsB", sig_dig = 8,
            origin = "Predecessor", derivation = "QS.QSSTRESC where QS.QSTESTCD=MPACCTRAILSB and QSBLFL = 'Y'."
          )
        )
      )
    } else {
      (.)
    }
  }

adsl_dic <- adsl_dic %>%
  mutate(
    dataset = "ADSL",
    structure = "One record per subject",
    dataset.label = "Subject-Level Analysis Dataset",
    key_seq = case_when(variable %in% "USUBJID" ~ 1),
    order = row_number(),
    keep = FALSE,
    core = case_when(
      variable %in% c(
        "STUDYID", "USUBJID", "SUBJID", "SITEID",
        "AGE", "AGEU", "SEX", "RACE", "ARM", "ENRLFL", "DX"
      ) ~ "Required",
      variable %in% c("DTHDTF") ~ "Conditionally Required",
      TRUE ~ "Permissible"
    ),
    supp_flag = FALSE,
    length = as.numeric(length),
    common = as.logical(ifelse(is.na(common), FALSE, common)),
    format = ifelse(is.na(format), NA_character_, format),
    sig_dig = as.integer(ifelse(is.na(sig_dig), NA_real_, sig_dig)),
    where = NA_character_,
    derivation_id = paste0(dataset, ".", variable)
  )
```

```{r prep-adae-spec, class.source = NULL}
# ADAE Specs ----
adae_dic <- bind_rows(
  c(variable = "STUDYID", type = "text", length = "4", label = "Study Identifier", common = TRUE, origin = "Predecessor", derivation = "ADSL.STUDYID"),
  c(variable = "USUBJID", type = "text", length = "20", label = "Unique Subject Identifier", common = TRUE, origin = "Predecessor", derivation = "ADSL.USUBJID"),
  c(
    variable = "SUBJID", type = "text", length = "15", label = "Subject Identifier for the Study", common = TRUE, origin = "Predecessor",
    derivation = "ADSL.SUBJID"
  ),
  c(variable = "SITEID", type = "text", length = "3", label = "Study Site Identifier", origin = "Predecessor", derivation = "ADSL.SITEID"),
  c(variable = "TRTA", type = "text", length = "40", label = "Actual Arm", origin = "Predecessor", derivation = "ADSL.ARM"),
  c(variable = "TRTP", type = "text", length = "40", label = "Planned Arm", origin = "Predecessor", derivation = "ADSL.ACTARM"),
  c(variable = "AGE", type = "float", length = "3", label = "Age (in Years)", common = TRUE, sig_dig = 1, origin = "Predecessor", derivation = "ADSL.AGE"),
  c(variable = "AGEU", type = "text", length = "5", label = "Age Units", common = TRUE, code_id = "AGEU", origin = "Predecessor", derivation = "ADSL.AGEU"),
  c(variable = "AGEGR1", type = "text", length = "8", label = "Pooled Age Group 1", code_id = "AGEGR1", origin = "Predecessor", derivation = "ADSL.AGEGR1"),
  c(variable = "AGEGR1N", type = "integer", length = "1", label = "Pooled Age Group 1 (N)", code_id = "AGEGR1N", origin = "Predecessor", derivation = "ADSL.AGEGR1N"),
  c(variable = "SEX", type = "text", length = "6", label = "Sex", common = TRUE, code_id = "SEX", origin = "Predecessor", derivation = "ADSL.SEX"),
  c(variable = "RACE", type = "text", length = "45", label = "Race", common = TRUE, code_id = "RACE", origin = "Predecessor", derivation = "ADSL.RACE"),
  c(variable = "RACEN", type = "integer", length = "3", label = "Race (N)", code_id = "RACEN", origin = "Predecessor", derivation = "ADSL.RACEN"),
  c(variable = "ETHNIC", type = "text", length = "40", label = "Ethnicity", code_id = "ETHNIC", origin = "Predecessor", derivation = "ADSL.ETHNIC"),
  c(variable = "DX", type = "text", length = "40", label = "Baseline Diagnostics Status", common = TRUE, code_id = "DX", origin = "Predecessor", derivation = "ADSL.DX"),
  c(variable = "ENRLFL", type = "text", length = "3", label = "Enrolled Population Flag", common = TRUE, code_id = "YN", origin = "Predecessor", derivation = "ADSL.ENRLFL"),
  c(variable = "ENRLDT", type = "date", length = "10", label = "Date of Enrollment", common = TRUE, format = "DATE10.", origin = "Predecessor", derivation = "ADSL.ENRLDT"),
  c(variable = "EOSDT", type = "date", length = "10", label = "End of Study Date", common = TRUE, format = "DATE10.", origin = "Predecessor", derivation = "ADSL.EOSDT"),
  c(variable = "EOSDTF", type = "text", length = "1", label = "End of Study Date Imputation Flag", common = TRUE, origin = "Predecessor", derivation = "ADSL.EOSDTF"),
  c(
    variable = "ASTDT", type = "date", length = "10", label = "Analysis Start Date", format = "DATE10.", origin = "Derived",
    derivation = paste0(
      "AE.AESTDTC, converted into date format. A maximum of month will be imputed for parital date.",
      "There are no events with completely missing start dates."
    )
  ),
  c(
    variable = "ASTDTF", type = "text", length = "1", label = "Analysis Start Date Imputation Flag", origin = "Assigned",
    derivation = "ASTDTF='D' if the day value within the character date is imputed. Note that only day values needed to be imputed for this study."
  ),
  c(
    variable = "ASTDY", type = "integer", length = "10", label = "Analysis Start Relative Day", origin = "Derived",
    derivation = "IF ASTDT>=ENRLDT then ASTDY=ASTDT-ENRLDT+1 Else if ENRLDT>ASTDT then ASTDY=ASTDT-ENRLDT"
  ),
  c(
    variable = "AENDT", type = "date", length = "1", label = "Analysis End Date", format = "DATE10.", origin = "Derived",
    derivation = "AE.AEENDTC, converted into date format. A maximum of month will be imputed for parital date."
  ),
  c(
    variable = "AENDTF", type = "text", length = "1", label = "Analysis End Date Imputation Flag", origin = "Assigned",
    derivation = "AENDT='M' if the day/month value within the character date is imputed."
  ),
  c(
    variable = "AENDY", type = "integer", length = "10", label = "Analysis End Relative Day", origin = "Derived",
    derivation = "IF AENDT>=ENRLDT then AENDY=AENDT-ENRLDT+1 Else if ENRLDT>AENDT then AENDY=AENDT-ENRLDT"
  ),
  c(
    variable = "ADURN", type = "integer", length = "1", label = "Adverse Events Duration (N)", origin = "Derived",
    derivation = "ADURN=AENDT-ASTDT+1"
  ),
  c(variable = "ADURU", type = "text", length = "8", label = "Analysis Duration Unit", origin = "Assigned", derivation = "If ADURN is not missing then ADURU='DAY'"),
  c(
    variable = "AETERM", type = "text", length = "200", label = "Reported Term for the Adverse Event", origin = "Predecessor",
    derivation = "AE.AETERM"
  ),
  c(variable = "AELLT", type = "text", length = "200", label = "Lowest Level Term", origin = "Predecessor", derivation = "AE.AELLT"),
  c(variable = "AELLTCD", type = "integer", length = "8", label = "Lowest Level Term Code", origin = "Predecessor", derivation = "AE.AELLTCD"),
  c(variable = "AEDECOD", type = "text", length = "200", label = "Dictionary-Derived Term", origin = "Predecessor", derivation = "AE.AEDECOD"),
  c(variable = "AEPTCD", type = "integer", length = "8", label = "Preferred Term Code", origin = "Predecessor", derivation = "AE.AEPTCD"),
  c(variable = "AEHLT", type = "text", length = "200", label = "High Level Term", origin = "Predecessor", derivation = "AE.AEHLT"),
  c(variable = "AEHLTCD", type = "integer", length = "8", label = "High Level Term Code", origin = "Predecessor", derivation = "AE.AEHLTCD"),
  c(variable = "AEHLGT", type = "text", length = "200", label = "High Level Group Term", origin = "Predecessor", derivation = "AE.AEHLGT"),
  c(variable = "AEHLGTCD", type = "integer", length = "8", label = "High Level Group Term Code", origin = "Predecessor", derivation = "AE.AEHLGTCD"),
  c(variable = "AESOC", type = "text", length = "200", label = "Primary Sytem Organ Class", origin = "Predecessor", derivation = "AE.AESOC"),
  c(variable = "AESOCCD", type = "integer", length = "8", label = "Primary Sytem Organ Class Code", origin = "Predecessor", derivation = "AE.AESOCCD"),
  c(variable = "AESEV", type = "text", length = "8", label = "Severity/Intensity", code_id = "AESEV", origin = "Predecessor", derivation = "AE.AESEV"),
  c(variable = "AESER", type = "text", length = "3", label = "Serious Event", code_id = "YN", origin = "Predecessor", derivation = "AE.AESER"),
  c(variable = "AESCONG", type = "text", length = "3", label = "Congenital Anomaly or Birth Defect", code_id = "YN", origin = "Predecessor", derivation = "AE.AESCONG"),
  c(variable = "AESDISAB", type = "text", length = "3", label = "Persist or Signif Disability/Incapacity", code_id = "YN", origin = "Predecessor", derivation = "AE.AESDISAB"),
  c(variable = "AESDTH", type = "text", length = "3", label = "Results in Death", code_id = "YN", origin = "Predecessor", derivation = "AE.AESDTH"),
  c(variable = "AESHOSP", type = "text", length = "3", label = "Requires or Prologns Hospitalization", code_id = "YN", origin = "Predecessor", derivation = "AE.AESHOSP"),
  c(variable = "AESLIFE", type = "text", length = "3", label = "Is Life Threating", code_id = "YN", origin = "Predecessor", derivation = "AE.AESLIFE"),
  c(variable = "AESMIE", type = "text", length = "3", label = "Other Medically Important Serious Event", origin = "Predecessor", derivation = "AE.AESMIE"),
  c(variable = "AERELCM", type = "text", length = "40", label = "Related to Concomitant Therapy", code_id = "REL", origin = "Predecessor", derivation = "AE.AERELCM"),
  c(variable = "AERELFLRBTBN", type = "text", length = "40", label = "Related to Florbetaben Tracer", code_id = "REL", origin = "Predecessor", derivation = "AE.AERELFLRBTBN"),
  c(variable = "AERELFLRBPR", type = "text", length = "40", label = "Related to Florbetapir Tracer", code_id = "REL", origin = "Predecessor", derivation = "AE.AERELFLRBPR"),
  c(variable = "AEHIMG", type = "text", length = "40", label = "Related to PET/MRI Imaging Procedure", code_id = "REL", origin = "Predecessor", derivation = "AE.AEHIMG"),
  c(variable = "AERELTAU", type = "text", length = "40", label = "Related to Flotaucipir Tracer", code_id = "REL", origin = "Predecessor", derivation = "AE.AERELTAU"),
  c(variable = "AERELNAV", type = "text", length = "40", label = "Related to NAV-4694 Tracer", code_id = "REL", origin = "Predecessor", derivation = "AE.AERELNAV"),
  c(variable = "AERELMK", type = "text", length = "40", label = "Related to MK-6240 Tracer", code_id = "REL", origin = "Predecessor", derivation = "AE.AERELMK"),
  c(variable = "AERELPI", type = "text", length = "40", label = "Related to PI-2620 Tracer", code_id = "REL", origin = "Predecessor", derivation = "AE.AERELPI"),
  c(variable = "AEHLUMB", type = "text", length = "40", label = "Related to Lumbar Puncture", code_id = "REL", origin = "Predecessor", derivation = "AE.AEHLUMB"),
  c(variable = "AERELCOVID", type = "text", length = "40", label = "Related to COVID-19 Illness", code_id = "RELCOVID", origin = "Predecessor", derivation = "AE.AERELCOVID"),
  c(variable = "AERELPAN", type = "text", length = "40", label = "Related to COVID-19 Pandemic Disruption", code_id = "REL", origin = "Predecessor", derivation = "AE.AERELPAN"),
  c(variable = "AERELAD", type = "text", length = "40", label = "Related to Early Stage/Progression of AD", code_id = "REL", origin = "Predecessor", derivation = "AE.AERELAD"),
  c(variable = "AERELATESP", type = "text", length = "40", label = "Related to Other Study Procedure(s)", code_id = "REL", origin = "Predecessor", derivation = "AE.AERELATESP"),
  c(variable = "AECONTRT", type = "text", length = "3", label = "Concomitant or Additional Trtmnt Given", code_id = "YN", origin = "Predecessor", derivation = "AE.AECONTRT"),
  c(variable = "AEOUT", type = "text", length = "60", label = "Outcome of Adverse Event", origin = "Predecessor", derivation = "AE.AEOUT"),
  c(variable = "AESEQ", type = "integer", length = "8", label = "Sequence Number", origin = "Predecessor", derivation = "AE.AESEQ")
) %>%
  as_tibble() %>%
  mutate(
    dataset = "ADAE",
    dataset.label = "Analysis Dataset of Adverse Events",
    structure = "One record per adverse event per subject",
    key_seq = case_when(
      variable %in% "USUBJID" ~ 1,
      variable %in% "AEOUT" ~ 2,
      variable %in% "ASTDT" ~ 3,
      variable %in% "AESEQ" ~ 4
    ),
    order = row_number(),
    keep = FALSE,
    core = case_when(
      variable %in% c("STUDYID", "USUBJID", "AEOUT", "AESEV", "AESER") ~ "Required",
      variable %in% c("ASTDTF", "AENDTF") ~ "Conditionally Required",
      TRUE ~ "Permissible"
    ),
    supp_flag = FALSE,
    length = as.numeric(length),
    common = as.logical(ifelse(is.na(common), FALSE, common)),
    format = ifelse(is.na(format), NA_character_, format),
    sig_dig = as.integer(ifelse(is.na(sig_dig), NA_real_, sig_dig)),
    where = NA_character_,
    derivation_id = paste0(dataset, ".", variable),
  )
```

```{r prep-adqs-spec, class.source = NULL}
# ADQS Specs ----
adqs_dic <- bind_rows(
  c(variable = "STUDYID", type = "text", length = "4", label = "Study Identifier", common = TRUE, origin = "Predecessor", derivation = "ADSL.STUDYID"),
  c(variable = "USUBJID", type = "text", length = "20", label = "Unique Subject Identifier", common = TRUE, origin = "Predecessor", derivation = "ADSL.USUBJID"),
  c(variable = "SUBJID", type = "text", length = "15", label = "Subject Identifier for the Study", common = TRUE, origin = "Predecessor", derivation = "ADSL.SUBJID"),
  c(variable = "SITEID", type = "text", length = "3", label = "Study Site Identifier", origin = "Predecessor", derivation = "ADSL.SITEID"),
  c(variable = "AGE", type = "float", length = "3", label = "Age (in Years)", common = TRUE, sig_dig = 1, origin = "Predecessor", derivation = "ADSL.AGE"),
  c(variable = "AGEU", type = "text", length = "5", label = "Age Units", common = TRUE, code_id = "AGEU", origin = "Predecessor", derivation = "ADSL.AGEU"),
  c(variable = "AGEGR1", type = "text", length = "8", label = "Pooled Age Group 1", code_id = "AGEGR1", origin = "Predecessor", derivation = "ADSL.AGEGR1"),
  c(variable = "AGEGR1N", type = "integer", length = "1", label = "Pooled Age Group 1 (N)", code_id = "AGEGR1N", origin = "Predecessor", derivation = "ADSL.AGEGR1N"),
  c(variable = "SEX", type = "text", length = "6", label = "Sex", common = TRUE, code_id = "SEX", origin = "Predecessor", derivation = "ADSL.SEX"),
  c(variable = "RACE", type = "text", length = "45", label = "Race", common = TRUE, code_id = "RACE", origin = "Predecessor", derivation = "ADSL.RACE"),
  c(variable = "BMI", type = "float", length = "3", label = "Body Mass Index", sig_dig = 2, origin = "Predecessor", derivation = "ADSL.BMI"),
  c(variable = "BMIBLU", type = "text", length = "8", label = "Body Mass Index Unit", origin = "Predecessor", derivation = "ADSL.BMIBLU"),
  c(variable = "DX", type = "text", length = "40", label = "Baseline Diagnostics Status", common = TRUE, code_id = "DX", origin = "Predecessor", derivation = "ADSL.DX"),
  c(variable = "AMYSTAT", type = "text", length = "40", label = "Baseline Amyloid Status", code_id = "AMYSTAT", origin = "Predecessor", derivation = "ADSL.AMYSTAT"),
  c(variable = "ENRLFL", type = "text", length = "3", label = "Enrolled Population Flag", common = TRUE, code_id = "YN", origin = "Predecessor", derivation = "ADSL.ENRLFL"),
  c(variable = "ENRLDT", type = "date", length = "10", label = "Date of Enrollment", common = TRUE, format = "DATE10.", origin = "Predecessor", derivation = "ADSL.ENRLDT"),
  c(variable = "EOSDT", type = "date", length = "10", label = "End of Study Date", common = TRUE, format = "DATE10.", origin = "Predecessor", derivation = "ADSL.EOSDT"),
  c(variable = "EOSDTF", type = "text", length = "1", label = "End of Study Date Imputation Flag", common = TRUE, origin = "Predecessor", derivation = "ADSL.EOSDTF"),
  #  TRTA and TRTP are included for illustration purpose
  c(variable = "TRTA", type = "text", length = "40", label = "Actual Arm", origin = "Predecessor", derivation = "ADSL.ARM"),
  c(variable = "TRTP", type = "text", length = "40", label = "Planned Arm", origin = "Predecessor", derivation = "ADSL.ACTARM"),
  c(variable = "AVISIT", type = "text", length = "10", label = "Analysis Visit", origin = "Derived"),
  c(variable = "AVISITN", type = "integer", length = "3", label = "Analysis Visit (N)", origin = "Derived"),
  c(variable = "VISIT", type = "text", length = "10", label = "Visit Name", origin = "Predecessor", derivation = "QS.VISIT"),
  c(variable = "VISITNUM", type = "integer", length = "3", label = "Visit Number", origin = "Predecessor", derivation = "QS.VISITNUM"),
  c(variable = "ADY", type = "integer", length = "40", label = "Analysis Relative Day", origin = "Derived", derivation = "ADQS.ENRLDT - ADQS.ADT"),
  c(variable = "ADT", type = "date", length = "10", label = "Analysis Date", format = "DATE10.", origin = "Predecessor", derivation = "QS.QSDTC"),
  c(variable = "PARAM", type = "text", length = "200", label = "Parameter", origin = "Predecessor", derivation = "QS.QSTEST"),
  c(variable = "PARAMCD", type = "text", length = "40", label = "Parameter Code", code_id = "PARAMCD", origin = "Predecessor", derivation = "QS.QSTESTCD"),
  c(variable = "PARAMN", type = "integer", length = "10", label = "Parameter (N)", code_id = "PARAMN", origin = "Derived", derivation = "Derived from ADQS.PARAMCD"),
  c(variable = "AVAL", type = "float", length = "10", label = "Analysis Value", sig_dig = 5, origin = "Predecessor", derivation = "QS.QSSTRESN"),
  c(variable = "AVALC", type = "text", length = "40", label = "Analysis Value (C)", origin = "Predecessor", derivation = "QS.QSSTRESC"),
  c(variable = "BASE", type = "float", length = "10", label = "Baseline Value", sig_dig = 5, origin = "Derived", derivation = "QS.QSORRES if QS.QSBLFL is 'Y'"),
  c(variable = "CHG", type = "float", length = "10", label = "Change from Baseline", sig_dig = 5, origin = "Derived", derivation = "ADQS.AVAL - ADQS.BASE"),
  c(
    variable = "PCHG", type = "float", length = "4", label = "Percent Change from Baseline", sig_dig = 2, origin = "Derived",
    derivation = "(ADQS.AVAL - ADQS.BASE)/ADQS.BASE. NULL if ADQS.BASE is missing."
  ),
  c(variable = "ABLFL", type = "text", length = "3", label = "Baseline Record Flag", code_id = "Y_BLANK", origin = "Predecessor", derivation = "QS.QSBLFL"),
  c(variable = "QSSEQ", type = "integer", length = "8", label = "Sequence Number", origin = "Predecessor", derivation = "QS.QSSEQ")
) %>%
  as_tibble() %>%
  mutate(
    dataset = "ADQS",
    dataset.label = "Analysis Dataset of Questionnaires",
    structure = "One record per subject per parameter per analysis visit per analysis date",
    key_seq = case_when(
      variable %in% "USUBJID" ~ 1,
      variable %in% "PARAMCD" ~ 2,
      variable %in% "AVISIT" ~ 3,
      variable %in% "ADT" ~ 4
    ),
    order = row_number(),
    keep = FALSE,
    core = case_when(
      variable %in% c("STUDYID", "USUBJID", "PARAM", "PARAMCD", "AVISIT") ~ "Required",
      variable %in% c("AVAL", "AVALC", "DTYPE") ~ "Conditionally Required",
      TRUE ~ "Permissible"
    ),
    supp_flag = FALSE,
    length = as.numeric(length),
    common = as.logical(ifelse(is.na(common), FALSE, common)),
    format = ifelse(is.na(format), NA_character_, format),
    sig_dig = as.integer(ifelse(is.na(sig_dig), NA_real_, sig_dig)),
    where = NA_character_,
    derivation_id = paste0(dataset, ".", variable)
  )
```

```{r prep-adrs-spec, class.source = NULL}
# ADRS Specs ----
# ADRS is created mainly as input for clinical progression summary Vignette
adrs_dic <- bind_rows(
  c(variable = "STUDYID", type = "text", length = "4", label = "Study Identifier", common = TRUE, origin = "Predecessor", derivation = "ADSL.STUDYID"),
  c(variable = "USUBJID", type = "text", length = "20", label = "Unique Subject Identifier", common = TRUE, origin = "Predecessor", derivation = "ADSL.USUBJID"),
  c(variable = "SUBJID", type = "text", length = "15", label = "Subject Identifier for the Study", common = TRUE, origin = "Predecessor", derivation = "ADSL.SUBJID"),
  c(variable = "COLPROT", type = "text", length = "15", label = "Study Protocol of Data Collection", origin = "Predecessor", derivation = "RS.RSGRPID"),
  c(
    variable = "PARAMCD", type = "text", length = "40", label = "Parameter Code", code_id = "PARAMCD", origin = "Derived",
    derivation = "RS.QSTESTCD = 'DX' and Death if ADSL.DTHFL = 'Yes'"
  ),
  c(
    variable = "PARAM", type = "text", length = "200", label = "Parameter", origin = "Derived",
    derivation = "QS.QSTEST and Death if ADSL.DTHFL = 'Yes'"
  ),
  c(
    variable = "PARAMN", type = "integer", length = "10", label = "Parameter (N)", code_id = "PARAMN", origin = "Derived",
    derivation = "Derived from ADRS.PARAMCD"
  ),
  c(variable = "AVISIT", type = "text", length = "10", label = "Analysis Visit", origin = "Derived"),
  c(variable = "AVISITN", type = "integer", length = "3", label = "Analysis Visit (N)", origin = "Derived"),
  c(variable = "VISIT", type = "text", length = "10", label = "Visit Name", origin = "Predecessor", derivation = "RS.VISIT"),
  c(variable = "VISITNUM", type = "integer", length = "3", label = "Visit Number", origin = "Predecessor", derivation = "RS.VISITNUM"),
  c(variable = "ADY", type = "integer", length = "40", label = "Analysis Relative Day", origin = "Derived"),
  c(
    variable = "ADT", type = "date", length = "10", label = "Analysis Date", format = "DATE10.", origin = "Derived",
    derivation = "RS.RSDTC and ADSL.DTHDT with date imputation"
  ),
  c(
    variable = "AVAL", type = "float", length = "10", label = "Analysis Value", sig_dig = 5, origin = "Derived",
    derivation = "RS.RSORRES and 'DEATH' if ADSL.DTHDT is non-missing"
  ),
  c(variable = "AVALC", type = "text", length = "40", label = "Analysis Value (C)", origin = "Predecessor", derivation = "ADRS.AVAL"),
  c(variable = "BASE", type = "float", length = "10", label = "Baseline Value", sig_dig = 5, origin = "Derived", derivation = "RS.RSORRES if RS.RSBLFL is 'Y'"),
  c(variable = "ABLFL", type = "text", length = "3", label = "Baseline Record Flag", code_id = "Y_BLANK", origin = "Predecessor", derivation = "RS.RSBLFL"),
  c(variable = "ENRLFL", type = "text", length = "3", label = "Enrolled Population Flag", common = TRUE, code_id = "YN", origin = "Predecessor", derivation = "ADSL.ENRLFL"),
  c(variable = "ENRLDT", type = "date", length = "10", label = "Date of Enrollment", common = TRUE, format = "DATE10.", origin = "Predecessor", derivation = "ADSL.ENRLDT"),
  c(variable = "EOSDT", type = "date", length = "10", label = "End of Study Date", common = TRUE, format = "DATE10.", origin = "Predecessor", derivation = "ADSL.EOSDT"),
  c(variable = "EOSDTF", type = "text", length = "1", label = "End of Study Date Imputation Flag", common = TRUE, origin = "Predecessor", derivation = "ADSL.EOSDTF"),
  # Additional flags used in the vignettes
  c(
    variable = "BLBFL", type = "text", length = "3", label = "Baseline Records Population Flag", code_id = "Y_BLANK", origin = "Derived",
    derivation = "Subjects that have a baseline record"
  ),
  c(
    variable = "FOLLOWPFL", type = "text", length = "3", label = "Follow-up Records Population Flag", origin = "Derived",
    derivation = "Subjects that have at least one diagnostics summary after baseline visit or death records"
  ),
  c(
    variable = "ANL01FL", type = "text", length = "3", label = "Analysis Flag", common = FALSE, code_id = "YN", origin = "Derived",
    derivation = "Enrolled subjects that have baseline record, at least one diagnostics summary after baseline visit or death records"
  ),
  c(variable = "AGE", type = "float", length = "3", label = "Age (in Years)", common = TRUE, sig_dig = 1, origin = "Predecessor", derivation = "ADSL.AGE"),
  c(variable = "AGEU", type = "text", length = "5", label = "Age Units", common = TRUE, code_id = "AGEU", origin = "Predecessor", derivation = "ADSL.AGEU"),
  c(variable = "AGEGR1", type = "text", length = "8", label = "Pooled Age Group 1", code_id = "AGEGR1", origin = "Predecessor", derivation = "ADSL.AGEGR1"),
  c(variable = "AGEGR1N", type = "integer", length = "1", label = "Pooled Age Group 1 (N)", code_id = "AGEGR1N", origin = "Predecessor", derivation = "ADSL.AGEGR1N"),
  c(variable = "SEX", type = "text", length = "6", label = "Sex", common = TRUE, code_id = "SEX", origin = "Predecessor", derivation = "ADSL.SEX"),
  c(variable = "RACE", type = "text", length = "45", label = "Race", common = TRUE, code_id = "RACE", origin = "Predecessor", derivation = "ADSL.RACE"),
  c(variable = "BMI", type = "float", length = "3", label = "Body Mass Index", sig_dig = 2, origin = "Predecessor", derivation = "ADSL.BMI"),
  c(variable = "BMIBLU", type = "text", length = "8", label = "Body Mass Index Unit", origin = "Predecessor", derivation = "ADSL.BMIBLU"),
  c(variable = "DX", type = "text", length = "40", label = "Baseline Diagnostics Status", common = TRUE, code_id = "DX", origin = "Predecessor", derivation = "ADSL.DX"),
  c(variable = "AMYSTAT", type = "text", length = "40", label = "Baseline Amyloid Status", code_id = "AMYSTAT", origin = "Predecessor", derivation = "ADSL.AMYSTAT")
) %>%
  as_tibble() %>%
  mutate(
    dataset = "ADRS",
    dataset.label = "Response Analysis Dataset",
    structure = "One record per subject per parameter per analysis visit per analysis date",
    key_seq = case_when(
      variable %in% "USUBJID" ~ 1,
      variable %in% "PARAMCD" ~ 2,
      variable %in% "AVISIT" ~ 3,
      variable %in% "ADT" ~ 4
    ),
    order = row_number(),
    keep = FALSE,
    core = case_when(
      variable %in% c("STUDYID", "USUBJID", "PARAM", "PARAMCD", "AVISIT") ~ "Required",
      variable %in% c("AVAL", "AVALC") ~ "Conditionally Required",
      TRUE ~ "Permissible"
    ),
    supp_flag = FALSE,
    length = as.numeric(length),
    common = as.logical(ifelse(is.na(common), FALSE, common)),
    format = ifelse(is.na(format), NA_character_, format),
    sig_dig = as.integer(ifelse(is.na(sig_dig), NA_real_, sig_dig)),
    where = NA_character_,
    derivation_id = paste0(dataset, ".", variable)
  )
```

```{r prep-dataset-spec, echo = TRUE}
# Based on manual codes
data_spec <- bind_rows(adsl_dic, adae_dic, adqs_dic, adrs_dic)
```

## Dataset-Level Information

```{r ds-spec, echo = TRUE}
ds_spec_labels <- c(
  dataset = "Dataset Name",
  structure = "Value Structure",
  label = "Dataset Label"
)

ds_spec <- data_spec %>%
  select(dataset, structure, label = dataset.label) %>%
  distinct() %>%
  assert_non_missing(everything()) %>%
  labelled::set_variable_labels(.labels = ds_spec_labels, .strict = TRUE)
```

```{r print-ds-spec}
datatable(ds_spec, paging = FALSE)
```

## Dataset and Variable Information

```{r ds-vars, echo = TRUE}
ds_vars_labels <- c(
  dataset = "Dataset Name",
  variable = "Variable Name",
  key_seq = "Sequence Key",
  order = "Variable Order",
  keep = "Keep (Boolean)",
  core = "ADaM core (Expected, Required, Permissible)",
  supp_flag = "Supplemental Flag"
)
ds_vars <- data_spec %>%
  select(dataset, variable, key_seq, order, keep, core, supp_flag) %>%
  assert_non_missing(!key_seq) %>%
  labelled::set_variable_labels(.labels = ds_vars_labels, .strict = TRUE)
```

```{r print-ds-vars}
datatable(
  ds_vars %>%
    mutate(across(where(is.logical), ~ toupper(.x))),
  paging = TRUE
)
```

## Variable-Level Information

```{r var-spec, echo = TRUE}
var_spec_labels <- c(
  variable = "Variable Name",
  length = "Variable Length",
  label = "Variable Label",
  type = "Variable Class",
  common = "Common Across ADaM",
  format = "Variable Format"
)
var_spec <- data_spec %>%
  select(all_of(names(var_spec_labels))) %>%
  assert_non_missing(!format) %>%
  mutate(type = toupper(type)) %>%
  labelled::set_variable_labels(.labels = var_spec_labels, .strict = TRUE)
```

```{r print-var-spec}
datatable(
  var_spec %>%
    mutate(across(where(is.logical), ~ toupper(.x))),
  paging = TRUE
)
```

## Value-Level Information

* Assumed value-level information are similar to variable level information. 

```{r value-spec, echo = TRUE}
value_spec_labels <- c(
  dataset = "Dataset Name",
  variable = "Variable Name",
  type = "Value Type",
  origin = "Value Source",
  sig_dig = "Significant Digits for Numeric Value",
  code_id = "ID of the Code List",
  where = "Value of the Variable",
  derivation_id = "ID of Derivation"
)
value_spec <- data_spec %>%
  select(all_of(names(value_spec_labels))) %>%
  assert_non_missing(dataset, variable, origin) %>%
  labelled::set_variable_labels(.labels = value_spec_labels, .strict = FALSE)
```

```{r print-value-spec}
datatable(value_spec, paging = TRUE)
```

## Derivations 

```{r derivations-spec, echo = TRUE}
derivations_labels <- c(
  derivation_id = "ID of Derivation",
  derivation = "Derivation"
)
derivations <- data_spec %>%
  select(all_of(names(derivations_labels))) %>%
  assert_non_missing(derivation_id)
```

```{r print-derivations-spec}
datatable(derivations, paging = TRUE)
```

## Code List 

```{r codelist-spec, class.source = NULL}
# Pre-specified variables with coded values, illustration purpose
# Age Group ----
AGEGR1_CODEDLIST <- tibble(
  code_id = "AGEGR1", name = "AGEGR1", type = "code_decode",
  tibble(
    code = c("<65", ">=65"),
    decode = c("<65", ">=65")
  )
)
AGEGR1N_CODEDLIST <- tibble(
  code_id = "AGEGR1N", name = "AGEGR1N", type = "code_decode",
  tibble(
    code = c("1", "2"),
    decode = c("<65", ">=65")
  )
)

AGEU_CODEDLIST <- tibble(
  code_id = "AGEU", name = "AGEU", type = "code_decode",
  tibble(
    code = c("Years"),
    decode = c("Years")
  )
)

# PHASE/ORIGPROT ----
PHASE_CODEDLIST <- tibble(
  code_id = "PHASE", name = "PHASE", type = "code_decode",
  tibble(
    code = c("ADNI1", "ADNIGO", "ADNI2", "ADNI3", "ADNI4"),
    decode = c("ADNI1", "ADNIGO", "ADNI2", "ADNI3", "ADNI4")
  )
)

# Sex ----
SEX_CODEDLIST <- tibble(
  code_id = "SEX", name = "SEX", type = "code_decode",
  tibble(
    code = c("Female", "Male"),
    decode = c("Female", "Male")
  )
)

# Race -----
race_code_value <- c(
  "American Indian or Alaskan Native", "Asian", "Black or African American",
  "Native Hawaiian or Other Pacific Islander", "Other Pacific Islander",
  "White", "More than one race", "Unknown"
)

RACE_CODEDLIST <- tibble(
  code_id = "RACE", name = "RACE", type = "code_decode",
  tibble(
    code = race_code_value,
    decode = race_code_value
  )
)

race_code_num <- c(seq_along(race_code_value)[-length(race_code_value)], 999)
RACEN_CODEDLIST <- tibble(
  code_id = "RACEN", name = "RACEN", type = "code_decode",
  tibble(
    code = as.character(race_code_num),
    decode = race_code_value
  )
)

# Ethnicity ----
ETHNIC_CODEDLIST <- tibble(
  code_id = "ETHNIC", name = "ETHNIC", type = "code_decode",
  tibble(
    code = c("Hispanic or Latino", "Not Hispanic or Latino", "Unknown"),
    decode = c("Hispanic or Latino", "Not Hispanic or Latino", "Unknown")
  )
)

# Diagnostics status ----
DX_CODEDLIST <- tibble(
  code_id = "DX", name = "DX", type = "code_decode",
  tibble(
    code = c("CN", "MCI", "DEM"),
    decode = c(
      "Cognitive Normal",
      "Mild Cognitive Impairment",
      "Dementia/Alzheimer's"
    )
  )
)

# Amyloid Status ----
AMYSTAT_CODEDLIST <- tibble(
  code_id = "AMYSTAT", name = "AMYSTAT", type = "code_decode",
  tibble(
    code = c("Elevated", "Non Elevated"),
    decode = c("Elevated", "Non Elevated")
  )
)

# BMI Units ----
BMIBLU_CODEDLIST <- tibble(
  code_id = "BMIBLU", name = "BMIBLU", type = "code_decode",
  tibble(
    code = c("kg/m^2"),
    decode = c("kg/m^2")
  )
)

# Adverse Events Possible Relation ----
REL_CODEDLIST <- tibble(
  code_id = "REL", name = "REL", type = "code_decode",
  tibble(
    code = c("1", "2", "3", "4", "5", "999"),
    decode = c(
      "Definitely", "Probably", "Possibly",
      "Unlikely", "Not related", "Uknown"
    )
  )
)

RELCOVID_CODEDLIST <- tibble(
  code_id = "RELCOVID", name = "RELCOVID", type = "code_decode",
  tibble(
    code = c("1", "2", "3", "4", "5", "999"),
    decode = c(
      "Definitely (Confirmed COVID-19 test)", "Probably", "Possibly",
      "Unlikely", "Not related", "Uknown"
    )
  )
)

# AE Severity  ---
AESEV_CODEDLIST <- tibble(
  code_id = "AESEV", name = "AESEV", type = "code_decode",
  tibble(
    code = c("1", "2", "3"),
    decode = c("Mild", "Moderate", "Severe")
  )
)

# Parameter code list ----
QSTESTCD_LIST <- c(
  "ADASTT11", "ADASTT13", "CDGLOBAL", "CDRSB", "DIGITSCR",
  "ECOGPTTT", "ECOGSPTT", "FAQTOTAL", "FCISCORE", "GDTOTAL", "LDELTOTL",
  "LIMMTOTL", "MMSCORE", "MOCA", "MPACCDIGIT", "MPACCTRAILSB",
  "NPIQTOTL", "NPITOTAL", "RAVLTFG", "RAVLTFGP", "RAVLTIMM",
  "RAVLTLRN", "TRABSCOR"
)
# To apply with data flow
# QSTESTCD_LIST <- QS %>%
#   select(QSTESTCD) %>%
#   distinct() %>%
#   drop_na() %>%
#   pull()

PARAMCD_CODEDLIST <- tibble(
  code_id = "PARAMCD", name = "PARAMCD", type = "code_decode",
  tibble(
    code = QSTESTCD_LIST,
    decode = QSTESTCD_LIST
  )
)

PARAMN_CODEDLIST <- tibble(
  code_id = "PARAMN", name = "PARAMN", type = "code_decode",
  tibble(
    code = as.character(seq_along(QSTESTCD_LIST)),
    decode = QSTESTCD_LIST
  )
)

# YN ----
YN_CODEDLIST <- tibble(
  code_id = "YN", name = "YN", type = "code_decode",
  tibble(
    code = c("Y", "N"),
    decode = c("Yes", "No")
  )
)

# Y_BLANK ----
Y_BLANK_CODEDLIST <- tibble(
  code_id = "Y_BLANK", name = "Y_BLANK", type = "code_decode",
  tibble(
    code = c("Y"),
    decode = c("Yes")
  )
)

meta_coded_id <- data_spec$code_id[!is.na(data_spec$code_id)]

codelist <- mget(ls()[str_detect(ls(), "_CODEDLIST")]) %>%
  bind_rows() %>%
  group_by(code_id, name, type) %>%
  nest(codes = all_of(c("code", "decode"))) %>%
  ungroup() %>%
  verify(all(code_id %in% meta_coded_id)) %>%
  verify(all(meta_coded_id %in% code_id))
```

```{r print-codelist-spec}
datatable(
  codelist %>%
    unnest(codes),
  paging = TRUE
)
```

## Metadata Specification

```{r meta-spec, echo = TRUE, message = TRUE, warning = FALSE}
# Generate Metadata Specs -  R6 class wrapper object
METACORES <- metacore(
  ds_spec = ds_spec,
  ds_vars = ds_vars,
  var_spec = var_spec,
  value_spec = value_spec,
  derivations = derivations,
  codelist = codelist
)
# Checking for mismatch labels
check_inconsistent_labels(METACORES) %>%
  is.null()
# Checking for mismatch format
check_inconsistent_formats(METACORES) %>%
  is.null()
# Checking for mismatch type
check_inconsistent_types(METACORES) %>%
  is.null()

# Checking for mismatch common variables: user defined function
# Please see the source file for more information
check_inconsistent_common(METACORES) %>%
  is.null()
```
